### 2021-06-10 analysis_eTMB.Rmd
Initial example

---
title: "Analysis of eTMB"
output:
  pdf_document: default
  code_folding: default
  html_document:
    code_folding: hide
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(MutationalPatterns)
library(NMF)
library(reshape2)
library(grid)
library(gridExtra)
```

## Preliminary functions

```{r}
# Loads a mutational profile with columns sample, ext.context and num
# Subsets such file to samples in sample.list and generates a matrix
# with the mutational profile that can be used by MutationalPatterns
load_mutational_matrix<- function(filename, sample.list) {

  mutational.signature <-  read_tsv(filename)
  order.tri <-  unique(mutational.signature$ext.context)
  sample.names <-  unique(mutational.signature$sample)

  mutational.signature <-  mutational.signature %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    filter(sample %in% sample.list) %>%
    select(sample,ext.context,num) %>%
    spread(sample,num) %>%
    select(one_of("ext.context",sample.names))

  mut.matrix <- as.matrix(sapply(mutational.signature[,2:ncol(mutational.signature)], as.numeric))  
  rownames(mut.matrix) <-  mutational.signature$ext.context

  # Add 1e-04 for use with NMF
  mut.matrix <-  mut.matrix + 1e-04

  return (mut.matrix)
}

load_old_mutational_matrix<- function(filename,mut.order="supplied_results/mut_sig.order.txt") {

  mutational.signature = read_tsv(filename)
  mut_mat = read_tsv(mut.order)

  mut.join = mutational.signature %>%
    inner_join(mut_mat)

  order.tri = unique(mut.join$ext.context)

  mut.join = mut.join %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    select(ext.context,prob)

  mut.matrix=as.matrix(sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
  rownames(mut.matrix) = mut.join$ext.context

  return (t(mut.matrix)[,1:96])
}

# Loads the published COSMIC mutational signatures in a format that can be
# used by MutationalPatterns
load_cosmic_matrix<- function(mut.order="supplied_results/mut_sig.order.txt") {

  sp_url <- paste("https://cancer.sanger.ac.uk/cancergenome/assets/",
                "signatures_probabilities.txt", sep = "")

  cancer_signatures <-  read.table(sp_url, sep = "\t", header = TRUE)
  # Match the order of the mutation types to MutationalPatterns standard
  mut_mat <-  read_tsv(mut.order)
  new_order <-  match(mut_mat$ext.context, cancer_signatures$Somatic.Mutation.Type) 
  # Reorder cancer signatures dataframe
  cancer_signatures <-  cancer_signatures[as.vector(new_order),]
  # Add trinucletiode changes names as row.names
  row.names(cancer_signatures) <-  cancer_signatures$Somatic.Mutation.Type
  # Keep only 96 contributions of the signatures in matrix
  cancer_signatures <-  as.matrix(cancer_signatures[,4:33])
  return(cancer_signatures)
}

# Returns a frame with columns Signature, Sample, Contribution with the absolute
# contribution of each signature for each sample
calc_abs_contribution <- function(nmf.res) {

  total_signatures <-  colSums(nmf.res$signatures) 
  abs_contribution <-  nmf.res$contribution * total_signatures
  rownames(abs_contribution) <- sig.names

  m_contribution <-  melt(abs_contribution)
  colnames(m_contribution) <-  c("Signature", "Sample", "Contribution")

  return (m_contribution)
}

# Returns a frames with the cosine similarity (fit) between the original profile and the
# reconstructed profile
calc_fit_contribution <- function(mut.matrix,nmf.res) {

  # calculate all pairwise cosine similarities
  cos.sim.all <- cos_sim_matrix(mut.matrix, nmf.res$reconstructed)
  # extract cosine similarities per sample between original and reconstructed
  fit.table<- as.data.frame(diag(cos_sim_ori_rec))
  fit.table$Sample <-  rownames(fit.table)
  colnames(fit.table) <-  c("fit","sample")

  return(fit.table)
}

# Returns with the total mutations per case after removing the FFPE component
remove_ffpe_contribution <- function(contribution) {

  norm.contribution <-  contribution %>%
    filter(Signature != "FFPE") %>%
    group_by(Sample) %>%
    summarize(norm.mut=sum(Contribution))

  return (norm.contribution)
}

calc_mmr_contribution <- function(contribution) {

  norm.contribution <-  contribution %>%
    filter(Signature == "MMR") %>%
    group_by(Sample) %>%
    summarize(mmr.mut=sum(Contribution))

  return (norm.contribution)
}

# Summarize cohort according to mmr.status, sex, tissue and tissue of origin
summarize_cohort <-function(mmr.table) {

  summary.age <-  mmr.table %>%
    group_by(mmr.type) %>%
      summarize(n=n(),
            avg.age=mean(Age.DX,na.rm=TRUE),std.age=sd(Age.DX,na.rm=TRUE),
            min.age=min(Age.DX,na.rm=TRUE),max.age=max(Age.DX,na.rm=TRUE))

  summary.sex <- 
    mmr.table %>%
      group_by(mmr.type,Sex) %>%
      summarize (n=n()) %>%
      spread (Sex,n)

  summary.tissue <-  
    mmr.table %>%
    group_by(mmr.type,Tissue) %>%
    summarize (n=n()) %>%
    spread(Tissue,n)

  summary <-  summary.age %>%
    inner_join(summary.sex) %>%
    inner_join(summary.tissue)

  return(summary)
}

```

## Sample filtering

We exclude samples that meet the following criteria:

 * Their sequencing had less than 10 million reads.
 * Samples from our pilot with incomplete information.
 * Have no corresponding normal tissue (only tumor).
 * Samples with unknown tissue type.
 * Samples with wrong MSI/MSS phenotype. 
 * Samples that did not meet age criteria.
 * Samples from ovarian cancer or small bowel cancer.

```{r message=FALSE}
mapping.stats <- read_tsv("supplied_results/master_statistics_table.txt") %>%
  select(sample.name,tumor.status,stat.Total_Reads)

samples.pilot <-  c("MSI_MLH1p_1")
samples.na.tissue <-  c("MSI_MLH1G_2","MSI_MLH1G_6","MSI_MSH2_11","MSS_1")
samples.exclude.age <-  c("MSI_MLH1HM_14")
samples.wrong.phenotype <-  c("MSI_MLH1G_9")
samples.small.bowel <- c("MSI_PMS2_12","MSS_9")

min.reads <- 1e7 #minimum number of reads

mapping.info.filter <-  mapping.stats %>%
  spread(tumor.status,stat.Total_Reads) %>%
  filter(!is.na(normal)) %>%
  filter(normal>min.reads & tumor>min.reads) %>%
  filter(!sample.name %in% samples.pilot) %>%
  filter(!sample.name %in% samples.na.tissue) %>%
  filter(!sample.name %in% samples.exclude.age) %>%
  filter(!sample.name %in% samples.wrong.phenotype) %>%
  filter(!sample.name %in% samples.small.bowel)

```

## Initial NMF analysis

We start the analysis by loading the mutational profiles from the samples that made it into `mapping.info.filter`

```{r warning=FALSE, message=FALSE}
mut.matrix <-  load_mutational_matrix("supplied_results/signatures.txt",
  mapping.info.filter$sample.name)
```

And we apply non-negative matrix factorization

```{r}
nmf.res <- extract_signatures(mut.matrix, rank = 4, nrun = 10)
```

And plot the 4 signatures that we found in the processes side

```{r}
plot_96_profile(nmf.res$signatures, condensed = TRUE)
```

To infer what process corresponds to what signature, we load the COSMIC v2 mutational signatures and compare them against our signatures

```{r message=FALSE}
# Plot COSMIC v2 according to their similarity to each other
cosmic.signatures <- load_cosmic_matrix()
hclust_cosmic <-  cluster_signatures(cosmic.signatures,
                                   method = "average")
cosmic_order <-  colnames(cosmic.signatures)[hclust_cosmic$order]
plot(hclust_cosmic)

# Do a heatmap showing similarity between COSMIC v2 and our signatures
colnames(nmf.res$signatures) <- seq (1:4)
cos_sim_samples_signatures <-  cos_sim_matrix(nmf.res$signatures, cosmic.signatures)
plot_cosine_heatmap(cos_sim_samples_signatures,
                    col_order = cosmic_order,
                    cluster_rows = FALSE)
```

From the figure we get that:

 * Our first signature corresponds to COSMIC signature 10 with a cosine similarity of `r round(cos_sim(cosmic.signatures[,10],nmf.res$signatures[,1]),2)`

 * The second signature corresponds to COSMIC signature 5 with a cosine similarity of `r round(cos_sim(cosmic.signatures[,5],nmf.res$signatures[,2]),2)`

 * The third signature has a high similarity with COSMIC signature 30 with a cosine similarity of `r round(cos_sim(cosmic.signatures[,30],nmf.res$signatures[,3]),2)`. We will come to that later on.

* The fourth signature corresponds to COSMIC signature 6 with a cosine similarity of `r round(cos_sim(cosmic.signatures[,6],nmf.res$signatures[,4]),2)`

On further inspection we compare our third signature against a previously identified FFPE signature

```{r warning=FALSE, message=FALSE}
ffpe.aditya <-  load_old_mutational_matrix("supplied_results/signatures/ffpe.aditya.txt")
```

And we obtain that it has a cosine similarity of `r round(cos_sim(ffpe.aditya,nmf.res$signatures[,3]),2)` against this signature.

Given this information we label each of our resulting signatures:

```{r}
sig.names <-  c("PolE","Aging","FFPE","MMR")
colnames(nmf.res$signatures) <- sig.names
rownames(nmf.res$contribution) <- sig.names

sig.order <-  c("MMR","Aging","PolE","FFPE")
inferred.signatures =nmf.res$signatures[,sig.order]
```

## Plotting samples with high TMB

It seems samples with unusually high TMB have similar mutational profiles. Let's plot the top-6 highest mutated samples in a single graph

```{r message=FALSE, warning=FALSE}
# Get tmb for each sample that was used for mutational analysis
burden.table <-  read_tsv("supplied_results/burden.final.txt") %>%
    mutate(burden = (somatic_snvs/tumor_doc_50X) * 1e6) %>%
    filter(sample %in% colnames(nmf.res$contribution))

# Format sample name to remove MSI prefix
sample.names <-  colnames(nmf.res$contribution) %>%
  str_replace('MSI_','') %>%
  str_replace('MSH6p','MSH6')

# Table with tmb info and short name as a factor sorted by tmb (desc)
burden.table <-  burden.table %>%
    mutate(short.sample = str_replace(sample,"MSI_","")) %>%
    mutate(short.sample = str_replace(short.sample,"MSH6p","MSH6")) %>%
    mutate(short.sample = factor(short.sample,levels=rev(sample.names))) %>%
    arrange(desc(burden))

# Samples sorted by tmb in descending order
pch1 <- burden.table %>%
  ggplot(aes(x=reorder(short.sample,burden),y=burden,fill=type))+
  geom_bar(stat="identity")+
  coord_flip()+
  ylab("Tumor mutational burden")+
  xlab("Sample")

high.burden.samples <-  burden.table %>%
  head(n=6) %>%
  select(sample)

mut.high = load_mutational_matrix("supplied_results/signatures.txt",
  high.burden.samples$sample)

pch2 <- plot_96_profile(mut.high[], condensed = TRUE)

supp.1.grob = grid.arrange(pch1,pch2, ncol=2)

supp.1.grob
#ggsave(file="figures/supp.1.png",
       #supp.1.grob, width=12, height=7, units="in", dpi=300)

```

## Plotting the NMF analysis

```{r}
pch1 <- plot_96_profile(inferred.signatures, condensed = TRUE)

# Get absolute contribution for each signature x sample
sig.abs.contribution = calc_abs_contribution (nmf.res)

# Calculate absolute tmb and pct tmb for each sample x signature 
sample.sig.table = sig.abs.contribution %>%
    inner_join(burden.table, by=c("Sample"="sample")) %>%
    mutate(burden.contribution = Contribution/tumor_doc_50X * 1e6,
           burden.pct = burden.contribution/burden,
           Signature = factor(Signature,levels=sig.order))

# Plot absolute tmb contribution for each sample x signature in waterfall diagram
pch2 <- ggplot(sample.sig.table,
               aes (x=reorder(short.sample,burden),
                    y=burden.contribution,fill=Signature))+
  geom_bar(stat="identity")+
  coord_flip()+
  ylab("Tumor mutational burden")+
  xlab("Sample")+
  theme(legend.position="none")

fig.1.grob = grid.arrange(pch1,pch2, ncol=2)
fig.1.grob
```

### 2021-06-15 Function.Rmd
Here's our work from yesterday on the Bayes function

---
title: "Function"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(MutationalPatterns)
library(BSgenome)
library(BiocManager)
install("BSgenome.Hsapiens.UCSC.hg19")
ref_genome <- "BSgenome.Hsapiens.UCSC.hg19"
library(ref_genome, character.only = TRUE)
```

```{r}
vcf_files1 <- list.files(system.file("extdata", package="MutationalPatterns"), pattern = ".vcf", full.names = TRUE)
vcf_files <- vcf_files1[-c(1,2,3,7)]
sample_names <- c("colon1", "colon2", "colon3", "intestine1", "intestine2", "intestine3", "liver1", "liver2", "liver3")
vcfs <- read_vcfs_as_granges(vcf_files, sample_names, ref_genome)
tissue <- c(rep("colon",3), rep("intestine",3), rep("liver",3))
```

```{r}
mut_mat <- mut_matrix(vcf_list = vcfs, ref_genome = ref_genome)
mut_mat <- mut_mat + 0.0001
```

```{r}
nmf_res <- extract_signatures(mut_mat, rank = 2, nrun = 10)
colnames(nmf_res$signatures) <- c("Signature A", "Signature B")
rownames(nmf_res$contribution) <- c("Signature A", "Signature B")
```

```{r}
#Goal: Pr(Signature|Mutation) in a particular sample

Bayes_function <- function(mutation, process, sample, nmf_matrix, mutation_matrix){
  #Pr(Mutation|Signature)
  a <- nmf_matrix$signatures[mutation,process]/sum(nmf_matrix$signatures[,process])
  #Pr(Signature)
  b <- nmf_matrix$contribution[process,sample]/sum(nmf_matrix$contribution[,sample])
  #Pr(Mutation)
  c <- mutation_matrix[mutation, sample]/sum(mutation_matrix[,sample])
  #Bayes Formula Calculation
  prob <- a*b/c
  return(prob)
}
```

```{r}
#Goal: Pr(SignatureA|A[C>A]G) and Pr(SignatureB|A[C>A]G) for sample colon1
Bayes_function("A[C>A]G", "Signature A", "colon1", nmf_res, mut_mat)
Bayes_function("A[C>A]G", "Signature B", "colon1", nmf_res, mut_mat)
```

### 2021-06-15 Function.Rmd
initial work on function

@@
 #Goal: Pr(Signature|Mutation) in a particular sample
-
-Bayes_function <- function(mutation, process, sample, nmf_matrix, mutation_matrix){
+#Change name for nmf_matrix
+#Split denominator into two pieces, add a parameter for which method we use
+#One function that doesn't need mutational signature
+Bayes_function <- function(mutation, signature, sample, nmf_matrix, mutation_matrix){
   #Pr(Mutation|Signature)
-  a <- nmf_matrix$signatures[mutation,process]/sum(nmf_matrix$signatures[,process])
+  a <- nmf_matrix$signatures[mutation,signature]/sum(nmf_matrix$signatures[,signature])
   #Pr(Signature)
-  b <- nmf_matrix$contribution[process,sample]/sum(nmf_matrix$contribution[,sample])
+  b <- nmf_matrix$contribution[signature,sample]/sum(nmf_matrix$contribution[,sample])
   #Pr(Mutation)
@@
 ```
+
+```{r}
+plot_96_profile(nmf_res$signatures)
+```

### 2021-06-15 bayes_function.R
Implementation of Bayes Theorem into a function

@@ -1,0 +1,52 @@
+library(tidyverse)
+library(MutationalPatterns)
+library(BSgenome)
+library(BiocManager)
+
+extract_all_prob <- function(mutation, sample, nmf_res) {
+  
+  # Creates empty list of length equal to amount of signatures
+  prob = vector(mode = "list", length=length(colnames(nmf_res$signatures)))
+  
+  # initailize accumulator variable and denominator of Bayes Theorem
+  i <- 1
+  denominator = 0
+  
+  # Fills prob with prior and likelihood for each signature
+  for (signature in colnames(nmf_res$signatures)) {
+    # Naming each element of prob
+    names(prob)[i] <- signature
+    
+    # Pr(Mutation|Signature) - likelihood
+    a <- nmf_res$signatures[mutation,signature]/sum(nmf_res$signatures[,signature])
+    
+    # Pr(Signature) - prior
+    b <- nmf_res$contribution[signature,sample]/sum(nmf_res$contribution[,sample])
+    
+    # Appending likelihood and prior into prob
+    prob[[i]] <- c(a, b)
+    
+    # Adding product of likelihood and prior to denominator
+    denominator = denominator + a * b
+    
+    # Accumulating
+    i <- i + 1
+  }
+  
+  # Intialize acummulator and vector to store posteriors
+  i <- 1
+  posteriors <- vector(mode = "double", length = length(colnames(nmf_res$signatures)))
+  
+  # Fills posteriors
+  for (x in colnames(nmf_res$signatures)) {
+    # Calculates numerator of Bayes formula for each signature
+    numerator <- prob[[x]][1] * prob[[x]][2]
+    # Finishes Bayes calculation and inserts value into the vector posteriors
+    posteriors[i] <- numerator / denominator
+    
+    # Accumulating
+    i <- i + 1
+  }
+  
+  return (posteriors)
+}

### 2021-06-16 bayes_function.R
function outputs list with signature names

@@ -1,7 +1,4 @@
-library(tidyverse)
-library(MutationalPatterns)
-library(BSgenome)
-library(BiocManager)
+
 
 extract_all_prob <- function(mutation, sample, nmf_res) {
   
@@ -35,14 +32,15 @@ extract_all_prob <- function(mutation, sample, nmf_res) {
   
   # Intialize acummulator and vector to store posteriors
   i <- 1
-  posteriors <- vector(mode = "double", length = length(colnames(nmf_res$signatures)))
+  posteriors <- vector(mode = "list", length = length(colnames(nmf_res$signatures)))
   
   # Fills posteriors
   for (x in colnames(nmf_res$signatures)) {
+    names(posteriors)[i] <- x
     # Calculates numerator of Bayes formula for each signature
     numerator <- prob[[x]][1] * prob[[x]][2]
     # Finishes Bayes calculation and inserts value into the vector posteriors
-    posteriors[i] <- numerator / denominator
+    posteriors[[i]] <- numerator / denominator
     
     # Accumulating
     i <- i + 1

### 2021-06-16 simulation.R
Started implementing function for sample simulating.

@@ -1,0 +1,35 @@
+library(readr)
+
+# Loads the published COSMIC mutational signatures in a format that can be
+# used by MutationalPatterns
+load_cosmic_matrix<- function(mut.order="supplied_results/mut_sig.order.txt") {
+  
+  sp_url <- paste("https://cancer.sanger.ac.uk/cancergenome/assets/",
+                  "signatures_probabilities.txt", sep = "")
+  
+  cancer_signatures <-  read.table(sp_url, sep = "\t", header = TRUE)
+  # Match the order of the mutation types to MutationalPatterns standard
+  mut_mat <-  read_tsv(mut.order)
+  new_order <-  match(mut_mat$ext.context, cancer_signatures$Somatic.Mutation.Type) 
+  # Reorder cancer signatures dataframe
+  cancer_signatures <-  cancer_signatures[as.vector(new_order),]
+  # Add trinucletiode changes names as row.names
+  row.names(cancer_signatures) <-  cancer_signatures$Somatic.Mutation.Type
+  # Keep only 96 contributions of the signatures in matrix
+  cancer_signatures <-  as.matrix(cancer_signatures[,4:33])
+  return(cancer_signatures)
+}
+
+# Prop = [0.1, 0.9] vector
+# Signatures = [21, 7] vector
+
+create_test_sample <- function(signatures, prop, num) {
+  cosmic_signatures <- load_cosmic_matrix()
+  process <- cosmic_signatures[, signatures]
+  
+  exposures <- matrix(prop, nrow = length(signatures), ncol = num)
+  
+  return ((process %*% exposures) * 10^5)
+}
+
+cosmic_signatures[, "Signature.21"]

### 2021-06-17 Function.Rmd

@@
 vcf_files1 <- list.files(system.file("extdata", package="MutationalPatterns"), pattern = ".vcf", full.names = TRUE)
-vcf_files <- vcf_files1[-c(1,2,3,7)]
-sample_names <- c("colon1", "colon2", "colon3", "intestine1", "intestine2", "intestine3", "liver1", "liver2", "liver3")
+#vcf_files <- vcf_files1[-c(1,2,3,7)]
+vcf_files <-vcf_files1[c(4,5,6)]
+sample_names <- c("colon1", "colon2", "colon3")
 vcfs <- read_vcfs_as_granges(vcf_files, sample_names, ref_genome)
-tissue <- c(rep("colon",3), rep("intestine",3), rep("liver",3))
+tissue <- c(rep("colon",3))
 ```
@@
 ```{r}
+Bayes_function <- function(mutation, signature, sample, nmf_matrix, mutation_matrix){
+  #Pr(Mutation|Signature)
+  prob_mut_given_sig <- nmf_matrix$signatures[mutation,signature]/sum(nmf_matrix$signatures[,signature])
+  #Pr(Signature)
+  prob_sig <- nmf_matrix$contribution[signature,sample]/sum(nmf_matrix$contribution[,sample])
+  #Pr(Mutation)
+
+  #Bayes Formula Calculation
+  prob <- a*b/c
+  return(prob)
+}
+```
+
+```{r}
 #Goal: Pr(SignatureA|A[C>A]G) and Pr(SignatureB|A[C>A]G) for sample colon1
@@
 ```
+
+```{r}
+Bayes_function("G[C>T]G", "Signature A", "colon1", nmf_res, mut_mat)
+Bayes_function("G[C>T]G", "Signature B", "colon1", nmf_res, mut_mat)
+```

### 2021-06-17 Bayes Function Testing.Rmd
testing function on samples created from extreme proportions of two signatures

---
title: "Bayes Function Testing"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MutationalPatterns)
library(BSgenome)
library(BiocManager)
ref_genome <- "BSgenome.Hsapiens.UCSC.hg19"
library(ref_genome, character.only = TRUE)
source("bayes_function.R")
```

```{r}
# Loads the published COSMIC mutational signatures in a format that can be
# used by MutationalPatterns
load_cosmic_matrix<- function(mut.order="supplied_results/mut_sig.order.txt") {

  sp_url <- paste("https://cancer.sanger.ac.uk/cancergenome/assets/",
                "signatures_probabilities.txt", sep = "")

  cancer_signatures <-  read.table(sp_url, sep = "\t", header = TRUE)
  # Match the order of the mutation types to MutationalPatterns standard
  mut_mat <-  read_tsv(mut.order)
  new_order <-  match(mut_mat$ext.context, cancer_signatures$Somatic.Mutation.Type) 
  # Reorder cancer signatures dataframe
  cancer_signatures <-  cancer_signatures[as.vector(new_order),]
  # Add trinucletiode changes names as row.names
  row.names(cancer_signatures) <-  cancer_signatures$Somatic.Mutation.Type
  # Keep only 96 contributions of the signatures in matrix
  cancer_signatures <-  as.matrix(cancer_signatures[,4:33])
  return(cancer_signatures)
}
```

```{r}
create_test_sample <- function(signatures, prop, num) {
  cosmic_signatures <- load_cosmic_matrix()
  process <- cosmic_signatures[, signatures]

  exposures <- matrix(prop, nrow = length(signatures), ncol = num)

  return ((process %*% exposures) * 10^5)
}
```

```{r}
#Create a set of 4 test samples with extreme propotions of COSMIC signatures 18 and 21
test_sample <- create_test_sample(c(18,21), c(1,0,0,1,.9,.1,.1,.9),4)
test_sample
```

```{r}
#Run NMF analysis on signatures
test_nmf <- extract_signatures(test_sample, rank = 2, nrun = 10)

#Plot signatures to match with signatures 18 and 21: V1 appears to match signature 21, V2 appears to match signature 18
plot_96_profile(test_nmf$signatures)

#set column and row names to names of signatures
colnames(test_nmf$signatures) <- c("Signature 21", "Signature 18")
rownames(test_nmf$contribution) <- c("Signature 21", "Signature 18")
```

```{r}
test_nmf$contribution
plot_contribution(test_nmf$contribution, test_nmf$signature, mode = "relative", coord_flip = TRUE)
```

```{r}
#G[C>A]A mutation is the most common in Signature 18
#Sample 1 is ~100% Signature 18
extract_all_prob("G[C>A]A", 1, test_nmf)
#Sample 3 is ~90% Signature 18
extract_all_prob("G[C>A]A", 3, test_nmf)
```

```{r}
#G[T>C]A mutation is the most common in Signature 21
#Sample 2 is ~100% Signature 21
extract_all_prob("G[T>C]A", 2, test_nmf)
#Sample 4 is ~90% Signature 21
extract_all_prob("G[T>C]A", 4, test_nmf)
```

### 2021-06-18 Bayes Function Testing.Rmd
testing bayes function on 4 samples created with extreme proportions of signatures 18 and 21
Updated testing with finer range of exposures and expected output for each chunk.

@@
 ```{r}
+#Creates samples/mutational profiles using selected signatures from COSMIC
+#and desired contributions from each signature.
 create_test_sample <- function(signatures, prop, num) {
@@
 ```{r}
-#Create a set of 4 test samples with extreme propotions of COSMIC signatures 18 and 21
-test_sample <- create_test_sample(c(18,21), c(1,0,0,1,.9,.1,.1,.9),4)
+#Create a set of 7 test samples with extreme proportions of COSMIC signatures 18 and 21
+test_sample <- create_test_sample(c(18,21), c(  1,  0,
+                                              0.9,0.1,
+                                              0.7,0.3,
+                                              0.5,0.5,
+                                              0.3,0.7,
+                                              0.1,0.9,
+                                                0,  1),
+                                  num = 7)
 test_sample
@@
 #set column and row names to names of signatures
-colnames(test_nmf$signatures) <- c("Signature 21", "Signature 18")
-rownames(test_nmf$contribution) <- c("Signature 21", "Signature 18")
+colnames(test_nmf$signatures) <- c("Signature 18", "Signature 21")
+rownames(test_nmf$contribution) <- c("Signature 18", "Signature 21")
 ```
@@
 
+## Proportions should lean in favor for signature 18 and decrease with its sample size
+
 ```{r}
 #G[C>A]A mutation is the most common in Signature 18
+#1287 times more likely in signature 18 than 21
 #Sample 1 is ~100% Signature 18
 extract_all_prob("G[C>A]A", 1, test_nmf)
-#Sample 3 is ~90% Signature 18
+#Sample 2 is ~90% Signature 18
+extract_all_prob("G[C>A]A", 2, test_nmf)
+#Sample 3 is ~70% Signature 18
 extract_all_prob("G[C>A]A", 3, test_nmf)
+#Sample 3 is ~50% Signature 18
+extract_all_prob("G[C>A]A", 3, test_nmf)
+#Sample 4 is ~30% Signature 18
+extract_all_prob("G[C>A]A", 4, test_nmf)
+#Sample 5 is ~10% Signature 18
+extract_all_prob("G[C>A]A", 5, test_nmf)
+#Sample 6 is ~0% Signature 18
+extract_all_prob("G[C>A]A", 6, test_nmf)
+```
+
+## Proportions should lean in favor for signature 21 and decrease with its sample size
+
+```{r}
+#G[T>C]A mutation is the most common in Signature 21
+#269 times more likely in signature 21 than 18
+#Sample 6 is ~100% Signature 21
+extract_all_prob("G[T>C]A", 6, test_nmf)
+#Sample 5 is ~90% Signature 21
+extract_all_prob("G[T>C]A", 5, test_nmf)
+#Sample 4 is ~70% Signature 21
+extract_all_prob("G[T>C]A", 4, test_nmf)
+#Sample 4 is ~50% Signature 21
+extract_all_prob("G[T>C]A", 4, test_nmf)
+#Sample 3 is ~30% Signature 21
+extract_all_prob("G[T>C]A", 3, test_nmf)
+#Sample 2 is ~10% Signature 21
+extract_all_prob("G[T>C]A", 2, test_nmf)
+#Sample 1 is ~0% Signature 21
+extract_all_prob("G[T>C]A", 1, test_nmf)
 ```
@@
 ```{r}
-#G[T>C]A mutation is the most common in Signature 21
-#Sample 2 is ~100% Signature 21
-extract_all_prob("G[T>C]A", 2, test_nmf)
-#Sample 4 is ~90% Signature 21
-extract_all_prob("G[T>C]A", 4, test_nmf)
+#Find smallest difference between contribution rate for every mutation type between
+#signature 18 and 21. Attempting to find mutation type w/ equal probabilities for both.
+cosmic_signatures <- load_cosmic_matrix()
+(process18 <- cosmic_signatures[, c(18)])
+(process21 <- cosmic_signatures[, c(21)])
+diff_signatures <- abs(process21-process18)
+diff_signatures[93]=1 #contribution for signature 21 is 0 (T[T>G]A)
+diff_signatures[85]=1 #contribution for signature 21 is 0
+which.min(diff_signatures)
 ```
+
+## Signature 21 probabilities should be zero for all samples
+
+```{r}
+#LOW probability of T[T>G]A mutation in signature 18
+#zero probability of T[T>G]A mutation in signature 21
+#100% Signature 18 and 0% Signature 21
+extract_all_prob("T[T>G]A", 1, test_nmf)
+#90% Signature 18 and 10% Signature 21
+extract_all_prob("T[T>G]A", 2, test_nmf)
+#70% Signature 18 and 30% Signature 21
+extract_all_prob("T[T>G]A", 3, test_nmf)
+#50% Signature 18 and 50% Signature 21
+extract_all_prob("T[T>G]A", 4, test_nmf)
+#30% Signature 18 and 70% Signature 21
+extract_all_prob("T[T>G]A", 5, test_nmf)
+#10% Signature 18 and 90% Signature 21
+extract_all_prob("T[T>G]A", 6, test_nmf)
+#0% Signature 18 and 100% Signature 21
+extract_all_prob("T[T>G]A", 7, test_nmf) #BAD RESULT
+```
+
+## Probabilities should almost reflect contribution proportions
+
+```{r}
+#Low chances of C[C>G]A mutations in both but roughly equal chances of mutations
+#(4.58 more likely in signature 18, as opposed to hundreds or thousands times more likely)
+#100% Signature 18 and 0% Signature 21
+extract_all_prob("C[C>G]A", 1, test_nmf)
+#90% Signature 18 and 10% Signature 21
+extract_all_prob("C[C>G]A", 2, test_nmf)
+#70% Signature 18 and 30% Signature 21
+extract_all_prob("C[C>G]A", 3, test_nmf)
+#50% Signature 18 and 50% Signature 21
+extract_all_prob("C[C>G]A", 4, test_nmf)
+#30% Signature 18 and 70% Signature 21
+extract_all_prob("C[C>G]A", 5, test_nmf)
+#10% Signature 18 and 90% Signature 21
+extract_all_prob("C[C>G]A", 6, test_nmf)
+#0% Signature 18 and 100% Signature 21
+extract_all_prob("C[C>G]A", 7, test_nmf)
+```

### 2021-06-21 simulation.R
Gives error for giving argument without column/row names

@@ -31,5 +31,3 @@ create_test_sample <- function(signatures, prop, num) {
   
   return ((process %*% exposures) * 10^5)
 }
-
-cosmic_signatures[, "Signature.21"]

### 2021-06-21 simulation.R
updated sample matrix function

@@ -79,5 +79,6 @@ create_signature_sample_matrix <- function(signature.matrix, number.of.mutations
   #creates a matrix of sample frequencies and defines row names as mutation types
   sample.matrix<- as.matrix(sample.df$frequencies)
   rownames(sample.matrix)= mutations
+  #returns matrix
   return(sample.matrix)
 }

### 2021-06-21 Sample Distribution Testing.Rmd
resolved function errors and tested on FFPE and PolE signature distributions

---
title: "Sample Distribution Testing"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(MutationalPatterns)
library(NMF)
source("simulation.R")
```

```{r}
load_old_mutational_matrix<- function(filename,mut.order="supplied_results/mut_sig.order.txt") {

  mutational.signature = read_tsv(filename)
  mut_mat = read_tsv(mut.order)

  mut.join = mutational.signature %>%
    inner_join(mut_mat)

  order.tri = unique(mut.join$ext.context)

  mut.join = mut.join %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    select(ext.context,prob)

  mut.matrix=as.matrix(sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
  rownames(mut.matrix) = mut.join$ext.context

  return (t(mut.matrix)[,1:96])
}

#load ffpe signature
ffpe.signature <-  load_old_mutational_matrix("supplied_results/ffpe.signature.txt")
ffpe.signature
```

```{r}
#convert ffpe.signature to a matrix to be compatible with MutationalPatterns
ffpe.signature.matrix <- as.matrix(ffpe.signature)
plot_96_profile(ffpe.signature.matrix)
```

```{r}
#create ffpe sample vector of 1000 mutations
ffpe.sample.vector <- create_signature_sample_vector(ffpe.signature.matrix, 1000)
ffpe.sample.vector
```

```{r}
#create sample matrix to match ffpe signature
ffpe.sample.matrix <- create_signature_sample_matrix(ffpe.signature.matrix, 10000)
ffpe.sample.matrix
#calculate the cosine similarity between sample matrix and original ffpe matrix
cos_sim_matrix(ffpe.sample.matrix, ffpe.signature.matrix)
#plot original ffpe matrix
plot_96_profile(ffpe.signature.matrix)
#plot our ffpe sample matrix
plot_96_profile(ffpe.sample.matrix)
```

```{r}
#load COSMIC signatures
cosmic.sigs <- load_cosmic_matrix()
```

```{r}
polE <- as.matrix(cosmic.sigs[,10])
polE
plot_96_profile(polE)

polE.sample.vector <- create_signature_sample_vector(polE, 1000)
polE.sample.vector

polE.sample.matrix <- create_signature_sample_matrix(polE, 1000)
polE.sample.matrix
cos_sim_matrix(polE, polE.sample.matrix)
plot_96_profile(polE.sample.matrix)
```

```

### 2021-06-21 simulation.R

@@ -40,7 +40,7 @@ create_signature_sample_vector <- function(signature.matrix, number.of.mutations
   #creates a vector of mutation types
   mutations <- rownames(signature.matrix)
   #creates a vector of probabilities of each mutation type
-  probabilites = as.vector(signature.matrix[,1])
+  probabilities = as.vector(signature.matrix[,1])
   
   #creates sample of desired number of total mutations to match the distribution of mutations in the signature
   signature.sample = sample(mutations, size = number.of.mutations, replace = TRUE, prob = probabilities)
@@ -54,7 +54,7 @@ create_signature_sample_matrix <- function(signature.matrix, number.of.mutations
   #creates a vector of mutation types
   mutations <- rownames(signature.matrix)
   #creates a vector of probabilities of each mutation type
-  probabilites = as.vector(signature.matrix[,1])
+  probabilities = as.vector(signature.matrix[,1])
   #creates sample of desired number of total mutations to match the distribution of mutations in the signature (vector)
   signature.sample = sample(mutations, size = number.of.mutations, replace = TRUE, prob = probabilities)
   

### 2021-06-22 Sample Distribution Testing.Rmd
implemented and tested a function that calculates the cosine similarity between a vector of mutations and mutational signature matrix

@@
 polE.sample.vector
+#function that calculates the cosine similarity between the vector of mutations and a signature matrix
+signature_cosine_similarity(polE.sample.vector, polE)
 
@@
 ```
-
-```

### 2021-06-22 simulation.R

@@ -82,3 +82,34 @@ create_signature_sample_matrix <- function(signature.matrix, number.of.mutations
   #returns matrix
   return(sample.matrix)
 }
+
+#Calculating the cosine similarity between a vector of mutations and a mutational signature
+
+signature_cosine_similarity <- function(mutations.vector, signature.matrix){
+  #create a table of probabilities for the vector of mutations
+  tab <- table(mutations.vector)/sum(table(mutations.vector))
+  #transform table to a data frame and assign column names to mutations and frequencies
+  mutations.df <- data.frame(tab)
+  colnames(mutations.df) <- c("mutations", "frequencies")
+  
+  #defines a function that joins two data frames and replaces NA with 0
+  left_join_NA <- function(x, y, by) {
+    left_join(x = x, y = y, by = by) %>% 
+      mutate_each(funs(replace(., which(is.na(.)), 0)))
+  }
+  
+  #creates a vector of the 96 mutation types
+  mutations <- rownames(signature.matrix)
+  #converts vector of 96 mutation types to a data frame
+  mutation.types.df <- data.frame(mutations)
+  #joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
+  complete.mutations.df <- left_join_NA(mutation.types.df, mutations.df, by = "mutations")
+  
+  #converts data frame of mutation types and frequencies to a data frame and assigns rownames
+  mutations.matrix <- as.matrix(complete.mutations.df$frequencies)
+  rownames(mutations.matrix) = mutations
+  
+  #calculates and returns cosine similarity to the matrix for the signature of interest 
+  cosine.similarity.matrix <- cos_sim_matrix(signature.matrix, mutations.matrix)
+  return(cosine.similarity.matrix[,1])
+}

### 2021-06-23 manual_classification.Rmd
Progressed in classification of signature in simulated mutations

---
title: "multiple_sig_sample"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(MutationalPatterns)
library(NMF)
library(tidyverse)

source("simulation.R")
source("bayes_function.R")
```

```{r}
cosmic.sigs <- load_cosmic_matrix()
ffpe.sig <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))
```

```{r}
cosmic.sig6 <- as.matrix(cosmic.sigs[, 6])
cosmic.sig6

sample.sig6 <- create_signature_sample_vector(cosmic.sig6, 100)
sig6.truth <- rep("MMR", 100)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)
ffpe.truth <- rep("FFPE", 100)
```

```{r}
# Create ground truth column

df.final <- data.frame(append(sample.sig6, sample.ffpe), append(sig6.truth, ffpe.truth))
df.final
```

```{r}
signature.sample <- append(sample.sig6, sample.ffpe)

#converts sample vector into a table of probabilities
tab <- table(signature.sample)/sum(table(signature.sample))
tab
#converts table of probabilities to a data frame and defines column names
sample.probs.df <- data.frame(tab)
colnames(sample.probs.df) <- c("mutations", "frequencies")
sample.probs.df

#defines a function that joins two data frames and replaces NA with 0
left_join_NA <- function(x, y, by) {
  left_join(x = x, y = y, by = by) %>% 
    mutate_each(funs(replace(., which(is.na(.)), 0)))
}

#creates a vector of mutation types
mutations <- rownames(cosmic.sig6)

#creates a data.frame of mutation types to join with sample probabilities
df <- data.frame(mutations)
df

#joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
sample.df <- left_join_NA(df, sample.probs.df, by = "mutations")
sample.df

#creates a matrix of sample frequencies and defines row names as mutation types
sample.matrix<- as.matrix(sample.df$frequencies) + 0.0001
rownames(sample.matrix)= mutations

plot_96_profile(sample.matrix)

nmf_res <- fit_to_signatures(sample.matrix, cbind(cosmic.sig6, ffpe.sig))
nmf_res$signatures <- cbind(cosmic.sig6, ffpe.sig)
nmf_res
```

```{r}
colnames(nmf_res$signatures) <- c("MMR", "FFPE")
rownames(nmf_res$contribution) <- c("MMR", "FFPE")

plot_96_profile(nmf_res$signatures)

mmr.prob <- vector(mode = "double")
ffpe.prob <- vector(mode = "double")

for (x in mutations) {
  probs <- extract_all_prob(x, 1, nmf_res)
  print(x)
  print(probs)
  mmr.prob <- c(mmr.prob, probs$MMR)
  ffpe.prob <- c(ffpe.prob, probs$FFPE)
}

mmr.prob
ffpe.prob
```

### 2021-06-23 simulation.R

@@ -20,6 +20,26 @@ load_cosmic_matrix<- function(mut.order="supplied_results/mut_sig.order.txt") {
   return(cancer_signatures)
 }
 
+load_old_mutational_matrix<- function(filename,mut.order="supplied_results/mut_sig.order.txt") {
+  
+  mutational.signature = read_tsv(filename)
+  mut_mat = read_tsv(mut.order)
+  
+  mut.join = mutational.signature %>%
+    inner_join(mut_mat)
+  
+  order.tri = unique(mut.join$ext.context)
+  
+  mut.join = mut.join %>%
+    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
+    select(ext.context,prob)
+  
+  mut.matrix=as.matrix(sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
+  rownames(mut.matrix) = mut.join$ext.context
+  
+  return (t(mut.matrix)[,1:96])
+}
+
 # Prop = [0.1, 0.9] vector
 # Signatures = [21, 7] vector
 
@@ -31,8 +51,6 @@ create_test_sample <- function(signatures, prop, num) {
   
   return ((process %*% exposures) * 10^5)
 }
-
-cosmic_signatures[, "Signature.21"]
 
 #Creating sample to match signature distribution (vector of all mutations)
 

### 2021-06-23 manual_classification.Rmd
Finished manual implementation of classification

@@
 df.final <- data.frame(append(sample.sig6, sample.ffpe), append(sig6.truth, ffpe.truth))
-df.final
+colnames(df.final) <- c("mutations", "truth")
 ```
@@
   probs <- extract_all_prob(x, 1, nmf_res)
-  print(x)
-  print(probs)
   mmr.prob <- c(mmr.prob, probs$MMR)
@@
 
-mmr.prob
-ffpe.prob
+df2 <- data.frame(mutations, mmr.prob, ffpe.prob)
+df2
+
+df.final <- left_join(df.final, df2, by = "mutations")
+
+df.final <- df.final %>%
+  mutate(
+    classify = ifelse(mmr.prob > ffpe.prob, "MMR", "FFPE"),
+    missClassification = ifelse(truth != classify, 1, 0)
+  )
+
+mean(df.final$missClassification)
 ```

### 2021-06-23 June 25 Presentation Demo.Rmd

---
title: "Friday Presentation Demonstration"
author: "Audrey Mitchell, Marco Ruiz, Soua Yang"
date: "6/23/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(MutationalPatterns)
source("bayes_function.R")
source("simulation.R")
```

#Show Bayes function, explain inputs and outputs

#Testing our function by generating mutational profiles with varying proportions of distinct mutational signatures (Signature 18 and Signature 21) from the COSMIC library
```{r}
test_sample <- create_test_sample(c(18,21), c(  1,  0,
                                              0.9,0.1,
                                              0.7,0.3,
                                              0.5,0.5,
                                              0.3,0.7,
                                              0.1,0.9,
                                                0,  1),
                                  num = 7)
test_sample
```

```{r}
#Run NMF analysis on signatures
test_nmf <- extract_signatures(test_sample, rank = 2, nrun = 10)

#Plot signatures to match with signatures 18 and 21: V1 appears to match signature 21, V2 appears to match signature 18
plot_96_profile(test_nmf$signatures)

#set column and row names to names of signatures
colnames(test_nmf$signatures) <- c("Signature 18", "Signature 21")
rownames(test_nmf$contribution) <- c("Signature 18", "Signature 21")
```

```{r}
test_nmf$contribution
plot_contribution(test_nmf$contribution, test_nmf$signature, mode = "relative", coord_flip = TRUE)
```

## Proportions should lean in favor for signature 21 and decrease with its sample size

```{r}
#G[T>C]A mutation is the most common in Signature 21
#269 times more likely in signature 21 than 18
#Sample 6 is ~100% Signature 21
extract_all_prob("G[T>C]A", 6, test_nmf)
#Sample 5 is ~90% Signature 21
extract_all_prob("G[T>C]A", 5, test_nmf)
#Sample 4 is ~70% Signature 21
extract_all_prob("G[T>C]A", 4, test_nmf)
#Sample 4 is ~50% Signature 21
extract_all_prob("G[T>C]A", 4, test_nmf)
#Sample 3 is ~30% Signature 21
extract_all_prob("G[T>C]A", 3, test_nmf)
#Sample 2 is ~10% Signature 21
extract_all_prob("G[T>C]A", 2, test_nmf)
#Sample 1 is ~0% Signature 21
extract_all_prob("G[T>C]A", 1, test_nmf)
```

## Proportions should lean in favor for signature 18 and decrease with its sample size

```{r}
#G[C>A]A mutation is the most common in Signature 18
#1287 times more likely in signature 18 than 21
#Sample 1 is ~100% Signature 18
extract_all_prob("G[C>A]A", 1, test_nmf)
#Sample 2 is ~90% Signature 18
extract_all_prob("G[C>A]A", 2, test_nmf)
#Sample 3 is ~70% Signature 18
extract_all_prob("G[C>A]A", 3, test_nmf)
#Sample 3 is ~50% Signature 18
extract_all_prob("G[C>A]A", 3, test_nmf)
#Sample 4 is ~30% Signature 18
extract_all_prob("G[C>A]A", 4, test_nmf)
#Sample 5 is ~10% Signature 18
extract_all_prob("G[C>A]A", 5, test_nmf)
#Sample 6 is ~0% Signature 18
extract_all_prob("G[C>A]A", 6, test_nmf)
```

## Probabilities should almost reflect contribution proportions

```{r}
#low levels of C[C>T]G mutations in both signatures
#Sample 3 is 90% Signature 18 and 10% Signature 21
extract_all_prob("T[C>T]G", 3, test_nmf)
#Sample 4 is 90% Signature 21 and 10% Signature 18
extract_all_prob("T[C>T]G", 4, test_nmf)
```

#Example applying our function to real data
```{r}
# Loads a mutational profile with columns sample, ext.context and num
# Subsets such file to samples in sample.list and generates a matrix
# with the mutational profile that can be used by MutationalPatterns
load_mutational_matrix<- function(filename, sample.list) {

  mutational.signature <-  read_tsv(filename)
  order.tri <-  unique(mutational.signature$ext.context)
  sample.names <-  unique(mutational.signature$sample)

  mutational.signature <-  mutational.signature %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    filter(sample %in% sample.list) %>%
    select(sample,ext.context,num) %>%
    spread(sample,num) %>%
    select(one_of("ext.context",sample.names))

  mut.matrix <- as.matrix(sapply(mutational.signature[,2:ncol(mutational.signature)], as.numeric))  
  rownames(mut.matrix) <-  mutational.signature$ext.context

  # Add 1e-04 for use with NMF
  mut.matrix <-  mut.matrix + 1e-04

  return (mut.matrix)
}

# Loads the published COSMIC mutational signatures in a format that can be
# used by MutationalPatterns
load_cosmic_matrix<- function(mut.order="supplied_results/mut_sig.order.txt") {

  sp_url <- paste("https://cancer.sanger.ac.uk/cancergenome/assets/",
                "signatures_probabilities.txt", sep = "")

  cancer_signatures <-  read.table(sp_url, sep = "\t", header = TRUE)
  # Match the order of the mutation types to MutationalPatterns standard
  mut_mat <-  read_tsv(mut.order)
  new_order <-  match(mut_mat$ext.context, cancer_signatures$Somatic.Mutation.Type) 
  # Reorder cancer signatures dataframe
  cancer_signatures <-  cancer_signatures[as.vector(new_order),]
  # Add trinucletiode changes names as row.names
  row.names(cancer_signatures) <-  cancer_signatures$Somatic.Mutation.Type
  # Keep only 96 contributions of the signatures in matrix
  cancer_signatures <-  as.matrix(cancer_signatures[,4:33])
  return(cancer_signatures)
}

mapping.stats <- read_tsv("supplied_results/master_statistics_table.txt") %>%
  select(sample.name,tumor.status,stat.Total_Reads)

samples.pilot <-  c("MSI_MLH1p_1")
samples.na.tissue <-  c("MSI_MLH1G_2","MSI_MLH1G_6","MSI_MSH2_11","MSS_1")
samples.exclude.age <-  c("MSI_MLH1HM_14")
samples.wrong.phenotype <-  c("MSI_MLH1G_9")
samples.small.bowel <- c("MSI_PMS2_12","MSS_9")

min.reads <- 1e7 #minimum number of reads

mapping.info.filter <-  mapping.stats %>%
  spread(tumor.status,stat.Total_Reads) %>%
  filter(!is.na(normal)) %>%
  filter(normal>min.reads & tumor>min.reads) %>%
  filter(!sample.name %in% samples.pilot) %>%
  filter(!sample.name %in% samples.na.tissue) %>%
  filter(!sample.name %in% samples.exclude.age) %>%
  filter(!sample.name %in% samples.wrong.phenotype) %>%
  filter(!sample.name %in% samples.small.bowel) 	
```

```{r}
mut.matrix <-  load_mutational_matrix("supplied_results/signatures.txt",
  mapping.info.filter$sample.name)
```

```{r}
nmf.res <- extract_signatures(mut.matrix, rank = 4, nrun = 10)
sig.names <-  c("PolE","Aging","FFPE","MMR")
colnames(nmf.res$signatures) <- sig.names
rownames(nmf.res$contribution) <- sig.names
```

```{r}
plot_96_profile(nmf.res$signatures)
```

```{r}
extract_all_prob("T[C>A]T", "MSS_6", nmf.res)
```

#Simulating data for which the process behind each mutation is known

```{r}
load_old_mutational_matrix<- function(filename,mut.order="supplied_results/mut_sig.order.txt") {

  mutational.signature = read_tsv(filename)
  mut_mat = read_tsv(mut.order)

  mut.join = mutational.signature %>%
    inner_join(mut_mat)

  order.tri = unique(mut.join$ext.context)

  mut.join = mut.join %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    select(ext.context,prob)

  mut.matrix=as.matrix(sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
  rownames(mut.matrix) = mut.join$ext.context

  return (t(mut.matrix)[,1:96])
}

#load ffpe signature
ffpe.signature <-  load_old_mutational_matrix("supplied_results/ffpe.signature.txt")
ffpe.signature.matrix <- as.matrix(ffpe.signature)
```

```{r}
#create ffpe sample vector of 1000 mutations
ffpe.sample.vector <- create_signature_sample_vector(ffpe.signature.matrix, 1000)
ffpe.sample.vector
signature_cosine_similarity(ffpe.sample.vector, ffpe.signature.matrix)
```

```{r}
#create sample matrix to match ffpe signature
ffpe.sample.matrix <- create_signature_sample_matrix(ffpe.signature.matrix, 10000)
ffpe.sample.matrix
#calculate the cosine similarity between sample matrix and original ffpe matrix
cos_sim_matrix(ffpe.sample.matrix, ffpe.signature.matrix)
#plot original ffpe matrix
plot_96_profile(ffpe.signature.matrix)
#plot our ffpe sample matrix
plot_96_profile(ffpe.sample.matrix)
```

### 2021-06-23 get_classification_df.R
Began implementation of classification function

@@ -1,0 +1,16 @@
+
+
+get_classification_df <- function(samples, sig.names, signatures) {
+  
+  names <- vector(mode = "character")
+  mutation <- vector(mode = "character")
+  
+  for (i in 1:length(samples)) {
+    names <- append(names, rep(sig.names[i], length(samples[[i]])))
+    mutation <- append(mutation, samples[[i]])
+  }
+  
+  df <- data.frame(mutation, names)
+  
+  return (df)
+}

### 2021-06-23 get_classification_df.R
updates to get_classification_df function

@@ -3,14 +3,65 @@
 get_classification_df <- function(samples, sig.names, signatures) {
   
   names <- vector(mode = "character")
-  mutation <- vector(mode = "character")
+  signature.sample <- vector(mode = "character")
   
   for (i in 1:length(samples)) {
     names <- append(names, rep(sig.names[i], length(samples[[i]])))
-    mutation <- append(mutation, samples[[i]])
+    signature.sample <- append(signature.sample, samples[[i]])
   }
   
-  df <- data.frame(mutation, names)
+  df <- data.frame(signature.sample, names)
   
-  return (df)
+  #converts sample vector into a table of probabilities
+  tab <- table(signature.sample)/sum(table(signature.sample))
+
+  #converts table of probabilities to a data frame and defines column names
+  sample.probs.df <- data.frame(tab)
+  colnames(sample.probs.df) <- c("mutations", "frequencies")
+  sample.probs.df
+  
+  #defines a function that joins two data frames and replaces NA with 0
+  left_join_NA <- function(x, y, by) {
+    left_join(x = x, y = y, by = by) %>% 
+      mutate_each(funs(replace(., which(is.na(.)), 0)))
+  }
+  
+  #creates a vector of mutation types
+  mutations <- rownames(signatures[[1]])
+  
+  #creates a data.frame of mutation types to join with sample probabilities
+  df <- data.frame(mutations)
+  df
+  
+  #joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
+  sample.df <- left_join_NA(df, sample.probs.df, by = "mutations")
+  sample.df
+  
+  sample.matrix<- as.matrix(sample.df$frequencies) + 0.0001
+  rownames(sample.matrix)= mutations
+  
+  nmf_res <- fit_to_signatures(sample.matrix, do.call(cbind, signatures))
+  nmf_res$signatures <-do.call(cbind, signatures)
+  
+  colnames(nmf_res$signatures) <- sig.names
+  rownames(nmf_res$contribution) <- sig.names
+  
+  #prob.list <- vector(mode = "list", length = length(samples))
+  
+  prob.list <- list()
+  
+  for (i in length(samples)){
+    prob.list[[i]]<-vector()
+  }
+
+  for (x in mutations) {
+    probs <- extract_all_prob(x, 1, nmf_res)
+    for (i in 1:length(samples)){
+    prob.list[i] <- append(prob.list[i], probs[[i]])
+    print(prob.list[i])
+    }
+  }
+    
+  
+  return (prob.list)
 }

### 2021-06-23 get_classification_df.R
function updates

@@ -10,7 +10,9 @@ get_classification_df <- function(samples, sig.names, signatures) {
     signature.sample <- append(signature.sample, samples[[i]])
   }
   
-  df <- data.frame(signature.sample, names)
+  df.final <- data.frame(signature.sample, names)
+  colnames(df.final) <- c("mutations", "truth")
+  
   
   #converts sample vector into a table of probabilities
   tab <- table(signature.sample)/sum(table(signature.sample))
@@ -46,7 +48,6 @@ get_classification_df <- function(samples, sig.names, signatures) {
   colnames(nmf_res$signatures) <- sig.names
   rownames(nmf_res$contribution) <- sig.names
   
-  #prob.list <- vector(mode = "list", length = length(samples))
   
   prob.list <- list()
   
@@ -54,14 +55,44 @@ get_classification_df <- function(samples, sig.names, signatures) {
     prob.list[[i]]<-vector()
   }
 
-  for (x in mutations) {
-    probs <- extract_all_prob(x, 1, nmf_res)
-    for (i in 1:length(samples)){
-    prob.list[i] <- append(prob.list[i], probs[[i]])
-    print(prob.list[i])
+  for (m in mutations) {
+    probs <- extract_all_prob(m, 1, nmf_res)
+    for (n in 1:length(samples)){
+    prob.list[[n]][m] <- probs[[n]]
     }
   }
-    
   
-  return (prob.list)
+  prob.list.df = data.frame(prob.list)
+  rownames(prob.list.df) <- NULL
+  colnames(prob.list.df) <- sig.names
+  
+  prob.list.df$mutations <- mutations
+  
+  df.final <- left_join(df.final, prob.list.df, by = "mutations")
+  
+  classify <- vector(mode = "character")
+  for (row in df.final){
+    x <- colnames(df.final[which.max(df.final[row,])])
+    append(classify, x)
+  }
+  
+ # df.final <- df.final %>%
+    #group_by(mutations) %>%
+    #mutate(classify = colnames(df.final)[which.max()]) %>%
+    #ungroup()
+  
+  #colnames(df.final)[which.max(df.final[1,])]
+  
+  #classify.df <- df.final %>%
+    #pivot_longer(sig.names, names_to = "process", values_to = "probability") %>%
+    #group_by(mutations) %>%
+    #filter(probability == max(probability)) %>%
+    #mutate(classify = process) %>%
+    #select(mutations, classify)
+  
[…]

### 2021-06-23 get_classification_df.R
completed get_classification_df function (will comment and clean up later) and tested

@@ -71,28 +71,21 @@ get_classification_df <- function(samples, sig.names, signatures) {
   df.final <- left_join(df.final, prob.list.df, by = "mutations")
   
   classify <- vector(mode = "character")
-  for (row in df.final){
-    x <- colnames(df.final[which.max(df.final[row,])])
-    append(classify, x)
+  
+  for (row in 1:nrow(df.final)){
+    probabilities <- df.final[row,3:(2+length(samples))]
+    #print(colnames(probabilities)[which.max(probabilities[1,])])
+    classifier <- colnames(probabilities)[which.max(probabilities[1,])]
+    classify[row]<- classifier
+    #print(df.final[row,3:(2+length(samples))])
   }
   
- # df.final <- df.final %>%
-    #group_by(mutations) %>%
-    #mutate(classify = colnames(df.final)[which.max()]) %>%
-    #ungroup()
+  df.final$classify <- classify
   
-  #colnames(df.final)[which.max(df.final[1,])]
+  df.final <- df.final %>%
+    mutate(misclassification = ifelse(classify != truth, 1, 0))
   
-  #classify.df <- df.final %>%
-    #pivot_longer(sig.names, names_to = "process", values_to = "probability") %>%
-    #group_by(mutations) %>%
-    #filter(probability == max(probability)) %>%
-    #mutate(classify = process) %>%
-    #select(mutations, classify)
-  
-  #df.final <- df.final %>%
-    #left_join(classify.df, by = "mutations")
-  
-  return (classify)
+
+  return (df.final)
 
 }

### 2021-06-25 get_classification_df.R
Commented and cleaned code

@@ -1,15 +1,26 @@
 
+#Arguments:
+#1) list where each element is a sample generated from a certain signature and
+#   each subelement is a trinucleotide mutation type (as a string)
+#2) vector of strings with names of signatures in the same order as the elements
+#   in the list from argument 1
+#3) list where each element is the true mutational signature vector
+
 
 get_classification_df <- function(samples, sig.names, signatures) {
   
+  #Empty vectors for names of mutation sources (signatures) and randomly generated samples
   names <- vector(mode = "character")
   signature.sample <- vector(mode = "character")
   
   for (i in 1:length(samples)) {
+    #Adds each signature name for every element/profile in the sample to empty vector
     names <- append(names, rep(sig.names[i], length(samples[[i]])))
+    #Adds each element/profile in the sample to empty vector
     signature.sample <- append(signature.sample, samples[[i]])
   }
   
+  #Puts the long generated mutations and their source into a data frame and names columns
   df.final <- data.frame(signature.sample, names)
   colnames(df.final) <- c("mutations", "truth")
   
@@ -20,8 +31,7 @@ get_classification_df <- function(samples, sig.names, signatures) {
   #converts table of probabilities to a data frame and defines column names
   sample.probs.df <- data.frame(tab)
   colnames(sample.probs.df) <- c("mutations", "frequencies")
-  sample.probs.df
-  
+
   #defines a function that joins two data frames and replaces NA with 0
   left_join_NA <- function(x, y, by) {
     left_join(x = x, y = y, by = by) %>% 
@@ -33,56 +43,71 @@ get_classification_df <- function(samples, sig.names, signatures) {
   
   #creates a data.frame of mutation types to join with sample probabilities
   df <- data.frame(mutations)
-  df
-  
+
   #joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
   sample.df <- left_join_NA(df, sample.probs.df, by = "mutations")
-  sample.df
-  
+
+  #Add pseudocount to column of probabilities 
   sample.matrix<- as.matrix(sample.df$frequencies) + 0.0001
   rownames(sample.matrix)= mutations
   
+  #Computes exposure matrix using sample matrix and true signature matrix
   nmf_res <- fit_to_signatures(sample.matrix, do.call(cbind, signatures))
+  #Add element with original signatures so that Bayes function accepts this object
   nmf_res$signatures <-do.call(cbind, signatures)
   
[…]

### 2021-06-29 Varying Signature Function Testing.Rmd
plotted average misclassification rate vs. cosine similarity of COSMIC signatures to FFPE, implemented a function to plot an ROC curve for a simulated sample containing mutations from FFPE and one other COSMIC signature

---
title: "Untitled"
author: "Audrey Mitchell"
date: "6/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MutationalPatterns)
library(NMF)
library(tidyverse)
library(ROCR)
source("bayes_function.R")
source("get_classification_df.R")
```

```{r}
#Load FFPE signature
##Function to load FFPE signature
load_old_mutational_matrix<- function(filename,mut.order="supplied_results/mut_sig.order.txt") {

  mutational.signature = read_tsv(filename)
  mut_mat = read_tsv(mut.order)

  mut.join = mutational.signature %>%
    inner_join(mut_mat)

  order.tri = unique(mut.join$ext.context)

  mut.join = mut.join %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    select(ext.context,prob)

  mut.matrix=as.matrix(sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
  rownames(mut.matrix) = mut.join$ext.context

  return (t(mut.matrix)[,1:96])
}

ffpe.signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))
```

```{r}
#create vector of the 96 mutation types
mutations <- rownames(ffpe.signature) 
```

```{r}
#Load COSMIC signatures (version 3)
cosmic.signatures <- get_known_signatures()
```

```{r}
#Create vector of the signature names
signature.names <- colnames(cosmic.signatures)
```

```{r}
#Create list of signature matrices with assigned rownames
signature.matrix.list <- list()
for (i in 1:60){
  sig <- cosmic.signatures[,i]
  sig.mat <- as.matrix(sig)
  rownames(sig.mat) <- mutations
  colnames(sig.mat) <- signature.names[i]
  signature.matrix.list[[i]] <- sig.mat
}
```

```{r}
#outer for loop: loop through each COSMIC v3 signature
average.misclassification.rates <- c()
for (x in 1:60){
  signature.misclass.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 180 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],180)
    #Create FFPE sample vector of 20 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 20)
    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    misclassification.rate <- mean(classification.df$misclassification)
    signature.misclass.vec[y] <- misclassification.rate
  }
  average.signature.misclass <- mean(signature.misclass.vec)
  average.misclassification.rates[x] <- average.signature.misclass
}
```

```{r}
#create vector of cosine similarities to the FFPE signature
ffpe.cosmic.cos.sims <- c()
for (i in 1:60){
  cos.sim <- cos_sim_matrix(signature.matrix.list[[i]], ffpe.signature)
  ffpe.cosmic.cos.sims[i] <- cos.sim[1,1]
}
```

```{r}
cos.sim.misclass.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.misclassification.rates)
colnames(cos.sim.misclass.data) <- c("signature", "cosine_similarity", "average_misclassification")
```

```{r}
ggplot(cos.sim.misclass.data, aes(x = cosine_similarity, y = average_misclassification))+
  geom_point()+
  geom_smooth()+
  labs(title = "Average misclassification rate vs. cosine similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average misclassification rate (10 iterations)")
```

```{r}
model <- lm(average_misclassification ~ log(cosine_similarity), data = cos.sim.misclass.data)
summary(model)
```

#ROC Curves

```{r}
df <- data.frame(signature.names, ffpe.cosmic.cos.sims)
colnames(df) <- c("signature", "cos_similarity") 
df <- df %>%
  filter(signature =="SBS17b" | signature == "SBS12" | signature == "SBS41" | signature == "SBS44" | signature == "SBS30")
df
```

```{r}
#ROC curve function
##function(COSMIC signature, number of COSMIC mutations, number of FFPE mutations, number of iterations)
plot_bayes_roc <- function(cosmic_sig_matrix, cosmic_sig_name, cosmic_mutation_count, ffpe_mutation_count, iterations){
  true.pos.vector <- c()
  false.pos.vector <- c()
  for (i in 1:iterations){
    rownames(cosmic_sig_matrix) <- mutations
    sample <-create_signature_sample_vector(cosmic_sig_matrix, cosmic_mutation_count)
    ffpe.sample <- create_signature_sample_vector(ffpe.signature, ffpe_mutation_count)
    class.df <- get_classification_df(list(sample, ffpe.sample), c(cosmic_sig_name, "FFPE"), list(cosmic_sig_matrix, ffpe.signature))
    true.pos.df <- class.df %>%
    filter(classify == cosmic_sig_name & truth == cosmic_sig_name)
  true.pos <- nrow(true.pos.df)/nrow(class.df)
  true.pos.vector[i] <- true.pos
  false.pos.df <- class.df %>%
    filter(classify == cosmic_sig_name & truth == "FFPE")
  false.pos <- nrow(false.pos.df)/nrow(class.df)
  false.pos.vector[i] <- false.pos
  }
  plot(false.pos.vector, true.pos.vector, xlab = "False Positivity Rate", ylab = "True Positivity Rate")
}

```
#MMR (Signature 6, cos similarity to ffpe = )
```{r}
plot_bayes_roc(cosmic.6, "MMR", 180, 20, 200)
```

##COSMIC signature SBS17b (minimum cosine similarity to FFPE signature: 0.00206)

```{r}
cosmic.17b <- as.matrix(cosmic.signatures[,"SBS17b"])
plot_bayes_roc(cosmic.17b, "SBS17b", 180, 20, 200)
```

##COSMIC signature SBS12 (~25th percentile for cosine similarity to FFPE: 0.05969503)

```{r}
cosmic.12 <- as.matrix(cosmic.signatures[,"SBS12"])
plot_bayes_roc(cosmic.12, "SBS12", 180, 20, 200)
```

##COSMIC signature SBS 41 (~median cosine similarity to FFPE signature: 0.148913)

```{r}
cosmic.41 <- as.matrix(cosmic.signatures[,"SBS41"])
plot_bayes_roc(cosmic.41, "SBS41", 180, 20, 200)
```

##COSMIC signature 44 (~75th percentile cosine similarity to FFPE signature: 0.486769)

```{r}
cosmic.44 <- as.matrix(cosmic.signatures[,"SBS44"])
plot_bayes_roc(cosmic.44, "SBS44", 180, 20, 200)
```

##COSMIC signature 30 (maximum cosine similarity to FFPE signature: 0.905587)

```{r}
cosmic.30 <- as.matrix(cosmic.signatures[,"SBS30"])
plot_bayes_roc(cosmic.30, "SBS30", 180, 20, 200)
```

### 2021-06-29 Varying Signature Function Testing.Rmd
plotted average misclassification rate for Bayes function vs. cosine similarities of COSMIC signatures to FFPE, implemented a function to plot an ROC curve for the Bayes function applied to a simulated sample of mutations from FFPE and one specified COSMIC signature

@@
 ---
-title: "Untitled"
+title: "Testing Success of Bayes Function on Varying COSMIC Signatures"
 author: "Audrey Mitchell"
@@
 ```{r}
+ggplot(data = cos.sim.misclass.data, aes(y = average_misclassification, x = signature))+
+  geom_boxplot()
+```
+
+```{r}
 ggplot(cos.sim.misclass.data, aes(x = cosine_similarity, y = average_misclassification))+
   geom_point()+
-  geom_smooth()+
-  labs(title = "Average misclassification rate vs. cosine similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average misclassification rate (10 iterations)")
+  geom_smooth(method = "lm")+
+  labs(title = "Average misclassification rate vs. cosine similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average misclassification rate (10 iterations)")+
+  scale_x_log10()
+  #label outliars on plot
+  #geom_label(aes(signature))
 ```
@@
 ```{r}
-plot_bayes_roc(cosmic.6, "MMR", 180, 20, 200)
+plot_bayes_roc(cosmic.6, "MMR", 100, 100, 200)
 ```
@@
 cosmic.17b <- as.matrix(cosmic.signatures[,"SBS17b"])
-plot_bayes_roc(cosmic.17b, "SBS17b", 180, 20, 200)
+plot_bayes_roc(cosmic.17b, "SBS17b", 100, 100, 200)
 ```
@@
 cosmic.30 <- as.matrix(cosmic.signatures[,"SBS30"])
-plot_bayes_roc(cosmic.30, "SBS30", 180, 20, 200)
+plot_bayes_roc(cosmic.30, "SBS30", 100, 100, 200)
 ```

### 2021-07-05 ROC_Curves/app.R
shiny app

@@ -1,0 +1,129 @@
+#
+# This is a Shiny web application. You can run the application by clicking
+# the 'Run App' button above.
+#
+# Find out more about building applications with Shiny here:
+#
+#    http://shiny.rstudio.com/
+#
+
+library(shiny)
+library(tidyverse)
+library(MutationalPatterns)
+library(ROCR)
+source("bayes_function.R")
+source("simulation.R")
+source("get_classification_df.R")
+
+# Define UI for application that draws a histogram
+ui <- fluidPage(
+   
+   # Application title
+   titlePanel("Bayesian Classification of COSMIC and FFPE mutations"),
+   
+   # Sidebar with a slider input for number of bins 
+   sidebarLayout(
+      sidebarPanel(
+         sliderInput("total_mutations",
+                     "Total number of mutations:",
+                     min = 100,
+                     max = 1000,
+                     value = 100),
+         sliderInput("proportion_ffpe",
+                     "Proportion of total mutations from the FFPE signature",
+                     min = 0,
+                     max = 1,
+                     value = 0.5),
+         selectInput(inputId = "signature",
+                     label = "COSMIC Signature:",
+                     choices = c ("SBS1", "SBS2", "SBS3","SBS4", "SBS5", "SBS6",
+                                  "SBS7a", "SBS7b", "SBS7c", "SBS7d", "SBS8",
+                                  "SBS9", "SBS10a", "SBS10b", "SBS10c", "SBS10d",
+                                  "SBS11", "SBS12", "SBS13", "SBS14", "SBS15",
+                                  "SBS16", "SBS17a", "SBS17b", "SBS18", "SBS19",
+                                  "SBS20", "SBS21", "SBS22", "SBS23", "SBS24", "SBS25",
+                                  "SBS26", "SBS28", "SBS29", "SBS30", "SBS31",
+                                  "SBS32", "SBS33", "SBS34", "SBS35", "SBS36", "SBS37",
+                                  "SBS38", "SBS39", "SBS40", "SBS41", "SBS42", "SBS44",
+                                  "SBS84", "SBS85", "SBS86", "SBS87", "SBS88", "SBS89",
+                                  "SBS90", "SBS91", "SBS92", "SBS93", "SBS94"),
+                     selected = "SBS1")
+      ),
+      
[…]

### 2021-07-05 ROC_Curves/get_classification_df.R

@@ -1,0 +1,116 @@
+
+#Arguments:
+#1) list where each element is a sample generated from a certain signature and
+#   each subelement is a trinucleotide mutation type (as a string)
+#2) vector of strings with names of signatures in the same order as the elements
+#   in the list from argument 1
+#3) list where each element is the true mutational signature vector
+
+
+get_classification_df <- function(samples, sig.names, signatures) {
+  
+  #Empty vectors for names of mutation sources (signatures) and randomly generated samples
+  names <- vector(mode = "character")
+  signature.sample <- vector(mode = "character")
+  
+  for (i in 1:length(samples)) {
+    #Adds each signature name for every element/profile in the sample to empty vector
+    names <- append(names, rep(sig.names[i], length(samples[[i]])))
+    #Adds each element/profile in the sample to empty vector
+    signature.sample <- append(signature.sample, samples[[i]])
+  }
+  
+  #Puts the long generated mutations and their source into a data frame and names columns
+  df.final <- data.frame(signature.sample, names)
+  colnames(df.final) <- c("mutations", "truth")
+  
+  
+  #converts sample vector into a table of probabilities
+  tab <- table(signature.sample)/sum(table(signature.sample))
+
+  #converts table of probabilities to a data frame and defines column names
+  sample.probs.df <- data.frame(tab)
+  colnames(sample.probs.df) <- c("mutations", "frequencies")
+
+  #defines a function that joins two data frames and replaces NA with 0
+  left_join_NA <- function(x, y, by) {
+    left_join(x = x, y = y, by = by) %>% 
+      mutate_each(funs(replace(., which(is.na(.)), 0)))
+  }
+  
+  #creates a vector of mutation types
+  mutations <- rownames(signatures[[1]])
+  
+  #creates a data.frame of mutation types to join with sample probabilities
+  df <- data.frame(mutations)
+
+  #joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
+  sample.df <- left_join_NA(df, sample.probs.df, by = "mutations")
+
+  #Add pseudocount to column of probabilities 
+  sample.matrix<- as.matrix(sample.df$frequencies) + 0.0001
+  rownames(sample.matrix)= mutations
+  
+  #Computes exposure matrix using sample matrix and true signature matrix
+  nmf_res <- fit_to_signatures(sample.matrix, do.call(cbind, signatures))
+  #Add element with original signatures so that Bayes function accepts this object
[…]

### 2021-07-05 ROC_Curves/simulation.R

@@ -1,0 +1,133 @@
+library(readr)
+
+# Loads the published COSMIC mutational signatures in a format that can be
+# used by MutationalPatterns
+load_cosmic_matrix<- function(mut.order="supplied_results/mut_sig.order.txt") {
+  
+  sp_url <- paste("https://cancer.sanger.ac.uk/cancergenome/assets/",
+                  "signatures_probabilities.txt", sep = "")
+  
+  cancer_signatures <-  read.table(sp_url, sep = "\t", header = TRUE)
+  # Match the order of the mutation types to MutationalPatterns standard
+  mut_mat <-  read_tsv(mut.order)
+  new_order <-  match(mut_mat$ext.context, cancer_signatures$Somatic.Mutation.Type) 
+  # Reorder cancer signatures dataframe
+  cancer_signatures <-  cancer_signatures[as.vector(new_order),]
+  # Add trinucletiode changes names as row.names
+  row.names(cancer_signatures) <-  cancer_signatures$Somatic.Mutation.Type
+  # Keep only 96 contributions of the signatures in matrix
+  cancer_signatures <-  as.matrix(cancer_signatures[,4:33])
+  return(cancer_signatures)
+}
+
+load_old_mutational_matrix<- function(filename,mut.order="supplied_results/mut_sig.order.txt") {
+  
+  mutational.signature = read_tsv(filename)
+  mut_mat = read_tsv(mut.order)
+  
+  mut.join = mutational.signature %>%
+    inner_join(mut_mat)
+  
+  order.tri = unique(mut.join$ext.context)
+  
+  mut.join = mut.join %>%
+    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
+    select(ext.context,prob)
+  
+  mut.matrix=as.matrix(sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
+  rownames(mut.matrix) = mut.join$ext.context
+  
+  return (t(mut.matrix)[,1:96])
+}
+
+# Prop = [0.1, 0.9] vector
+# Signatures = [21, 7] vector
+
+create_test_sample <- function(signatures, prop, num) {
+  cosmic_signatures <- load_cosmic_matrix()
+  process <- cosmic_signatures[, signatures]
+  
+  exposures <- matrix(prop, nrow = length(signatures), ncol = num)
+  
+  return ((process %*% exposures) * 10^5)
+}
+
+#Creating sample to match signature distribution (vector of all mutations)
+
[…]

### 2021-07-06 ROC_Curves/app.R
updates to shiny app

@@ -15,6 +15,15 @@ source("bayes_function.R")
 source("simulation.R")
 source("get_classification_df.R")
 
+ffpe.signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))
+
+#create a vector of the 96 snv mutation types:
+##this is used to assign rownames to COSMIC signature matrices later on
+mutations <- rownames(ffpe.signature)
+
+#load COSMIC version 3 signatures
+cosmic.v3 <- get_known_signatures()
+
 # Define UI for application that draws a histogram
 ui <- fluidPage(
    
@@ -24,30 +33,31 @@ ui <- fluidPage(
    # Sidebar with a slider input for number of bins 
    sidebarLayout(
       sidebarPanel(
-         sliderInput("total_mutations",
-                     "Total number of mutations:",
+         sliderInput(inputId = "total_mutations",
+                     label = "Total number of mutations:",
                      min = 100,
                      max = 1000,
                      value = 100),
-         sliderInput("proportion_ffpe",
-                     "Proportion of total mutations from the FFPE signature",
+         sliderInput(inputId = "proportion_ffpe",
+                     label = "Proportion of total mutations from the FFPE signature",
                      min = 0,
                      max = 1,
                      value = 0.5),
          selectInput(inputId = "signature",
-                     label = "COSMIC Signature:",
-                     choices = c ("SBS1", "SBS2", "SBS3","SBS4", "SBS5", "SBS6",
-                                  "SBS7a", "SBS7b", "SBS7c", "SBS7d", "SBS8",
-                                  "SBS9", "SBS10a", "SBS10b", "SBS10c", "SBS10d",
-                                  "SBS11", "SBS12", "SBS13", "SBS14", "SBS15",
-                                  "SBS16", "SBS17a", "SBS17b", "SBS18", "SBS19",
-                                  "SBS20", "SBS21", "SBS22", "SBS23", "SBS24", "SBS25",
-                                  "SBS26", "SBS28", "SBS29", "SBS30", "SBS31",
-                                  "SBS32", "SBS33", "SBS34", "SBS35", "SBS36", "SBS37",
[…]

### 2021-07-07 plots.Rmd
functioning shiny app, new function to remove intermediate steps in the shiny app

---
title: "Plots"
author: "Audrey Mitchell"
date: "6/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(NMF)
library(tidyverse)
library(ROCR)
source("bayes_function.R")
source("get_classification_df.R")
```

```{r}
#load FFPE signature as a matrix

##Function to load FFPE signature
load_old_mutational_matrix<- function(filename,mut.order="supplied_results/mut_sig.order.txt") {

  mutational.signature = read_tsv(filename)
  mut_mat = read_tsv(mut.order)

  mut.join = mutational.signature %>%
    inner_join(mut_mat)

  order.tri = unique(mut.join$ext.context)

  mut.join = mut.join %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    select(ext.context,prob)

  mut.matrix=as.matrix(sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
  rownames(mut.matrix) = mut.join$ext.context

  return (t(mut.matrix)[,1:96])
}

ffpe.signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))
```

```{r}
#create vector of the 96 mutation types
mutations <- rownames(ffpe.signature) 
```

```{r}
#Load COSMIC signatures (version 3)
cosmic.signatures <- get_known_signatures()

#get MMR signature (SBS6)
cosmic.6 <- as.matrix(cosmic.signatures[,"SBS6"])
rownames(cosmic.6) <- mutations
```

```{r}
mmr.sample <-create_signature_sample_vector(cosmic.6, 50)

ffpe.sample <- create_signature_sample_vector(ffpe.signature, 50)

class.df <- get_classification_df(list(mmr.sample, ffpe.sample), c("MMR", "FFPE"), list(cosmic.6, ffpe.signature))
```

```{r}
class.df2 <- class.df %>%
  mutate(mmr.indicator = ifelse(truth == "MMR", 1, 0))
```

```{r}
predictions <- class.df2$MMR
labels <- class.df2$mmr.indicator
pred <- prediction(predictions, labels)
pred
```

```{r}
perf <- performance(pred, "tpr", "fpr")
perf
```

```{r}
plot(perf,
     avg="threshold")
```

```{r}
auc.perf = performance(pred, measure = "auc")
auc.perf@y.values[[1]][1]
```

#Plotting Reconstructed Profiles
- Create a mutational profile without FFPE -> plot mutational profile
- Add FFPE -> plot mutational profile
- Run get_classification_df() function on the FFPE-added mutational profile
- Filter out all the mutations that our function classifies as FFPE (Pr(FFPE) > .5)
- Calculate cosine similarity between original mutational profile and reconstructed

##Create a mutational profile without FFPE
```{r}
#load COSMIC v3 
cosmic.v3 <- get_known_signatures(muttype = "snv")

#create POLE signature matrix
pole.signature <- as.matrix(cosmic.v3[,"SBS10b"])
rownames(pole.signature) <- mutations

#create POLE sample vector
pole.sample.vector <- create_signature_sample_vector(pole.signature, 100)

#create MMR signature matrix
mmr.signature <- as.matrix(cosmic.v3[,"SBS6"])
rownames(mmr.signature) <- mutations

#create MMR sample vector
mmr.sample.vector <- create_signature_sample_vector(mmr.signature, 100)

#Combine POLE and MMR vector
sample.vector <- c(pole.sample.vector, mmr.sample.vector)

#transform sample.vector into a matrix of probabilities
tab <- table(sample.vector)/sum(table(sample.vector))
##converts table of probabilities to a data frame and defines column names
sample.probs.df <- data.frame(tab)
colnames(sample.probs.df) <- c("mutations", "frequencies")

#defines a function that joins two data frames and replaces NA with 0
left_join_NA <- function(x, y, by) {
  left_join(x = x, y = y, by = by) %>% 
  mutate_each(funs(replace(., which(is.na(.)), 0)))
}

#creates a data.frame of mutation types to join with sample probabilities
df <- data.frame(mutations)

#joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
sample.df <- left_join_NA(df, sample.probs.df, by = "mutations")

#Add pseudocount to column of probabilities 
pole.mmr.matrix<- as.matrix(sample.df$frequencies) + 0.0001
rownames(pole.mmr.matrix)= mutations

plot_96_profile(pole.mmr.matrix)
```

#Add FFPE to the sample
```{r}
ffpe.sample.vector <- create_signature_sample_vector(ffpe.signature, 50)

new.sample.vector <- c(sample.vector, ffpe.sample.vector)

#transform sample.vector into a matrix of probabilities
tab <- table(new.sample.vector)/sum(table(new.sample.vector))
##converts table of probabilities to a data frame and defines column names
new.sample.probs.df <- data.frame(tab)
colnames(new.sample.probs.df) <- c("mutations", "frequencies")
#joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
new.sample.df <- left_join_NA(df, new.sample.probs.df, by = "mutations")

#Add pseudocount to column of probabilities 
pole.mmr.ffpe.matrix<- as.matrix(new.sample.df$frequencies) + 0.0001
rownames(pole.mmr.ffpe.matrix)= mutations

plot_96_profile(pole.mmr.ffpe.matrix)
```

##Run classifier function on the FFPE-added sample, remove FFPE mutations
```{r}
classification.df <- get_classification_df(list(pole.sample.vector, mmr.sample.vector, ffpe.sample.vector), c("POLE", "MMR", "FFPE"), list(pole.signature, mmr.signature, ffpe.signature))

ffpe.removed <- classification.df %>%
  filter(classify != "FFPE")

reconstructed.vector <- ffpe.removed$mutations
```

```{r}
#transform sample.vector into a matrix of probabilities
tab <- table(reconstructed.vector)/sum(table(reconstructed.vector))
##converts table of probabilities to a data frame and defines column names
reconstructed.probs.df <- data.frame(tab)
colnames(reconstructed.probs.df) <- c("mutations", "frequencies")
#joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
reconstructed.df <- left_join_NA(df, reconstructed.probs.df, by = "mutations")

#Add pseudocount to column of probabilities 
reconstructed.matrix<- as.matrix(reconstructed.df$frequencies) + 0.0001
rownames(reconstructed.matrix)= mutations
```

```{r}
plot_96_profile(reconstructed.matrix)
```

```{r}
cos_sim_matrix(pole.mmr.matrix, reconstructed.matrix)
```

### 2021-07-07 ROC_Curves/ROCapp.R

@@ -1,0 +1,79 @@
+#
+# This is a Shiny web application. You can run the application by clicking
+# the 'Run App' button above.
+#
+# Find out more about building applications with Shiny here:
+#
+#    http://shiny.rstudio.com/
+#
+
+library(shiny)
+library(tidyverse)
+library(MutationalPatterns)
+library(ROCR)
+source("bayes_function.R")
+source("simulation.R")
+source("get_classification_df.R")
+source("get_performance_object.R")
+
+ffpe.signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))
+
+#create a vector of the 96 snv mutation types:
+##this is used to assign rownames to COSMIC signature matrices later on
+mutations <- rownames(ffpe.signature)
+
+#load COSMIC version 3 signatures
+cosmic.v3 <- get_known_signatures(muttype = "snv")
+
+cosmic <- colnames(cosmic.v3)
+
+# Define UI for application that draws a histogram
+ui <- fluidPage(
+   
+   # Application title
+   titlePanel("Bayesian Classification of COSMIC and FFPE mutations"),
+   
+   # Sidebar with a slider input for number of bins 
+   sidebarLayout(
+      sidebarPanel(
+         sliderInput(inputId = "total_mutations",
+                     label = "Total number of mutations:",
+                     min = 100,
+                     max = 1000,
+                     value = 100),
+         sliderInput(inputId = "proportion_ffpe",
+                     label = "Proportion of total mutations from the FFPE signature",
+                     min = 0,
+                     max = 1,
+                     value = 0.5),
+         selectInput(inputId = "signature",
+                      label = "COSMIC Signature:",
+                      choices = cosmic,
+                      selected = "SBS1")
+      ),
+      
+      # Show a plot of the generated distribution
+      mainPanel(
+         plotOutput("plot")
+      )
+   )
+)
+
+# Define server logic required to draw a histogram
+server <- function(input,output) {
+  
+  
+  performance.object <- reactive(get_performance_object(input$signature, input$total_mutations,
+                                                        input$proportion_ffpe))
+  
+  
+  output$plot <- renderPlot({
+    p <- plot(performance.object(), avg = "threshold")
+    p
+  })
+
+}
+
+# Run the application 
+shinyApp(ui = ui, server = server)
+

### 2021-07-07 ROC_Curves/app.R

@@ -78,7 +78,7 @@ server <- function(input, output) {
     cosmic_sig_name <- input$signature
   })
 
-  cosmic.signature <- as.matrix(cosmic.v3[,cosmic_sig_name])
+  cosmic.signature <- reactive(as.matrix(cosmic.v3[,cosmic_sig_name()]))
   rownames(cosmic.signature) <- mutations
   
   ffpe.mutation.number <- reactive({
@@ -117,6 +117,7 @@ server <- function(input, output) {
           avg="threshold")
      p
    })
+
 }
 
 # Run the application 

### 2021-07-07 ROC_Curves/get_performance_object.R

@@ -1,0 +1,22 @@
+#ffpe.signature and cosmic.v3 must be loaded ahead of time
+get_performance_object <- function(cosmic_name, total_mut, prop_ffpe){
+  cosmic.sig.mat <- as.matrix(cosmic.v3[,cosmic_name])
+  rownames(cosmic.sig.mat) <- mutations
+  cosmic.mut.no = total_mut-(total_mut*prop_ffpe)
+  ffpe.mut.no = total_mut*prop_ffpe
+  
+  cosmic.samp <- create_signature_sample_vector(cosmic.sig.mat, cosmic.mut.no)
+  ffpe.samp <- create_signature_sample_vector(ffpe.signature, ffpe.mut.no)
+  
+  class.df.1 <- get_classification_df(list(cosmic.samp, ffpe.samp), c(cosmic_name, "FFPE"),
+                                      list(cosmic.sig.mat, ffpe.signature))
+  
+  class.df.2 <- class.df.1 %>%
+    mutate(indicator = ifelse(truth == cosmic_name, 1, 0))
+  
+  pred <- prediction(class.df.2[3], class.df.2[7])
+  perf <- performance(pred, "tpr", "fpr")
+  
+  return(perf)
+}
+

### 2021-07-07 ROC_Curves/ROCapp.R
edited sampling process to set seed for shiny app

@@ -54,7 +54,10 @@ ui <- fluidPage(
       
       # Show a plot of the generated distribution
       mainPanel(
-         plotOutput("plot")
+         tabsetPanel(type = "tabs",
+                     tabPanel("ROC Curve", plotOutput("rocplot"), textOutput("text")),
+                     tabPanel("Reconstructed Profile", plotOutput("reconplot"))
+         )
       )
    )
 )
@@ -67,9 +70,13 @@ server <- function(input,output) {
                                                         input$proportion_ffpe))
   
   
-  output$plot <- renderPlot({
-    p <- plot(performance.object(), avg = "threshold")
+  output$rocplot <- renderPlot({
+    p <- plot(performance.object()[[1]], avg = "threshold")
     p
+  })
+  
+  output$text <- renderText({
+    paste0("Area Under Curve: ", performance.object()[[2]])
   })
 
 }

### 2021-07-07 ROC_Curves/get_performance_object.R

@@ -5,8 +5,8 @@ get_performance_object <- function(cosmic_name, total_mut, prop_ffpe){
   cosmic.mut.no = total_mut-(total_mut*prop_ffpe)
   ffpe.mut.no = total_mut*prop_ffpe
   
-  cosmic.samp <- create_signature_sample_vector(cosmic.sig.mat, cosmic.mut.no)
-  ffpe.samp <- create_signature_sample_vector(ffpe.signature, ffpe.mut.no)
+  cosmic.samp <- create_signature_sample_vector_seed(cosmic.sig.mat, cosmic.mut.no)
+  ffpe.samp <- create_signature_sample_vector_seed(ffpe.signature, ffpe.mut.no)
   
   class.df.1 <- get_classification_df(list(cosmic.samp, ffpe.samp), c(cosmic_name, "FFPE"),
                                       list(cosmic.sig.mat, ffpe.signature))
@@ -16,7 +16,13 @@ get_performance_object <- function(cosmic_name, total_mut, prop_ffpe){
   
   pred <- prediction(class.df.2[3], class.df.2[7])
   perf <- performance(pred, "tpr", "fpr")
+  auc.perf <- performance(pred, measure = "auc")
+  auc.val <- auc.perf@y.values[[1]][1]
   
-  return(perf)
+  return(list(perf, auc.val))
 }
+
+
+
+
 

### 2021-07-07 ROC_Curves/simulation.R

@@ -67,6 +67,21 @@ create_signature_sample_vector <- function(signature.matrix, number.of.mutations
   return(signature.sample)
 }
 
+create_signature_sample_vector_seed <- function(signature.matrix, number.of.mutations){
+  #creates a vector of mutation types
+  mutations <- rownames(signature.matrix)
+  #creates a vector of probabilities of each mutation type
+  probabilities = as.vector(signature.matrix[,1])
+  
+  
+  #creates sample of desired number of total mutations to match the distribution of mutations in the signature
+  set.seed(1)
+  signature.sample = sample(mutations, size = number.of.mutations, replace = TRUE, prob = probabilities)
+  
+  #returns a vector with length = number.of.mutations, each element is a mutation type (string)
+  return(signature.sample)
+}
+
 #Creating sample to match signature distribution (matrix of probabilities)
 create_signature_sample_matrix <- function(signature.matrix, number.of.mutations){
   #creates a vector of mutation types

### 2021-07-07 ROC_Curves/ROCapp.R
Added Profile Reconstruction tab to Shiny App

@@ -11,6 +11,7 @@ library(shiny)
 library(tidyverse)
 library(MutationalPatterns)
 library(ROCR)
+library(gridExtra)
 source("bayes_function.R")
 source("simulation.R")
 source("get_classification_df.R")
@@ -55,8 +56,8 @@ ui <- fluidPage(
       # Show a plot of the generated distribution
       mainPanel(
          tabsetPanel(type = "tabs",
-                     tabPanel("ROC Curve", plotOutput("rocplot"), textOutput("text")),
-                     tabPanel("Reconstructed Profile", plotOutput("reconplot"))
+                     tabPanel("ROC Curve", plotOutput("rocplot"), textOutput("auc")),
+                     tabPanel("Reconstructed Profile", plotOutput("reconplot"), textOutput("cos_sim"))
          )
       )
    )
@@ -69,14 +70,36 @@ server <- function(input,output) {
   performance.object <- reactive(get_performance_object(input$signature, input$total_mutations,
                                                         input$proportion_ffpe))
   
+  profile.matrices <- reactive(get_profile_matrices(input$signature, input$total_mutations,
+                                           input$proportion_ffpe))
   
   output$rocplot <- renderPlot({
     p <- plot(performance.object()[[1]], avg = "threshold")
     p
   })
   
-  output$text <- renderText({
+  output$auc <- renderText({
     paste0("Area Under Curve: ", performance.object()[[2]])
+  })
+  
+  
+  output$reconplot <- renderPlot({
+    p1 <- plot_96_profile(profile.matrices()[[1]])+
+      labs(title = "True COSMIC Signature Profile")
+    #plot_96_profile for COSMIC sample vector
+    p2 <- plot_96_profile(profile.matrices()[[2]])+
+      labs(title = "Mutational Profile of COSMIC Sample")
+    #plot_96_profile for COSMIC and FFPE sample vectors
+    p3 <- plot_96_profile(profile.matrices()[[3]])+
+      labs(title = "Mutational Profile of COSMIC Sample with FFPE Added")
+    #plot_96_profile for reconstructed COSMIC profile
+    p4 <- plot_96_profile(profile.matrices()[[4]])+
+      labs(title = "Reconstructed Profile After Removing FFPE According to Bayesian Classifier")
+    grid.arrange(p1,p2,p3, p4)
+  })
+  
+  output$cos_sim <- renderText({
[…]

### 2021-07-07 ROC_Curves/get_performance_object.R

@@ -22,6 +22,59 @@ get_performance_object <- function(cosmic_name, total_mut, prop_ffpe){
   return(list(perf, auc.val))
 }
 
+vector_to_matrix <- function(mutations.vector){
+  tab <- table(mutations.vector)/sum(table(mutations.vector))
+  #transform table to a data frame and assign column names to mutations and frequencies
+  mutations.df <- data.frame(tab)
+  colnames(mutations.df) <- c("mutations", "frequencies")
+  
+  #defines a function that joins two data frames and replaces NA with 0
+  left_join_NA <- function(x, y, by) {
+    left_join(x = x, y = y, by = by) %>% 
+      mutate_each(funs(replace(., which(is.na(.)), 0)))
+  }
+  
+  #converts vector of 96 mutation types to a data frame
+  mutation.types.df <- data.frame(mutations)
+  #joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
+  complete.mutations.df <- left_join_NA(mutation.types.df, mutations.df, by = "mutations")
+  
+  #converts data frame of mutation types and frequencies to a data frame and assigns rownames
+  mutations.matrix <- as.matrix(complete.mutations.df$frequencies)
+  rownames(mutations.matrix) = mutations
+  
+  return(mutations.matrix)
+}
+
+get_profile_matrices <- function(cosmic_name, total_mut, prop_ffpe) {
+  cosmic.sig.mat <- as.matrix(cosmic.v3[,cosmic_name])
+  rownames(cosmic.sig.mat) <- mutations
+  cosmic.mut.no = total_mut-(total_mut*prop_ffpe)
+  ffpe.mut.no = total_mut*prop_ffpe
+  
+  cosmic.samp <- create_signature_sample_vector_seed(cosmic.sig.mat, cosmic.mut.no)
+  ffpe.samp <- create_signature_sample_vector_seed(ffpe.signature, ffpe.mut.no)
+  
+  cosmic.matrix <- vector_to_matrix(cosmic.samp)
+  colnames(cosmic.matrix) <- c("Original")
+  cosmic.ffpe.matrix <- vector_to_matrix(c(cosmic.samp, ffpe.samp))
+  colnames(cosmic.ffpe.matrix) <- c("FFPE Added")
+  
+  class.df.1 <- get_classification_df(list(cosmic.samp, ffpe.samp), c(cosmic_name, "FFPE"),
[…]

### 2021-07-08 ROC_Curves/ROCapp.R
Updated shiny app with an accuracy chart and points marking our 0.5 threshold/cutoff (code for those in get_performance_object)

@@ -39,7 +39,7 @@ ui <- fluidPage(
       sidebarPanel(
          sliderInput(inputId = "total_mutations",
                      label = "Total number of mutations:",
-                     min = 100,
+                     min = 10,
                      max = 1000,
                      value = 100),
          sliderInput(inputId = "proportion_ffpe",
@@ -57,7 +57,10 @@ ui <- fluidPage(
       mainPanel(
          tabsetPanel(type = "tabs",
                      tabPanel("ROC Curve", plotOutput("rocplot"), textOutput("auc")),
-                     tabPanel("Reconstructed Profile", plotOutput("reconplot"), textOutput("cos_sim"))
+                     tabPanel("Accuracy", plotOutput("accplot"), textOutput("acc50"),
+                              textOutput("accmax"), textOutput("cutoff")),
+                     tabPanel("Reconstructed Profile", plotOutput("reconplot"), 
+                              textOutput("cos_sim"))
          )
       )
    )
@@ -74,12 +77,12 @@ server <- function(input,output) {
                                            input$proportion_ffpe))
   
   output$rocplot <- renderPlot({
-    p <- plot(performance.object()[[1]], avg = "threshold")
-    p
+    plot(performance.object()[["rocplot"]], colorize=TRUE, print.cutoffs.at=0.5, 
+              text.adj=c(-0.2,1.7))
   })
   
   output$auc <- renderText({
-    paste0("Area Under Curve: ", performance.object()[[2]])
+    paste0("Area Under Curve: ", performance.object()[["auc"]])
   })
   
   
@@ -99,7 +102,27 @@ server <- function(input,output) {
   })
   
   output$cos_sim <- renderText({
-    paste0("Cosine Similarity Between Original COSMIC Sample Profile and Reconstructed Profile: ", profile.matrices()[[5]])
+    paste0("Cosine Similarity Between Original COSMIC Sample Profile and Reconstructed Profile: ", 
+           profile.matrices()[[5]])
+  })
+  
+  output$accplot <- renderPlot({
+    plot(performance.object()[["accplot"]])
+    points(0.5, performance.object()[["acc50"]])
+    text(0.5, performance.object()[["acc50"]]- 0.02, "0.5")
+  })
+  
+  output$acc50 <- renderText({
+    paste0("Accuracy at 50% threshold/cutoff: ", performance.object()[["acc50"]])
+  })
+  
+  output$accmax <- renderText({
[…]

### 2021-07-08 ROC_Curves/get_performance_object.R

@@ -18,8 +18,24 @@ get_performance_object <- function(cosmic_name, total_mut, prop_ffpe){
   perf <- performance(pred, "tpr", "fpr")
   auc.perf <- performance(pred, measure = "auc")
   auc.val <- auc.perf@y.values[[1]][1]
+  acc.perf <- performance(pred, measure = "acc")
+  #find accuracy for 0.5 threshold/cutoff
+  #get line segment using threshold immediately before and after 0.5
+  low.ind = max(which(slot(acc.perf, "x.values")[[1]] >=0.5))
+  high.ind = min(which(slot(acc.perf, "x.values")[[1]] <=0.5))
+  #point slop formula of a line
+  slope = (slot(acc.perf, "y.values")[[1]][high.ind] - 
+             slot(acc.perf, "y.values")[[1]][low.ind])/
+    (slot(acc.perf, "x.values")[[1]][high.ind] - slot(acc.perf, "x.values")[[1]][low.ind])
+  acc.50 = slope*(0.5 - slot(acc.perf, "x.values")[[1]][low.ind]) + 
+    slot(acc.perf, "y.values")[[1]][low.ind]
+  #find the cutoff that makes the max accuracy
+  ind = which.max( slot(acc.perf, "y.values")[[1]] )
+  acc.max = slot(acc.perf, "y.values")[[1]][ind]
+  cutoff = slot(acc.perf, "x.values")[[1]][ind]
   
-  return(list(perf, auc.val))
+  return(list(rocplot = perf, auc = auc.val, accplot = acc.perf, 
+              acc50 = acc.50, acc.max = acc.max, cutoff = cutoff))
 }
 
 vector_to_matrix <- function(mutations.vector){

### 2021-07-12 plots.Rmd
Added point labels to average misclassification rate vs. cosine similarity to FFPE scatterplot, started working on AUC vs cosine similarity to FFPE scatterplot

@@
 library(tidyverse)
+library(ggrepel)
 library(ROCR)
@@
 source("get_classification_df.R")
+source("simulation.R")
 ```
@@
 #load FFPE signature as a matrix
-
-##Function to load FFPE signature
-load_old_mutational_matrix<- function(filename,mut.order="supplied_results/mut_sig.order.txt") {
-
-  mutational.signature = read_tsv(filename)
-  mut_mat = read_tsv(mut.order)
-
-  mut.join = mutational.signature %>%
-    inner_join(mut_mat)
-
-  order.tri = unique(mut.join$ext.context)
-
-  mut.join = mut.join %>%
-    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
-    select(ext.context,prob)
-
-  mut.matrix=as.matrix(sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
-  rownames(mut.matrix) = mut.join$ext.context
-
-  return (t(mut.matrix)[,1:96])
-}
-
 ffpe.signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))
@@
 #Load COSMIC signatures (version 3)
-cosmic.signatures <- get_known_signatures()
+cosmic.signatures <- get_known_signatures(muttype = "snv")
 
@@
 ```
+
+#ROC Curve for Sample of 50 MMR mutations and 50 FFPE mutations
 
@@
 ```
+
+#Scatterplot of misclassification rate vs. cosine similarity of COSMIC signature to FFPE signature
+
+```{r}
+#Create vector of the signature names
+signature.names <- colnames(cosmic.signatures)
+```
+
+```{r}
+#Create list of signature matrices with assigned rownames
+signature.matrix.list <- list()
+for (i in 1:60){
+  sig <- cosmic.signatures[,i]
+  sig.mat <- as.matrix(sig)
+  rownames(sig.mat) <- mutations
+  colnames(sig.mat) <- signature.names[i]
+  signature.matrix.list[[i]] <- sig.mat
+}
+```
+
+```{r}
+#outer for loop: loop through each COSMIC v3 signature
+average.misclassification.rates <- c()
+average.auc.values <- c()
+for (x in 1:60){
+  signature.misclass.vec <- c()
+  signature.auc.vec <- c()
+  #inner for loop: calculate misclassification rate 10x for each signature
+  for (y in 1:10){
+    #create a sample vector of 180 mutations for the COSMIC signature corresponding to number x
+    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],180)
+    #Create FFPE sample vector of 20 mutations
+    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 20)
+    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
+    misclassification.rate <- mean(classification.df$misclassification)
+    signature.misclass.vec[y] <- misclassification.rate
+    class.df.colnames <- colnames(classification.df)
+    sig.name <- class.df.colnames[3]
+    classification.df.2 <- classification.df %>%
+      mutate(signature.indicator = ifelse(classify == sig.name, 1, 0))
+    pred <- prediction(classification.df.2[3], classification.df.2$signature.indicator)
+    auc.perf <- performance(pred, measure = "auc")
+    auc.value <- auc.perf@y.values[[1]][1]
+    signature.auc.vec[y] <- auc.value
+  }
+  average.signature.misclass <- mean(signature.misclass.vec)
+  average.misclassification.rates[x] <- average.signature.misclass
+  average.signature.auc <- mean(signature.auc.vec)
+  average.auc.values[x] <- average.signature.auc
+}
+```
+
+```{r}
+#Example of loop for SBS5: all mutations are classified as SBS5 -> problem with ROCR
+sample.vec <- create_signature_sample_vector(signature.matrix.list[[5]],180)
+    #Create FFPE sample vector of 20 mutations
+ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 20)
+classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[5],"FFPE"), list(signature.matrix.list[[5]], ffpe.signature))
+misclassification.rate <- mean(classification.df$misclassification)
+signature.misclass.vec[10] <- misclassification.rate
+class.df.colnames <- colnames(classification.df)
+sig.name <- class.df.colnames[3]
+classification.df.2 <- classification.df %>%
+  mutate(signature.indicator = ifelse(classify == sig.name, 1, 0))
+pred <- prediction(classification.df.2[3], classification.df.2$signature.indicator)
+    auc.perf <- performance(pred, measure = "auc")
+    auc.value <- auc.perf@y.values[[1]][1]
+    signature.auc.vec[10] <- auc.value
+```
+
+```{r}
+#create vector of cosine similarities to the FFPE signature
+ffpe.cosmic.cos.sims <- c()
+for (i in 1:60){
+  cos.sim <- cos_sim_matrix(signature.matrix.list[[i]], ffpe.signature)
+  ffpe.cosmic.cos.sims[i] <- cos.sim[1,1]
+}
+```
+
+```{r}
+cos.sim.misclass.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.misclassification.rates)
+colnames(cos.sim.misclass.data) <- c("signature", "cosine_similarity", "average_misclassification")
+```
+
+```{r}
+ggplot(cos.sim.misclass.data, aes(x = cosine_similarity, y = average_misclassification))+
+  geom_point()+
+  geom_smooth(method = "lm")+
+  labs(title = "Average misclassification rate vs. cosine similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average misclassification rate (10 iterations)")+
+  scale_x_log10()+
+  geom_label_repel(aes(label = signature), max.overlaps = 7, label.size =   0.1)
+```

### 2021-07-12 firevat/firevat_testing.Rmd

---
title: "firevat_testing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(FIREVAT)
library(bedr)
library(stringr)
library(MutationalPatterns)

source("../simulation.R")
```

```{r}
cosmic.sigs <- get_known_signatures()
cosmic.sig6 <- as.matrix(cosmic.sigs[,6])
ffpe.sig <- as.matrix(load_old_mutational_matrix("../supplied_results/ffpe.signature.txt")) # Doesn't work

df.sigs <- data.frame(GetPCAWGMutSigs()[1:3], cosmic.sigs, ffpe.sig)
rownames(df.sigs) <- seq(1, 96)
```

```{r}
# Output directory
output.dir <- "./" # assign this path

# VCF file
sample.vcf.file <- system.file("extdata", "DCC_PCAWG_Cell_Lines_HCC1954.vcf", package = "FIREVAT")
sample.vcf.file <- "../sample3.vcf"
sample.vcf.file <- "./short-example.vcf"
sample.vcf.file <- "../sample-combined.vcf"
sample.vcf.file <- system.file("extdata", "intestine1-sample.vcf", package = "MutationalPatterns")

# Configuration file
config.file <- system.file("config", "PCAWG_DKFZ_Cell_Line_Filtering_Params.json", package = "FIREVAT")
config.file <- "../custom_config.json"

colnames(df.sigs)[4:length(colnames(df.sigs))]
```

```{r}
run <- function(signature = "SBS6", default = 0) {

  if (default == 0) {
    # Run FIREVAT
    results <- RunFIREVAT(vcf.file = sample.vcf.file,
                          vcf.file.genome = 'hg38', # for mouse variants: 'mm10'
                          config.file = config.file,
                          df.ref.mut.sigs = df.sigs,
                          target.mut.sigs = c(signature, "ffpe.sig"),
                          sequencing.artifact.mut.sigs = c("ffpe.sig"),
                          output.dir = output.dir,
                          objective.fn = Default.Obj.Fn,
                          num.cores = 2,
                          ga.pop.size = 100,
                          ga.max.iter = 5,
                          ga.run = 5,
                          perform.strand.bias.analysis = FALSE,
                          annotate = FALSE)
  } else if (default == 1) {
    # Run FIREVAT
    results <- RunFIREVAT(vcf.file = sample.vcf.file,
                          vcf.file.genome = 'hg19', # for mouse variants: 'mm10'
                          config.file = config.file,
                          df.ref.mut.sigs = GetPCAWGMutSigs(),
                          target.mut.sigs = GetPCAWGMutSigsNames(),
                          sequencing.artifact.mut.sigs = PCAWG.All.Sequencing.Artifact.Signatures,
                          output.dir = output.dir,
                          objective.fn = Default.Obj.Fn,
                          num.cores = 2,
                          ga.pop.size = 100,
                          ga.max.iter = 5,
                          ga.run = 5,
                          perform.strand.bias.analysis = FALSE,
                          annotate = FALSE)
  }
}
```

```{r}
run("SBS4")
```

```{r}
GetPCAWGMutSigs()
GetPCAWGMutSigsNames()
PCAWG.All.Sequencing.Artifact.Signatures
```

```{r}
vcf.obj <- ReadVCF(sample.vcf.file)

mutalisk.results <- RunMutalisk(vcf.obj = vcf.obj,
                                df.ref.mut.sigs = df.sigs,
                                target.mut.sigs = c("SBS6", "ffpe.sig"),
                                n.sample = 2,
                                n.iter = 10)

g <- PlotMutaliskResults(
    mutalisk.results = mutalisk.results,
    signatures = unique(mutalisk.results$identified.mut.sigs),
    df.ref.sigs.groups.colors = GetPCAWGMutSigsEtiologiesColors(),
    trinuc.max.y = 1,
    trinuc.min.y = 0,
    mut.type.max.y = 1,
    title = "")

print(g$f1)

```

### 2021-07-12 ROC_Curves/get_classification_df.R
Fixed error.

@@ -97,7 +97,12 @@ get_classification_df <- function(samples, sig.names, signatures) {
     #store posterior probabilities in each row for all signatures
     probabilities <- df.final[row,3:(2+length(samples))]
     #finds the index for the maximum probability and retrieves name of signature that is the max
-    classifier <- colnames(probabilities)[which.max(probabilities[1,])]
+    if(max(is.na(probabilities)) == 0){
+      classifier <- colnames(probabilities)[which.max(probabilities[1,])]
+    }
+    else{
+      classifier <- NA
+    }    
     #each element is signature name (string) that has the maximum posterior probability
     classify[row]<- classifier
   }

### 2021-07-15 firevat/firevat_testing.Rmd
More firevat testing and firevat classification table contains label for both signatures

@@
 library(FIREVAT)
+library(tidyverse)
 library(bedr)
@@
 library(MutationalPatterns)
+library(BSgenome)
+ref_genome <- "BSgenome.Hsapiens.UCSC.hg19"
+library(ref_genome, character.only = TRUE)
+library(ROCR)
 
 source("../simulation.R")
-```
+source("../sample_conversions.R")
 
-```{r}
-cosmic.sigs <- get_known_signatures()
-cosmic.sig6 <- as.matrix(cosmic.sigs[,6])
-ffpe.sig <- as.matrix(load_old_mutational_matrix("../supplied_results/ffpe.signature.txt")) # Doesn't work
-
-df.sigs <- data.frame(GetPCAWGMutSigs()[1:3], cosmic.sigs, ffpe.sig)
-rownames(df.sigs) <- seq(1, 96)
-```
-
-```{r}
-# Output directory
-output.dir <- "./" # assign this path
-
-# VCF file
-sample.vcf.file <- system.file("extdata", "DCC_PCAWG_Cell_Lines_HCC1954.vcf", package = "FIREVAT")
-sample.vcf.file <- "../sample3.vcf"
-sample.vcf.file <- "./short-example.vcf"
-sample.vcf.file <- "../sample-combined.vcf"
-sample.vcf.file <- system.file("extdata", "intestine1-sample.vcf", package = "MutationalPatterns")
-
-# Configuration file
-config.file <- system.file("config", "PCAWG_DKFZ_Cell_Line_Filtering_Params.json", package = "FIREVAT")
-config.file <- "../custom_config.json"
-
-colnames(df.sigs)[4:length(colnames(df.sigs))]
+seq <- getSeq(Hsapiens, "chr1")
 ```
@@
 }
+
+get_roc_point <- function(df) {
+  tp <- df %>%
+    filter(truth == classify, classify != "FFPE") %>%
+    count()
+  tp <- tp$n
+
+  tn <- df %>%
+    filter(truth == classify, classify == "FFPE") %>%
+    count()
+  tn <- tn$n
+
+  fp <- df %>%
+    filter(truth != classify, classify != "FFPE") %>%
+    count()
+  fp <- fp$n
+
+  fn <- df %>%
+    filter(truth != classify, classify == "FFPE") %>%
+    count()
+  fn <- fn$n
+
+  tpr <- tp / (tp + fn)
+  fpr <- fp / (fp + tn)
+
+  return(c(fpr, tpr))
+}
 ```
@@
 ```{r}
+cosmic.sigs <- get_known_signatures()
+cosmic.sig6 <- as.matrix(cosmic.sigs[,6])
+cosmic.sig4 <- as.matrix(cosmic.sigs[,4])
+ffpe.sig <- as.matrix(load_old_mutational_matrix("../ROC_Curves/supplied_results/ffpe.signature.txt", "../supplied_results/mut_sig.order.txt")) # Doesn't work
+
+df.sigs <- data.frame(GetPCAWGMutSigs()[1:3], cosmic.sigs, ffpe.sig)
+rownames(df.sigs) <- seq(1, 96)
+```
+
+```{r}
+sample.sig6 <- create_signature_sample_vector(cosmic.sig6, 100)
+sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)
+
+sample1.df <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR", "FFPE"), list(cosmic.sig6, ffpe.sig))
+
+write_sample_to_vcf(sample1.df, "sample1.vcf", Hsapiens, "chr1")
+
+sample.vcf.file <- "sample1.vcf"
+config.file <- "../custom_config.json"
+output.dir <- "./"
+
+run("SBS6")
+```
+
+```{r}
+sample1.firevat.df <- create_classify_df("sample1.vcf", seq)
+
+mean(sample1.firevat.df$misclassification)
+mean(sample1.df$misclassification)
+
+roc.point <- get_roc_point(sample1.firevat.df)
+get_roc_point(sample1.df)
+
+sample1.df <- sample1.df %>%
+  mutate(binaryTruth = ifelse(truth == "MMR", 1, 0))
+
+sample1.prediction <- prediction(sample1.df$MMR, sample1.df$binaryTruth)
+sample1.perf <- performance(sample1.prediction, "tpr", "fpr")
+plot(sample1.perf)
+points(x=roc.point[1], y=roc.point[2])
+```
+
+```{r}
+sample.sig6 <- create_signature_sample_vector(cosmic.sig6, 500)
+sample.ffpe <- create_signature_sample_vector(ffpe.sig, 500)
+
+sample2.df <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR", "FFPE"), list(cosmic.sig6, ffpe.sig))
+
+write_sample_to_vcf(sample2.df, "sample2.vcf", Hsapiens, "chr1")
+
+sample.vcf.file <- "sample2.vcf"
+config.file <- "../custom_config.json"
+output.dir <- "./"
+
+run("SBS6")
+```
+
+```{r}
+sample2.firevat.df <- create_classify_df("sample2.vcf", seq)
+
+mean(sample2.firevat.df$misclassification)
+mean(sample2.df$misclassification)
+
+roc.point <- get_roc_point(sample2.firevat.df)
+get_roc_point(sample2.df)
+
+sample2.df <- sample2.df %>%
+  mutate(binaryTruth = ifelse(truth == "MMR", 1, 0))
+
+sample2.prediction <- prediction(sample2.df$MMR, sample2.df$binaryTruth)
+sample2.perf <- performance(sample2.prediction, "tpr", "fpr")
+plot(sample2.perf)
+points(x=roc.point[1], y=roc.point[2])
+```
+
+```{r}
+sample.sig4 <- create_signature_sample_vector(cosmic.sig4, 100)
+sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)
+
+sample3.df <- get_classification_df(list(sample.sig4, sample.ffpe), c("MMR", "FFPE"), list(cosmic.sig4, ffpe.sig))
+
+write_sample_to_vcf(sample3.df, "sample3.vcf", Hsapiens, "chr1")
+
+sample.vcf.file <- "sample3.vcf"
+config.file <- "../custom_config.json"
+output.dir <- "./"
+
 run("SBS4")
@@
 ```{r}
-GetPCAWGMutSigs()
-GetPCAWGMutSigsNames()
-PCAWG.All.Sequencing.Artifact.Signatures
+sample3.firevat.df <- create_classify_df("sample3.vcf", seq)
+
+mean(sample3.firevat.df$misclassification)
+mean(sample3.df$misclassification)
+
+roc.point <- get_roc_point(sample3.firevat.df)
+get_roc_point(sample3.df)
+
+sample3.df <- sample3.df %>%
+  mutate(binaryTruth = ifelse(truth == "MMR", 1, 0))
+
+sample3.prediction <- prediction(sample3.df$MMR, sample3.df$binaryTruth)
+sample3.perf <- performance(sample3.prediction, "tpr", "fpr")
+plot(sample3.perf)
+points(x=roc.point[1], y=roc.point[2])
 ```
@@
 ```{r}
-vcf.obj <- ReadVCF(sample.vcf.file)
+sample.sig4 <- create_signature_sample_vector(cosmic.sig4, 500)
+sample.ffpe <- create_signature_sample_vector(ffpe.sig, 500)
 
-mutalisk.results <- RunMutalisk(vcf.obj = vcf.obj,
-                                df.ref.mut.sigs = df.sigs,
-                                target.mut.sigs = c("SBS6", "ffpe.sig"),
-                                n.sample = 2,
-                                n.iter = 10)
+sample4.df <- get_classification_df(list(sample.sig4, sample.ffpe), c("MMR", "FFPE"), list(cosmic.sig4, ffpe.sig))
 
-g <- PlotMutaliskResults(
-    mutalisk.results = mutalisk.results,
-    signatures = unique(mutalisk.results$identified.mut.sigs),
-    df.ref.sigs.groups.colors = GetPCAWGMutSigsEtiologiesColors(),
-    trinuc.max.y = 1,
-    trinuc.min.y = 0,
-    mut.type.max.y = 1,
-    title = "")
+write_sample_to_vcf(sample4.df, "sample4.vcf", Hsapiens, "chr1")
 
-print(g$f1)
+sample.vcf.file <- "sample4.vcf"
+config.file <- "../custom_config.json"
+output.dir <- "./"
 
+run("SBS4")
 ```
+
+```{r}
+sample4.firevat.df <- create_classify_df("sample4.vcf", seq)
+
+mean(sample4.firevat.df$misclassification)
+mean(sample4.df$misclassification)
+
+roc.point <- get_roc_point(sample4.firevat.df)
+get_roc_point(sample4.df)
+
+sample4.df <- sample4.df %>%
+  mutate(binaryTruth = ifelse(truth == "MMR", 1, 0))
+
+sample4.prediction <- prediction(sample4.df$MMR, sample4.df$binaryTruth)
+sample4.perf <- performance(sample4.prediction, "tpr", "fpr")
+plot(sample4.perf)
+points(x=roc.point[1], y=roc.point[2])
+```

### 2021-07-21 ROC_Curves/get_classification_df.R
TryCatch function to account for random error (classify[row] <- classifier: replacement is length zero).

@@ -27,11 +27,11 @@ get_classification_df <- function(samples, sig.names, signatures) {
   
   #converts sample vector into a table of probabilities
   tab <- table(signature.sample)/sum(table(signature.sample))
-
+  
   #converts table of probabilities to a data frame and defines column names
   sample.probs.df <- data.frame(tab)
   colnames(sample.probs.df) <- c("mutations", "frequencies")
-
+  
   #defines a function that joins two data frames and replaces NA with 0
   left_join_NA <- function(x, y, by) {
     left_join(x = x, y = y, by = by) %>% 
@@ -43,10 +43,10 @@ get_classification_df <- function(samples, sig.names, signatures) {
   
   #creates a data.frame of mutation types to join with sample probabilities
   df <- data.frame(mutations)
-
+  
   #joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
   sample.df <- left_join_NA(df, sample.probs.df, by = "mutations")
-
+  
   #Add pseudocount to column of probabilities 
   sample.matrix<- as.matrix(sample.df$frequencies) + 0.0001
   rownames(sample.matrix)= mutations
@@ -68,7 +68,7 @@ get_classification_df <- function(samples, sig.names, signatures) {
     #each element of empty list is vector of probabilities for every element/profile from mutational sample
     prob.list[[i]]<-vector()
   }
-
+  
   for (m in mutations) {
     #each mutation type (string) is sent to Bayes function along with "nmf" results
     probs <- extract_all_prob(m, 1, nmf_res)
@@ -96,15 +96,18 @@ get_classification_df <- function(samples, sig.names, signatures) {
   for (row in 1:nrow(df.final)){
     #store posterior probabilities in each row for all signatures
     probabilities <- df.final[row,3:(2+length(samples))]
-    #finds the index for the maximum probability and retrieves name of signature that is the max
-    if(max(is.na(probabilities)) == 0){
-      classifier <- colnames(probabilities)[which.max(probabilities[1,])]
-    }
-    else{
-      classifier <- NA
-    }    
-    #each element is signature name (string) that has the maximum posterior probability
-    classify[row]<- classifier
+    tryCatch(
+      expr = {
+        #finds the index for the maximum probability and retrieves name of signature that is the max
+        classifier <- colnames(probabilities)[which.max(probabilities[1,])]
+        #each element is signature name (string) that has the maximum posterior probability
+        classify[row]<- classifier
+      },
+      error = function(e){
[…]

### 2021-07-21 get_classification_df.R

@@ -96,15 +96,18 @@ get_classification_df <- function(samples, sig.names, signatures) {
   for (row in 1:nrow(df.final)){
     #store posterior probabilities in each row for all signatures
     probabilities <- df.final[row,3:(2+length(samples))]
-    #finds the index for the maximum probability and retrieves name of signature that is the max
-    if(max(is.na(probabilities)) == 0){
-      classifier <- colnames(probabilities)[which.max(probabilities[1,])]
-    }
-    else{
-      classifier <- NA
-    }
-    #each element is signature name (string) that has the maximum posterior probability
-    classify[row]<- classifier
+    tryCatch(
+      expr = {
+        #finds the index for the maximum probability and retrieves name of signature that is the max
+        classifier <- colnames(probabilities)[which.max(probabilities[1,])]
+        #each element is signature name (string) that has the maximum posterior probability
+        classify[row]<- classifier
+      },
+      error = function(e){
+        #when mutation that should not have been generated appears, probabilities and classifier are NA
+        classifier <- NA
+      }
+    )
   }
   
   #add column to the data frame that represents Bayesian classifier

### 2021-07-21 Abstract.Rmd

---
title: "Abstract Plots"
author: "Audrey Mitchell"
date: "7/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(NMF)
library(MutationalPatterns)
library(tidyverse)
library(ggrepel)
library(ROCR)
source("bayes_function.R")
source("get_classification_df.R")
source("simulation.R")
```

```{r}
#load FFPE signature as a matrix
ffpe.signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))

#create vector of the 96 mutation types
mutations <- rownames(ffpe.signature)

#Load COSMIC signatures (version 3)
cosmic.signatures <- get_known_signatures(muttype = "snv")

#Create vector of the signature names
signature.names <- colnames(cosmic.signatures)

#Create list of signature matrices with assigned rownames
signature.matrix.list <- list()
for (i in 1:60){
  sig <- cosmic.signatures[,i]
  sig.mat <- as.matrix(sig)
  rownames(sig.mat) <- mutations
  colnames(sig.mat) <- signature.names[i]
  signature.matrix.list[[i]] <- sig.mat
}

#create vector of cosine similarities to the FFPE signature
ffpe.cosmic.cos.sims <- c()
for (i in 1:60){
  cos.sim <- cos_sim_matrix(signature.matrix.list[[i]], ffpe.signature)
  ffpe.cosmic.cos.sims[i] <- cos.sim[1,1]
}
```

#Scatterplot of Average AUC vs. Cosine Similarity to FFPE
```{r}
#outer for loop: loop through each COSMIC v3 signature
average.auc.values <- c()
for (x in 1:60){
  signature.auc.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df.1 <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df.1)
    sig.name <- class.df.colnames[3]
    classification.df.2 <- classification.df.1 %>%
      mutate(signature.indicator = ifelse(truth == sig.name, 1, 0))
    predictions <- classification.df.2[3]
    labels <- classification.df.2$signature.indicator
    pred <- prediction(predictions, labels)
    auc.perf <- performance(pred, measure = "auc")
    signature.auc.vec[y] <- auc.perf@y.values[[1]][1]
  }
  average.signature.auc <- mean(signature.auc.vec)
  average.auc.values[x] <- average.signature.auc
}
```

```{r}
cos.sim.auc.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.auc.values)
colnames(cos.sim.auc.data) <- c("signature", "cosine_similarity", "average_auc")
```

```{r}
model2 <- lm(average_auc ~ cosine_similarity, data = cos.sim.auc.data)

#summary(model2)

model2_summary <- summary(model2)

cos.sim.auc.data$residuals <- model2_summary$residuals

cos.sim.auc.outliers <- cos.sim.auc.data %>%
  filter(residuals >= quantile(cos.sim.auc.data$residuals, 0.95) | 
         residuals <= quantile(cos.sim.auc.data$residuals, 0.05))
```

```{r}
ggplot(cos.sim.auc.data, aes(x = cosine_similarity, y = average_auc))+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(title = "Average area under curve vs. cosine similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average area under curve (10 iterations)")+
  geom_point(data = cos.sim.auc.outliers, color = "red")+
  geom_label_repel(data = cos.sim.auc.outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)

```

#Scatterplot of Average Sensitivity and Specificity vs. Cosine Similarity to FFPE

```{r}
#outer for loop: loop through each COSMIC v3 signature
average.sensitivity.values <- c()
average.specificity.values <- c()
for (x in 1:60){
  signature.sensitivity.vec <- c()
  signature.specificity.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df)
    sig.name <- class.df.colnames[3]
    #Count number of true positives (for sensititivity)
    sensitivity.df <- classification.df %>%
      filter(truth == sig.name & classify == sig.name)
    #Calculate sensitivity
    sensitivity.val = nrow(sensitivity.df)/100
    #Add to the vector of 10 sensitivity values
    signature.sensitivity.vec[y] = sensitivity.val
    #Count number of true negatives (for specificity)
    specificity.df <- classification.df %>%
      filter(truth == "FFPE" & classify == "FFPE")
    #Calculate specificity
    specificity.val = nrow(specificity.df)/100
    #Add to the vector of 10 specificity values
    signature.specificity.vec[y] = specificity.val
  }
  average.signature.sensitivity <- mean(signature.sensitivity.vec)
  average.sensitivity.values[x] <- average.signature.sensitivity
  average.signature.specificity <- mean(signature.specificity.vec)
  average.specificity.values[x] <- average.signature.specificity
}
```

```{r}
sens.spec.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.sensitivity.values, average.specificity.values)

colnames(sens.spec.data) <- c("signature", "cosine_similarity", "sensitivity", "specificity")

sens.spec.data.2 <- sens.spec.data %>%
  pivot_longer(c(sensitivity, specificity), names_to = "measure", values_to = "value")
```

```{r}
ggplot(data = sens.spec.data.2, aes(x = cosine_similarity, y = value, color = measure))+
  geom_point()+
  geom_smooth(method = lm, se = FALSE)+
  labs(title = "Sensitivity & Specificity vs. Cosine Similarity to FFPE Signature", x = "Cosine Similarity of COSMIC Signature to FFPE", y = "Average Sensitivity and Specificity (10 iterations)")
```

### 2021-07-21 plots.Rmd

--- 2021-07-07 plots.Rmd
@@
 output: pdf_document
+editor_options: 
+  chunk_output_type: inline
 ---
@@
 library(NMF)
+library(MutationalPatterns)
 library(tidyverse)
+library(ggrepel)
 library(ROCR)
@@
 ```
+
+#Scatterplot of misclassification rate vs. cosine similarity of COSMIC signature to FFPE signature
+
+```{r}
+#Create vector of the signature names
+signature.names <- colnames(cosmic.signatures)
+```
+
+```{r}
+#Create list of signature matrices with assigned rownames
+signature.matrix.list <- list()
+for (i in 1:60){
+  sig <- cosmic.signatures[,i]
+  sig.mat <- as.matrix(sig)
+  rownames(sig.mat) <- mutations
+  colnames(sig.mat) <- signature.names[i]
+  signature.matrix.list[[i]] <- sig.mat
+}
+```
+
+```{r}
+#outer for loop: loop through each COSMIC v3 signature
+average.misclassification.rates <- c()
+for (x in 1:60){
+  signature.misclass.vec <- c()
+  #inner for loop: calculate misclassification rate 10x for each signature
+  for (y in 1:10){
+    #create a sample vector of 180 mutations for the COSMIC signature corresponding to number x
+    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
+    #Create FFPE sample vector of 20 mutations
+    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
+    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
+    misclassification.rate <- mean(classification.df$misclassification)
+    signature.misclass.vec[y] <- misclassification.rate
+  }
+  average.signature.misclass <- mean(signature.misclass.vec)
+  average.misclassification.rates[x] <- average.signature.misclass
+}
+```
+
+```{r}
+#create vector of cosine similarities to the FFPE signature
+ffpe.cosmic.cos.sims <- c()
+for (i in 1:60){
+  cos.sim <- cos_sim_matrix(signature.matrix.list[[i]], ffpe.signature)
+  ffpe.cosmic.cos.sims[i] <- cos.sim[1,1]
+}
+```
+
+```{r}
+cos.sim.misclass.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.misclassification.rates)
+colnames(cos.sim.misclass.data) <- c("signature", "cosine_similarity", "average_misclassification")
+```
+
+```{r}
+model1 <- lm(average_misclassification ~ log(cosine_similarity), data = cos.sim.misclass.data)
+
+summary(model1)
+
+model1_summary <- summary(model1)
+
+cos.sim.misclass.data$residuals <- model1_summary$residuals
+
+cos.sim.misclass.outliers <- cos.sim.misclass.data %>%
+  filter(residuals >= quantile(cos.sim.misclass.data$residuals, 0.90) | 
+         residuals <= quantile(cos.sim.misclass.data$residuals, 0.1))
+```
+
+```{r}
+ggplot(cos.sim.misclass.data, aes(x = cosine_similarity, y = average_misclassification))+
+  geom_point()+
+  geom_smooth(method = "lm")+
+  labs(title = "Average misclassification rate vs. cosine similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average misclassification rate (10 iterations)")+
+  scale_x_log10()+
+  geom_point(data = cos.sim.misclass.outliers, color = "red")+
+  geom_label_repel(data = cos.sim.misclass.outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)
+
+```
+
+#Scatterplot of Average AUC vs. Cosine Similarity to FFPE
+```{r}
+#outer for loop: loop through each COSMIC v3 signature
+average.auc.values <- c()
+for (x in 1:60){
+  signature.auc.vec <- c()
+  #inner for loop: calculate misclassification rate 10x for each signature
+  for (y in 1:10){
+    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
+    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
+    #Create FFPE sample vector of 100 mutations
+    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
+    classification.df.1 <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
+    class.df.colnames <- colnames(classification.df.1)
+    sig.name <- class.df.colnames[3]
+    classification.df.2 <- classification.df.1 %>%
+      mutate(signature.indicator = ifelse(truth == sig.name, 1, 0))
+    predictions <- classification.df.2[3]
+    labels <- classification.df.2$signature.indicator
+    pred <- prediction(predictions, labels)
+    auc.perf <- performance(pred, measure = "auc")
+    signature.auc.vec[y] <- auc.perf@y.values[[1]][1]
+  }
+  average.signature.auc <- mean(signature.auc.vec)
+  average.auc.values[x] <- average.signature.auc
+}
+```
+
+```{r}
+cos.sim.auc.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.auc.values)
+colnames(cos.sim.auc.data) <- c("signature", "cosine_similarity", "average_auc")
+```
+
+```{r}
+model2 <- lm(average_auc ~ cosine_similarity, data = cos.sim.auc.data)
+
+summary(model2)
+
+model2_summary <- summary(model2)
+
+cos.sim.auc.data$residuals <- model2_summary$residuals
+
+cos.sim.auc.outliers <- cos.sim.auc.data %>%
+  filter(residuals >= quantile(cos.sim.auc.data$residuals, 0.95) | 
+         residuals <= quantile(cos.sim.auc.data$residuals, 0.05))
+```
+
+```{r}
+ggplot(cos.sim.auc.data, aes(x = cosine_similarity, y = average_auc))+
+  geom_point()+
+  geom_smooth(method = "lm")+
+  labs(title = "Average area under curve vs. cosine similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average area under curve (10 iterations)")+
+  geom_point(data = cos.sim.auc.outliers, color = "red")+
+  geom_label_repel(data = cos.sim.auc.outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)
+
+```
+
+#Varying total number of mutations (Signature 4)
+
+```{r}
+signature.4.mat <- as.matrix(cosmic.signatures[,"SBS4"])
+rownames(signature.4.mat) <- mutations
+plot_96_profile(signature.4.mat)
+```
+
+```{r}
+average.auc.values <- c()
+i = 1
+for(total_mutations in seq(100, 1000, by = 50)){
+  auc.values <- c()
+  for(x in 1:10){
+    mutation.number <- total_mutations/2
+    sig.4.sample.vec <- create_signature_sample_vector(signature.4.mat, mutation.number)
+    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, mutation.number)
+    classification <- get_classification_df(list(sig.4.sample.vec, ffpe.sample.vec), c("SBS4","FFPE"), list(signature.4.mat, ffpe.signature))
+    classification <- classification %>%
+      mutate(indicator = ifelse(truth == "SBS4", 1, 0))
+    predictions <- classification$SBS4
+    labels <- classification$indicator
+    pred = prediction(predictions, labels)
+    auc.perf <- performance(pred, measure = "auc")
+    auc.val <- auc.perf@y.values[[1]][1]
+    auc.values[x] <- auc.val
+  }
+  average.auc.val <- mean(auc.values)
+  average.auc.values[i] <- average.auc.val
+  i = i + 1
+
+}
+```
+
+```{r}
+total_mutations <- seq(100, 1000, by = 50)
+total_mutations_df <- data.frame(total_mutations, average.auc.values)
+colnames(total_mutations_df) <- c("total_mutations", "average_auc")
+
+#plot(total_mutations_df$total_mutations, total_mutations_df$average.auc.values)
+```
+
+```{r}
+ggplot(data = total_mutations_df, aes(x = total_mutations, y = average_auc))+
+  geom_point()+
+  geom_smooth(method = "lm", se = TRUE)
+```
+
+#Varying total number of mutations (Signature 6)
+
+```{r}
+signature.6.mat <- as.matrix(cosmic.signatures[,"SBS6"])
+rownames(signature.6.mat) <- mutations
+plot_96_profile(signature.6.mat)
+```
+
+```{r}
+average.auc.values6 <- c()
+i = 1
+for(total_mutations in seq(100, 1000, by = 50)){
+  auc.values6 <- c()
+  for(x in 1:10){
+    mutation.number <- total_mutations/2
+    sig.6.sample.vec <- create_signature_sample_vector(signature.6.mat, mutation.number)
+    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, mutation.number)
+    classification <- get_classification_df(list(sig.6.sample.vec, ffpe.sample.vec), c("SBS6","FFPE"), list(signature.6.mat, ffpe.signature))
+    classification <- classification %>%
+      mutate(indicator = ifelse(truth == "SBS6", 1, 0))
+    predictions <- classification$SBS6
+    labels <- classification$indicator
+    pred = prediction(predictions, labels)
+    auc.perf <- performance(pred, measure = "auc")
+    auc.val <- auc.perf@y.values[[1]][1]
+    auc.values[x] <- auc.val
+  }
+  average.auc.val <- mean(auc.values)
+  average.auc.values6[i] <- average.auc.val
+  i = i + 1
+
+}
+```
+
+```{r}
+total_mutations6 <- seq(100, 1000, by = 50)
+total_mutations_df6 <- data.frame(total_mutations6, average.auc.values6)
+colnames(total_mutations_df6) <- c("total_mutations", "average_auc")
+
+#plot(total_mutations_df$total_mutations, total_mutations_df$average.auc.values)
+```
+
+```{r}
+ggplot(data = total_mutations_df6, aes(x = total_mutations, y = average_auc))+
+  geom_point()
+  #geom_smooth(method = "lm")
+```
+
+#Diagram of Method
+- We start with a mutational profile: create a sample of 200 MMR, 200 Tobacco, 100 FFPE, plot the mutational profile
+
+```{r}
+mmr.sample <- create_signature_sample_vector(signature.6.mat, 200)
+tobacco.sample <- create_signature_sample_vector(signature.4.mat, 200)
+ffpe.sample <- create_signature_sample_vector(ffpe.signature, 100)
+
+ex.profile.mat <- vector_to_matrix(c(mmr.sample, tobacco.sample))
+rownames(ex.profile.mat) <- mutations
+
+plot_96_profile(ex.profile.mat)+
+  labs(title = "Mutational Profile of Original Simulated Sample")+
+  theme(aspect.ratio = 1)
+```
+
+```{r}
+ex.ffpe.profile.mat <- vector_to_matrix(c(mmr.sample, tobacco.sample, ffpe.sample))
+rownames(ex.ffpe.profile.mat) <- mutations
+
+plot_96_profile(ex.ffpe.profile.mat)+
+  labs(title = "Mutational Profile After Addition of FFPE Mutations")+
+  theme(aspect.ratio = 1)
+```
+
+- We use extract_signatures or fit_to_signatures from the MutationalPatterns package to determine mutational signatures and their contributions: plot signatures and contributions
+
+```{r}
+sig.mat <- cbind(signature.4.mat, signature.6.mat, ffpe.signature)
+colnames(sig.mat) <- c("Signature 4", "Signature 6", "FFPE")
+fts.res <- fit_to_signatures(ex.ffpe.profile.mat, sig.mat)
+
+plot_contribution(fts.res$contribution, sig.mat, coord_flip = TRUE)+
+  labs(title = "Relative Contributions of Signatures")+
+  theme(aspect.ratio = 0.3)
+```
+
+```{r}
+plot_96_profile(sig.mat)+
+  labs(title = "Mutational Signatures")
+```
+
+- We use Bayes formula to calculate the posterior probability of each signature given each of the 96 mutation types; prior, likelihood, and marginalization are obtained from the mutational signatures and contributions
+- We classify which signature each mutation is due to based on the signature with the hightest posterior probability: show part of the classification data frame
+
+```{r}
+classifications <- get_classification_df(list(tobacco.sample, mmr.sample, ffpe.sample), c("Signature 4", "Signature 6", "FFPE"), list(signature.4.mat, signature.6.mat, ffpe.signature))
+
+head(classifications)
+```
+
+- Mutations classified as FFPE are filtered out, and we obtain a reconstructed profile: plot reconstructed profile
+
+```{r}
+filter.ffpe <- classifications %>%
+  filter(classify != "FFPE")
+
+recon.vec <- filter.ffpe$mutations
+recon.mat <- vector_to_matrix(recon.vec)
+rownames(recon.mat) <- mutations
+
+plot_96_profile(recon.mat)+
+  labs(title = "Reconstructed Profile After Removing FFPE Classifed Mutations")
+```
+
+```{r}
+cos_sim_matrix(ex.profile.mat, recon.mat)
+```
+
+#Description of Simulation
+- Plot original signature
+- Plot profile of sample created to match the signature distribution
+- cosine similarity of sample to original signature
+
+```{r}
+x1 <- plot_96_profile(signature.4.mat)+
+  labs(title = "Signature 4")
+x2 <- plot_96_profile(signature.6.mat)+
+  labs(title = "Signature 6")
+tobacco.1000 <- create_signature_sample_matrix(signature.4.mat, 1000)
+mmr.1000 <- create_signature_sample_matrix(signature.6.mat, 1000)
+x3 <- plot_96_profile(tobacco.1000)+
+  labs(title = "Signature 4 Simulated Sample Profile")
+x4 <- plot_96_profile(mmr.1000)+
+  labs(title = "Signature 6 Simulated Sample Profile")
+grid.arrange(x1, x2, x3, x4, nrow = 2)
+```
+
+#Profile Reconstruction for MMR
+- Create sample of 500 MMR mutations
+- Plot mutational profile of sample
+- Add 500 FFPE mutations to the sample
+- Plot FFPE-added profile
+- Run classifier function
+- Filter out FFPE mutations
+- Plot reconstructed profile
+- Calculate the cosine similarity between the original and recosntructed
+
+```{r}
+mmr.500 <- create_signature_sample_vector(signature.6.mat, 500)
+mmr.mat <- vector_to_matrix(mmr.500)
+rownames(mmr.mat) <- mutations
+
+m1 <- plot_96_profile(mmr.mat)+
+  labs(title = "Mutational Profile of Signature 6 Simulated Sample")+
+  theme(axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        plot.title = element_text(size = 10))
+```
+
+```{r}
+ffpe.500 <- create_signature_sample_vector(ffpe.signature, 500)
+mmr.ffpe.samp <- c(mmr.500, ffpe.500)
+mmr.ffpe.mat <- vector_to_matrix(mmr.ffpe.samp)
+rownames(mmr.ffpe.mat) <- mutations
+m2 <- plot_96_profile(mmr.ffpe.mat)+
+  labs(title = "Mutational Profile After Addition of FFPE")+
+  theme(axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        plot.title = element_text(size = 10))
+```
+
+```{r}
+mmr.ffpe.class <- get_classification_df(list(mmr.500, ffpe.500), c("Signature 6", "FFPE"), list(signature.6.mat, ffpe.signature))
+
+mmr.only <- mmr.ffpe.class %>%
+  filter(classify != "FFPE")
+
+recon.mmr.vec <- mmr.only$mutations
+recon.mmr.matrix <- vector_to_matrix(recon.mmr.vec)
+rownames(recon.mmr.matrix) <- mutations
+
+m3 <- plot_96_profile(recon.mmr.matrix)+
+  labs(title = "Reconstructed Mutational Profile (FFPE Mutations Removed)")+
+  theme(axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        plot.title = element_text(size = 10))
+```
+
+```{r}
+grid.arrange(m1,m2,m3, nrow = 3)
+```
+
+#Profile Reconstruction for Signature 4
+- Create sample of 500 Signature 4 mutations
+- Plot mutational profile of sample
+- Add 500 FFPE mutations to the sample
+- Plot FFPE-added profile
+- Run classifier function
+- Filter out FFPE mutations
+- Plot reconstructed profile
+- Calculate the cosine similarity between the original and recosntructed
+
+```{r}
+tobacco.500 <- create_signature_sample_vector(signature.4.mat, 500)
+tobacco.mat <- vector_to_matrix(tobacco.500)
+
+t1 <- plot_96_profile(tobacco.mat)+
+  labs(title = "Mutational Profile of Signature 4 Simulated Sample", x = "Dinucleotide Context", y = "Probability")+
+  theme(axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        plot.title = element_text(size = 10))
+t1
+```
+
+```{r}
+tobacco.ffpe.vec <- c(tobacco.500, ffpe.500)
+tobacco.ffpe.mat <- vector_to_matrix(tobacco.ffpe.vec)
+rownames(tobacco.ffpe.mat) <- mutations
+
+t2 <- plot_96_profile(tobacco.ffpe.mat)+
+  labs(title = "Mutational Profile After Addition of FFPE", x = "Dinucleotide Context", y = "Probability")+
+  theme(axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        plot.title = element_text(size = 10))
+```
+
+```{r}
+tobacco.ffpe.class <- get_classification_df(list(tobacco.500, ffpe.500), c("Signature 4", "FFPE"), list(signature.4.mat, ffpe.signature))
+
+tobacco.only <- tobacco.ffpe.class %>%
+  filter(classify != "FFPE")
+
+recon.tobacco.vec <- tobacco.only$mutations
+recon.tobacco.mat <- vector_to_matrix(recon.tobacco.vec)
+rownames(recon.tobacco.mat) <- mutations
+
+t3 <- plot_96_profile(recon.tobacco.mat)+
+  labs(title = "Reconstructed Mutational Profile (FFPE Mutations Removed)", x = "Dinucleotide Context", y = "Probability")+
+  theme(axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        plot.title = element_text(size = 10))
+```
+
+```{r}
+grid.arrange(t1, t2, t3)
+```
+
+#Scatterplot of Cosine similarity between recosntructed signature and original vs. cosine similarity between signature and the FFPE signature
+- For each signature, iterate through sample creation and profile reconstruction 10 times, calculate cosine similarity between original and reconstruction
+- Take average of 10 cosine similarities, plot against the cosine similarity of the COSMIC signatures to FFPE
+
+```{r}
+#outer for loop: loop through each COSMIC v3 signature
+average.auc.values <- c()
+for (x in 1:60){
+  signature.auc.vec <- c()
+  #inner for loop: calculate misclassification rate 10x for each signature
+  for (y in 1:10){
+    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
+    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
+    #Create FFPE sample vector of 100 mutations
+    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
+    classification.df.1 <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
+    class.df.colnames <- colnames(classification.df.1)
+    sig.name <- class.df.colnames[3]
+    classification.df.2 <- classification.df.1 %>%
+      mutate(signature.indicator = ifelse(truth == sig.name, 1, 0))
+    predictions <- classification.df.2[3]
+    labels <- classification.df.2$signature.indicator
+    pred <- prediction(predictions, labels)
+    auc.perf <- performance(pred, measure = "auc")
+    signature.auc.vec[y] <- auc.perf@y.values[[1]][1]
+  }
+  average.signature.auc <- mean(signature.auc.vec)
+  average.auc.values[x] <- average.signature.auc
+}
+```
+
+#Scatterplot of Sensitivity and Specificity vs. Cosine similarity of COSMIC signature to FFPE
+
+```{r}
+#outer for loop: loop through each COSMIC v3 signature
+average.sensitivity.values <- c()
+average.specificity.values <- c()
+for (x in 1:60){
+  signature.sensitivity.vec <- c()
+  signature.specificity.vec <- c()
+  #inner for loop: calculate misclassification rate 10x for each signature
+  for (y in 1:10){
+    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
+    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
+    #Create FFPE sample vector of 100 mutations
+    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
+    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
+    class.df.colnames <- colnames(classification.df)
+    sig.name <- class.df.colnames[3]
+    #Count number of true positives (for sensititivity)
+    sensitivity.df <- classification.df %>%
+      filter(truth == sig.name & classify == sig.name)
+    #Calculate sensitivity
+    sensitivity.val = nrow(sensitivity.df)/100
+    #Add to the vector of 10 sensitivity values
+    signature.sensitivity.vec[y] = sensitivity.val
+    #Count number of true negatives (for specificity)
+    specificity.df <- classification.df %>%
+      filter(truth == "FFPE" & classify == "FFPE")
+    #Calculate specificity
+    specificity.val = nrow(specificity.df)/100
+    #Add to the vector of 10 specificity values
+    signature.specificity.vec[y] = specificity.val
+  }
+  average.signature.sensitivity <- mean(signature.sensitivity.vec)
+  average.sensitivity.values[x] <- average.signature.sensitivity
+  average.signature.specificity <- mean(signature.specificity.vec)
+  average.specificity.values[x] <- average.signature.specificity
+}
+```
+
+```{r}
+sens.spec.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.sensitivity.values, average.specificity.values)
+
+colnames(sens.spec.data) <- c("signature", "cosine_similarity", "sensitivity", "specificity")
+
+sens.spec.data.2 <- sens.spec.data %>%
+  pivot_longer(c(sensitivity, specificity), names_to = "measure", values_to = "value")
+```
+
+```{r}
+ggplot(data = sens.spec.data.2, aes(x = cosine_similarity, y = value, color = measure))+
+  geom_point()+
+  geom_smooth(method = lm)+
+  labs(title = "Sensitivity & Specificity vs. Cosine Similarity to FFPE Signature", x = "Cosine Similarity of COSMIC Signature to FFPE", y = "Sensitivity and Specificity")
+```

### 2021-07-22 Abstract.Rmd
Additions to abstract plots

@@
 }
+
+signature.4.mat <- as.matrix(cosmic.signatures[,"SBS4"])
+rownames(signature.4.mat) <- mutations
+
+signature.6.mat <- as.matrix(cosmic.signatures[,"SBS6"])
+rownames(signature.6.mat) <- mutations
+```
+
+#Diagram of Methods
+
+```{r}
+mmr.sample <- create_signature_sample_vector(signature.6.mat, 200)
+tobacco.sample <- create_signature_sample_vector(signature.4.mat, 200)
+ffpe.sample <- create_signature_sample_vector(ffpe.signature, 100)
+
+ex.ffpe.profile.mat <- vector_to_matrix(c(mmr.sample, tobacco.sample, ffpe.sample))
+rownames(ex.ffpe.profile.mat) <- mutations
+
+plot_96_profile(ex.ffpe.profile.mat)+
+  labs(title = "Mutational Profile")
+```
+
+```{r}
+sig.mat <- cbind(signature.4.mat, signature.6.mat, ffpe.signature)
+colnames(sig.mat) <- c("Signature 4", "Signature 6", "FFPE")
+fts.res <- fit_to_signatures(ex.ffpe.profile.mat, sig.mat)
+
+plot_contribution(fts.res$contribution, sig.mat, coord_flip = TRUE)+
+  labs(title = "Relative Contributions of Signatures")+
+  theme(aspect.ratio = 0.3)
+```
+
+```{r}
+plot_96_profile(sig.mat)+
+  labs(title = "Mutational Signatures")
+```
+
+#Diagram of Simulating Data
+```{r}
+plot_96_profile(signature.4.mat)+
+  labs(title = "Signature 4")
+```
+
+```{r}
+plot_96_profile(ffpe.signature)+
+  labs(title = "FFPE Signature")
+```
+
+```{r}
+sample.4 <- create_signature_sample_vector(signature.4.mat, 100)
+sample.ffpe <- create_signature_sample_vector(ffpe.signature, 100)
+
+ex.sim.sample.vec <- c(sample.4, sample.ffpe)
+ex.sim.sample <- vector_to_matrix(ex.sim.sample.vec)
+plot_96_profile(ex.sim.sample)+
+  labs(title = "Simulated Sample Profile")
 ```

### 2021-07-22 Abstract.Rmd
Added specificity/sensitivity analysis for dilutions. Also changed loops that go through all signatures to be more universal (get_known_signature() does not always retrieve 60 signatures).

--- 2021-07-21 Abstract.Rmd
@@
 signature.matrix.list <- list()
-for (i in 1:60){
+for (i in 1:length(signature.names)){
   sig <- cosmic.signatures[,i]
@@
 ffpe.cosmic.cos.sims <- c()
-for (i in 1:60){
+for (i in 1:length(signature.names)){
   cos.sim <- cos_sim_matrix(signature.matrix.list[[i]], ffpe.signature)
@@
 average.auc.values <- c()
-for (x in 1:60){
+for (x in 1:length(signature.names)){
   signature.auc.vec <- c()
@@
 average.specificity.values <- c()
-for (x in 1:60){
+for (x in 1:length(signature.names)){
   signature.sensitivity.vec <- c()
@@
 ```
+
+# Dilution performance characteristics
+
+```{r}
+cosmic.sig4 = signature.matrix.list[[4]]
+cosmic.sig6 = signature.matrix.list[[6]]
+ffpe.sig = ffpe.signature
+```
+
+## Dilution predictions
+
+```{r}
+num.mut=1000
+probabilities.MMR <- vector(mode = "list", length = 99)
+truth.MMR <- vector(mode = "list", length = 99)
+
+for (x in seq(10, num.mut-1, by = 10)){
+  sample.sig6 <- create_signature_sample_vector(cosmic.sig6, x)
+  sig6.truth <- rep("MMR", x)
+  sample.ffpe <- create_signature_sample_vector(ffpe.sig, num.mut-x)
+  ffpe.truth <- rep("FFPE", num.mut-x)
+  test <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"),
+                                list(cosmic.sig6, ffpe.sig))
+  classification <- test %>%
+    mutate(trueBinary = ifelse(truth == "MMR",1,0))
+
+  probabilities.MMR[[x/10]] <- classification$MMR
+  truth.MMR[[x/10]] <- classification$trueBinary
+}
+```
+
+```{r}
+num.mut=1000
+probabilities.tob <- vector(mode = "list", length = 99)
+truth.tob <- vector(mode = "list", length = 99)
+
+for (x in seq(10, num.mut-1, by = 10)){
+  sample.sig4 <- create_signature_sample_vector(cosmic.sig4, x)
+  sig4.truth <- rep("Tobacco", x)
+  sample.ffpe <- create_signature_sample_vector(ffpe.sig, num.mut-x)
+  ffpe.truth <- rep("FFPE", num.mut-x)
+  test <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"),
+                                list(cosmic.sig4, ffpe.sig))
+  classification <- test %>%
+    mutate(trueBinary = ifelse(truth == "Tobacco",1,0))
+
+  probabilities.tob[[x/10]] <- classification$Tobacco
+  truth.tob[[x/10]] <- classification$trueBinary
+}
+```
+
+```{r}
+test.MMR <- prediction(probabilities.MMR, truth.MMR)
+test.tob <- prediction(probabilities.tob, truth.tob)
+```
+
+## Sensitivity
+
+### Sensitivity/MMR
+
+```{r}
+AVG.sens.MMR <- vector(mode = "list", length = 99)
+
+for (iteration in 1:20){
+  sens.MMR <- performance(test.MMR, measure = "sens")
+
+  for(ind in 1:length(sens.MMR@x.values)){
+    thresh = as.vector(slot(sens.MMR, "x.values")[[ind]])
+    sens = as.vector(slot(sens.MMR, "y.values")[[ind]])
+    if (length(thresh) > 2){
+      if (identical(which(thresh<=0.5), integer(0))){
+        AVG.sens.MMR[[ind]][iteration] = NA
+      }
+      else{
+        low.ind = max(which(thresh >=0.5))
+        high.ind = min(which(thresh <=0.5))
+        #point slop formula of a line
+        slope = (sens[high.ind] - sens[low.ind]) / (thresh[high.ind] - thresh[low.ind])
+        AVG.sens.MMR[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + sens[low.ind]
+      }
+    }
+    else{
+      AVG.sens.MMR[[ind]][iteration] = NA
+    }
+  }
+}
+
+sens.MMR.vec = c()
+for (avg in 1:length(AVG.sens.MMR)){
+  sens.MMR.vec[avg] = mean(AVG.sens.MMR[[avg]], na.rm = TRUE)
+}
+```
+
+### Sensitivity/Tobacco
+
+```{r}
+AVG.sens.tob <- vector(mode = "list", length = 99)
+
+for(iteration in 1:20){
+  sens.tob <- performance(test.tob, measure = "sens")
+
+  for(ind in 1:length(sens.tob@x.values)){
+    thresh = as.vector(slot(sens.tob, "x.values")[[ind]])
+    sens = as.vector(slot(sens.tob, "y.values")[[ind]])
+    if (length(thresh) > 2){
+      if (identical(which(thresh<=0.5), integer(0))){
+        AVG.sens.tob[[ind]][iteration] = NA
+      }
+      else{
+        low.ind = max(which(thresh >=0.5))
+        high.ind = min(which(thresh <=0.5))
+        #point slop formula of a line
+        slope = (sens[high.ind] - sens[low.ind]) / (thresh[high.ind] - thresh[low.ind])
+        AVG.sens.tob[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + sens[low.ind]
+      }
+    }
+    else{
+      AVG.sens.tob[[ind]][iteration] = NA
+    }
+  }
+}
+
+sens.tob.vec = c()
+for (avg in 1:length(AVG.sens.tob)){
+  sens.tob.vec[avg] = mean(AVG.sens.tob[[avg]], na.rm = TRUE)
+}
+```
+
+### Sensitivity across signatures
+
+```{r}
+plot(1:99, sens.MMR.vec, ylim = c(0,1), xlab = "Percent of each signature", 
+     ylab = "Sensitivity value", type = "l", main = "Comparing performance characteristics")
+lines(1:99, sens.tob.vec, lty = 2)
+abline(h = 0.85, col = "red")
+legend("bottomright", c("MMR sensitivity", "Tobacco sensitivity", 
+                        "Satisfactory Sensitivity"),
+       lty = c(1,2,1), col = c("black", "black", "red"))
+```
+
+## Specificity
+
+### Specificity/MMR
+
+```{r}
+AVG.spec.MMR <- vector(mode = "list", length = 99)
+
+for(iteration in 1:20){
+  spec.MMR <- performance(test.MMR, measure = "spec")
+
+  for(ind in 1:length(spec.MMR@x.values)){
+    thresh = as.vector(slot(spec.MMR, "x.values")[[ind]])
+    spec = as.vector(slot(spec.MMR, "y.values")[[ind]])
+    if (length(thresh) > 2){
+      if (identical(which(thresh<=0.5), integer(0))){
+        AVG.spec.MMR[[ind]][iteration] = NA
+      }
+      else{
+        low.ind = max(which(thresh >=0.5))
+        high.ind = min(which(thresh <=0.5))
+        #point slop formula of a line
+        slope = (spec[high.ind] - spec[low.ind]) / (thresh[high.ind] - thresh[low.ind])
+        AVG.spec.MMR[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + spec[low.ind]
+      }
+    }
+    else{
+      AVG.spec.MMR[[ind]][iteration] = NA
+    }
+  }
+}
+
+spec.MMR.vec = c()
+for (avg in 1:length(AVG.spec.MMR)){
+  spec.MMR.vec[avg] = mean(AVG.spec.MMR[[avg]], na.rm = TRUE)
+}
+```
+
+### Specificity/Tobacco
+
+```{r}
+AVG.spec.tob <- vector(mode = "list", length = 99)
+
+for (iteration in 1:20){
+  spec.tob <- performance(test.tob, measure = "spec")
+
+  for(ind in 1:length(spec.tob@x.values)){
+    thresh = as.vector(slot(spec.tob, "x.values")[[ind]])
+    spec = as.vector(slot(spec.tob, "y.values")[[ind]])
+    if (length(thresh) > 2){
+      if (identical(which(thresh<=0.5), integer(0))){
+        AVG.spec.tob[[ind]][iteration] = NA
+      }
+      else{
+        low.ind = max(which(thresh >=0.5))
+        high.ind = min(which(thresh <=0.5))
+        #point slop formula of a line
+        slope = (spec[high.ind] - spec[low.ind]) / (thresh[high.ind] - thresh[low.ind])
+        AVG.spec.tob[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + spec[low.ind]
+      }
+    }
+    else{
+      AVG.spec.tob[[ind]][iteration] = NA
+    }
+  }
+}
+
+spec.tob.vec = c()
+for (avg in 1:length(AVG.spec.tob)){
+  spec.tob.vec[avg] = mean(AVG.spec.tob[[avg]], na.rm = TRUE)
+}
+```
+
+### Specificity across all signatures
+
+```{r}
+plot(1:99, spec.MMR.vec, ylim = c(0,1), xlab = "Percent of each signature", 
+     ylab = "Specificity value", type = "l", main = "Comparing performance characteristics")
+lines(1:99, spec.tob.vec, lty = 2)
+abline(h = 0.85, col = "red")
+legend("bottomleft", c("MMR specificity", "Tobacco specificity", 
+                       "Satisfactory Specificity"),
+       lty = c(1,2,1), col = c("black", "black", "red"))
+```

### 2021-07-22 Abstract.Rmd
New plots

--- 2021-07-22 Abstract.Rmd
@@
 source("simulation.R")
+```
+
+```{r}
+vector_to_matrix <- function(mutations.vector){
+  tab <- table(mutations.vector)/sum(table(mutations.vector))
+  #transform table to a data frame and assign column names to mutations and frequencies
+  mutations.df <- data.frame(tab)
+  colnames(mutations.df) <- c("mutations", "frequencies")
+
+  #defines a function that joins two data frames and replaces NA with 0
+  left_join_NA <- function(x, y, by) {
+    left_join(x = x, y = y, by = by) %>% 
+      mutate_each(funs(replace(., which(is.na(.)), 0)))
+  }
+
+  #converts vector of 96 mutation types to a data frame
+  mutation.types.df <- data.frame(mutations)
+  #joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
+  complete.mutations.df <- left_join_NA(mutation.types.df, mutations.df, by = "mutations")
+
+  #converts data frame of mutation types and frequencies to a data frame and assigns rownames
+  mutations.matrix <- as.matrix(complete.mutations.df$frequencies)
+  rownames(mutations.matrix) = mutations
+
+  return(mutations.matrix)
+}
 ```

### 2021-07-22 Abstract.Rmd
Cleaned dilution plots.

--- 2021-07-22 Abstract.Rmd
@@
 for (x in seq(10, num.mut-1, by = 10)){
-  sample.sig6 <- create_signature_sample_vector(cosmic.sig6, x)
-  sig6.truth <- rep("MMR", x)
-  sample.ffpe <- create_signature_sample_vector(ffpe.sig, num.mut-x)
-  ffpe.truth <- rep("FFPE", num.mut-x)
+  sample.ffpe <- create_signature_sample_vector(ffpe.sig, x)
+  ffpe.truth <- rep("FFPE", x)
+  sample.sig6 <- create_signature_sample_vector(cosmic.sig6, num.mut-x)
+  sig6.truth <- rep("MMR", num.mut-x)
   test <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"),
@@
 for (x in seq(10, num.mut-1, by = 10)){
-  sample.sig4 <- create_signature_sample_vector(cosmic.sig4, x)
-  sig4.truth <- rep("Tobacco", x)
-  sample.ffpe <- create_signature_sample_vector(ffpe.sig, num.mut-x)
-  ffpe.truth <- rep("FFPE", num.mut-x)
+  sample.ffpe <- create_signature_sample_vector(ffpe.sig, x)
+  ffpe.truth <- rep("FFPE", x)
+  sample.sig4 <- create_signature_sample_vector(cosmic.sig4, num.mut-x)
+  sig4.truth <- rep("Tobacco", num.mut-x)
   test <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"),
@@
 
-for (iteration in 1:20){
+for (iteration in 1:50){
   sens.MMR <- performance(test.MMR, measure = "sens")
@@
 
-for(iteration in 1:20){
+for(iteration in 1:50){
   sens.tob <- performance(test.tob, measure = "sens")
@@
 ```{r}
-plot(1:99, sens.MMR.vec, ylim = c(0,1), xlab = "Percent of each signature", 
-     ylab = "Sensitivity value", type = "l", main = "Comparing performance characteristics")
-lines(1:99, sens.tob.vec, lty = 2)
-abline(h = 0.85, col = "red")
-legend("bottomright", c("MMR sensitivity", "Tobacco sensitivity", 
-                        "Satisfactory Sensitivity"),
-       lty = c(1,2,1), col = c("black", "black", "red"))
+sensitivity = append(sens.MMR.vec, sens.tob.vec)
+signature.sens = append(rep("MMR", length(sens.MMR.vec)),
+                        rep("Tobacco", length(sens.tob.vec)))
+dilution = append(1:99, 1:99)
+sens.df <- data.frame(dilution, sensitivity, signature.sens)
+
+ggplot(data = sens.df, mapping = aes(x = dilution, 
+                                     y = sensitivity, 
+                                     color = signature.sens)) +
+  geom_point() +
+  geom_smooth() +
+  labs(title = "Sensitivity as a function of FFPE noise",
+       x = "Percent of FFPE noise in samples",
+       y = "Sensitivity value",
+       color = "True signal")
 ```
@@
 ```{r}
-plot(1:99, spec.MMR.vec, ylim = c(0,1), xlab = "Percent of each signature", 
-     ylab = "Specificity value", type = "l", main = "Comparing performance characteristics")
-lines(1:99, spec.tob.vec, lty = 2)
-abline(h = 0.85, col = "red")
-legend("bottomleft", c("MMR specificity", "Tobacco specificity", 
-                       "Satisfactory Specificity"),
-       lty = c(1,2,1), col = c("black", "black", "red"))
+specificity = append(spec.MMR.vec, spec.tob.vec)
+signature.spec = append(rep("MMR", length(sens.MMR.vec)),
+                        rep("Tobacco", length(spec.tob.vec)))
+dilution = append(1:99, 1:99)
+spec.df <- data.frame(dilution, specificity, signature.spec)
+
+ggplot(data = spec.df, mapping = aes(x = dilution, 
+                                     y = specificity, 
+                                     color = signature.sens)) +
+  geom_point() +
+  geom_smooth() +
+  labs(title = "Specificity as a function of FFPE noise",
+       x = "Percent of FFPE noise in samples",
+       y = "Specificity value",
+       color = "True signal")
 ```

### 2021-07-23 sample_conversions.R
Added commented code that'll distribute genotype information based on the ground truth

@@ -18,6 +18,8 @@ create_gr_from_sample <- function(sample, specie, chromosome) {
   position <- vector(mode = "double")
   id <- vector(mode = "character")
   info <- vector(mode = "character")
+  sample1 <- vector(mode = "character")
+  sample2 <- vector(mode = "character")
   # Original nucleotide
   ref <- vector(mode = "character")
   # New nucleotide
@@ -37,7 +39,7 @@ create_gr_from_sample <- function(sample, specie, chromosome) {
     alt <- c(alt, str_sub(mut.reformatted, 3, 3))
     str_sub(mut.reformatted, 3, 3) <- ""
     
-    info <- c(info, paste("SOMATIC;TRUTH=", sample$truth[m], ";PROB", "=", round(sample[m, 3], 3), ",", round(sample[m, 4], 3), sep = ""))
+    info <- c(info, paste("SOMATIC", ";TRUTH=", sample$truth[m], ";PROB", "=", round(sample[m, 3], 3), ",", round(sample[m, 4], 3), sep = ""))
     
     # Searches chromosome sequence for mutation match
     while (mut.reformatted != toString(seq[i:(i+2)])) {
@@ -47,6 +49,16 @@ create_gr_from_sample <- function(sample, specie, chromosome) {
     # Once matched, store position
     position <- c(position, i+1)
     id <- c(id, paste(str_sub(chromosome, 4, 4), ":", i+1, "_", str_sub(mut.reformatted, 1, 1), "/", str_sub(mut.reformatted, 3, 3), sep=""))
+    # if (sample$truth[m] == "FFPE") {
+    #   sample1 <- c(sample1, paste("0/1:", sample(10:50, 1), sep = ""))
+    #   sample2 <- c(sample2, paste("0/1:", sample(50:100, 1), sep = ""))
+    # } else {
+    #   sample1 <- c(sample1, paste("0/1:", sample(51:100, 1), sep = ""))
+    #   sample2 <- c(sample2, paste("0/1:", sample(10:50, 1, replace=T), sep = ""))
+    # }
+    
+    sample1 <- c(sample1, paste("0/1:", sample(30:60, 1), sep = ""))
+    sample2 <- c(sample2, paste("0/1:", sample(30:60, 1), sep = ""))
   }
   
   print("...Done!")
[…]

### 2021-07-26 Abstract.Rmd
Added ROC curve for tobacco and MMR.

@@
   labs(title = "Simulated Sample Profile")
+```
+
+# ROC curve
+
+```{r}
+sample.sig6 <- create_signature_sample_vector(signature.6.mat, 500)
+sig6.truth <- rep("MMR", 500)
+sample.ffpe <- create_signature_sample_vector(ffpe.signature, 500)
+ffpe.truth <- rep("FFPE", 500)
+MMR.ROC <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"),
+                              list(signature.matrix.list[[6]], ffpe.sig))
+
+sample.sig4 <- create_signature_sample_vector(signature.4.mat, 500)
+sig4.truth <- rep("Tobacco", 500)
+tob.ROC <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"),
+                              list(signature.matrix.list[[4]], ffpe.sig))
+```
+
+```{r}
+MMR.ROC <- MMR.ROC %>%
+  #create binary labels
+  mutate(truthMMR = ifelse(truth == "MMR",1,0))
+
+tob.ROC <- tob.ROC %>%
+  mutate(truthTob = ifelse(truth == "Tobacco",1,0))
+
+#provide predictions and labels
+MMR.pred <- prediction(MMR.ROC$MMR, MMR.ROC$truthMMR)
+MMR.perf <- performance(test.pred, measure = "tpr", x.measure = "fpr")
+tob.pred <- prediction(tob.ROC$Tobacco, tob.ROC$truthTob)
+tob.perf <- performance(tob.pred, measure = "tpr", x.measure = "fpr")
+plot(MMR.perf, colorize=TRUE, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7),
+     main = "ROC curve of classifier", lty = 2) #ROC
+plot(tob.perf, colorize=TRUE, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7), 
+     add = TRUE, lty = 3)
+abline(a=0, b= 1)
+legend("bottom", c("MMR with FFPE noise", "Tobacco with FFPE noise"),
+       lty = 2:3)
 ```
@@
        y = "Sensitivity value",
-       color = "True signal")
+       color = "True signal") +
+  coord_cartesian(ylim = c(0,1))
 ```
@@
        y = "Specificity value",
-       color = "True signal")
+       color = "True signal") +
+  coord_cartesian(ylim = c(0,1))
 ```

### 2021-07-27 Abstract.Rmd
changes to plots

--- 2021-07-22 Abstract.Rmd
@@
 cos.sim.auc.outliers <- cos.sim.auc.data %>%
-  filter(residuals >= quantile(cos.sim.auc.data$residuals, 0.95) | 
-         residuals <= quantile(cos.sim.auc.data$residuals, 0.05))
+  filter(residuals >= quantile(cos.sim.auc.data$residuals, 0.90) | 
+         residuals <= quantile(cos.sim.auc.data$residuals, 0.10) |
+           signature %in% c("SBS4"))
 ```
@@
 for (x in seq(10, num.mut-1, by = 10)){
-  sample.ffpe <- create_signature_sample_vector(ffpe.sig, x)
-  ffpe.truth <- rep("FFPE", x)
-  sample.sig6 <- create_signature_sample_vector(cosmic.sig6, num.mut-x)
-  sig6.truth <- rep("MMR", num.mut-x)
+  sample.sig6 <- create_signature_sample_vector(cosmic.sig6, x)
+  sig6.truth <- rep("MMR", x)
+  sample.ffpe <- create_signature_sample_vector(ffpe.sig, num.mut-x)
+  ffpe.truth <- rep("FFPE", num.mut-x)
   test <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"),
@@
 for (x in seq(10, num.mut-1, by = 10)){
-  sample.ffpe <- create_signature_sample_vector(ffpe.sig, x)
-  ffpe.truth <- rep("FFPE", x)
-  sample.sig4 <- create_signature_sample_vector(cosmic.sig4, num.mut-x)
-  sig4.truth <- rep("Tobacco", num.mut-x)
+  sample.sig4 <- create_signature_sample_vector(cosmic.sig4, x)
+  sig4.truth <- rep("Tobacco", x)
+  sample.ffpe <- create_signature_sample_vector(ffpe.sig, num.mut-x)
+  ffpe.truth <- rep("FFPE", num.mut-x)
   test <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"),
@@
 
-for (iteration in 1:50){
+for (iteration in 1:20){
   sens.MMR <- performance(test.MMR, measure = "sens")
@@
 
-for(iteration in 1:50){
+for(iteration in 1:20){
   sens.tob <- performance(test.tob, measure = "sens")
@@
 ```{r}
-sensitivity = append(sens.MMR.vec, sens.tob.vec)
-signature.sens = append(rep("MMR", length(sens.MMR.vec)),
-                        rep("Tobacco", length(sens.tob.vec)))
-dilution = append(1:99, 1:99)
-sens.df <- data.frame(dilution, sensitivity, signature.sens)
-
-ggplot(data = sens.df, mapping = aes(x = dilution, 
-                                     y = sensitivity, 
-                                     color = signature.sens)) +
-  geom_point() +
-  geom_smooth() +
-  labs(title = "Sensitivity as a function of FFPE noise",
-       x = "Percent of FFPE noise in samples",
-       y = "Sensitivity value",
-       color = "True signal")
+plot(1:99, sens.MMR.vec, ylim = c(0,1), xlab = "Percent of each signature", 
+     ylab = "Sensitivity value", type = "l", main = "Comparing performance characteristics")
+lines(1:99, sens.tob.vec, lty = 2)
+abline(h = 0.85, col = "red")
+legend("bottomright", c("MMR sensitivity", "Tobacco sensitivity", 
+                        "Satisfactory Sensitivity"),
+       lty = c(1,2,1), col = c("black", "black", "red"))
 ```
@@
 ```{r}
-specificity = append(spec.MMR.vec, spec.tob.vec)
-signature.spec = append(rep("MMR", length(sens.MMR.vec)),
-                        rep("Tobacco", length(spec.tob.vec)))
-dilution = append(1:99, 1:99)
-spec.df <- data.frame(dilution, specificity, signature.spec)
-
-ggplot(data = spec.df, mapping = aes(x = dilution, 
-                                     y = specificity, 
-                                     color = signature.sens)) +
-  geom_point() +
-  geom_smooth() +
-  labs(title = "Specificity as a function of FFPE noise",
-       x = "Percent of FFPE noise in samples",
-       y = "Specificity value",
-       color = "True signal")
+plot(1:99, spec.MMR.vec, ylim = c(0,1), xlab = "Percent of each signature", 
+     ylab = "Specificity value", type = "l", main = "Comparing performance characteristics")
+lines(1:99, spec.tob.vec, lty = 2)
+abline(h = 0.85, col = "red")
+legend("bottomleft", c("MMR specificity", "Tobacco specificity", 
+                       "Satisfactory Specificity"),
+       lty = c(1,2,1), col = c("black", "black", "red"))
 ```

### 2021-07-27 plots.Rmd
changes to plots

@@
     #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
-    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
+    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],500)
     #Create FFPE sample vector of 100 mutations
-    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
+    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 500)
     classification.df.1 <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
@@
 cos.sim.auc.outliers <- cos.sim.auc.data %>%
-  filter(residuals >= quantile(cos.sim.auc.data$residuals, 0.95) | 
-         residuals <= quantile(cos.sim.auc.data$residuals, 0.05))
+  filter(residuals >= quantile(cos.sim.auc.data$residuals, 0.90) | 
+         residuals <= quantile(cos.sim.auc.data$residuals, 0.10) | 
+         signature %in% c("SBS4"))
 ```
@@
 ```{r}
-ggplot(cos.sim.auc.data, aes(x = cosine_similarity, y = average_auc))+
+auc.scatter <- ggplot(cos.sim.auc.data, aes(x = cosine_similarity, y = average_auc))+
   geom_point()+
-  geom_smooth(method = "lm")+
-  labs(title = "Average area under curve vs. cosine similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average area under curve (10 iterations)")+
-  geom_point(data = cos.sim.auc.outliers, color = "red")+
-  geom_label_repel(data = cos.sim.auc.outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)
+  geom_smooth(method = "lm", color = mycolor[2], se = FALSE)+
+  labs(title = "Average Area Under Curve vs. Cosine Similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average area under curve (10 iterations)")+
+  geom_point(data = cos.sim.auc.outliers, color = mycolor [1])+
+  geom_label_repel(data = cos.sim.auc.outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
+  theme(plot.title = element_text(hjust = 0.5))
 
+ggsave("aucscatter.png", plot = auc.scatter, width = 175, height = 108, units = "mm")
+
+```
+
+```{r}
+cos.sim.histogram <- ggplot(data = cos.sim.auc.data, aes(x = cosine_similarity))+
+  geom_histogram(fill = mycolor[2], bins = 20)+
+  labs(title = "Distribution of Cosine Simlarities Between COSMIC Signatures and FFPE", x = "Cosine Similarity", y = "Count")+
+  theme(plot.title = element_text(hjust = 0.5))
+
+ggsave("cossimhistogram.png", plot = cos.sim.histogram, width = 175, height = 108, units = "mm")
+```
+
+```{r}
+auc.histogram <- ggplot(data = cos.sim.auc.data, aes(x = average_auc))+
+  geom_histogram(fill = mycolor[2], bins = 20)+
+  labs(title = "Distribution of Average AUC Values", x = "Average AUC", y = "Count")+
+  theme(plot.title = element_text(hjust = 0.5))
+
+ggsave("auchistogram.png", plot = auc.histogram, width = 175, height = 108, units = "mm")
 ```
@@
     #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
-    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
+    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],500)
     #Create FFPE sample vector of 100 mutations
-    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
+    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 500)
     classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
@@
     #Calculate sensitivity
-    sensitivity.val = nrow(sensitivity.df)/100
+    sensitivity.val = nrow(sensitivity.df)/500
     #Add to the vector of 10 sensitivity values
@@
     #Calculate specificity
-    specificity.val = nrow(specificity.df)/100
+    specificity.val = nrow(specificity.df)/500
     #Add to the vector of 10 specificity values
@@
   geom_point()+
-  geom_smooth(method = lm)+
-  labs(title = "Sensitivity & Specificity vs. Cosine Similarity to FFPE Signature", x = "Cosine Similarity of COSMIC Signature to FFPE", y = "Sensitivity and Specificity")
+  geom_smooth(method = lm, se = FALSE)+
+  labs(title = "Sensitivity & Specificity vs. Cosine Similarity to FFPE Signature", x = "Cosine Similarity of COSMIC Signature to FFPE", y = "Average Sensitivity and Specificity (10 iterations)")+
+  scale_color_brewer(palette = "Set1")+
+  theme(plot.title = element_text(hjust = 0.5))
 ```
+
+```{r}
+mycolor <- brewer.pal(3, "Set1")
+
+sensitivity.histogram <- ggplot(data = sens.spec.data, aes(x = sensitivity))+
+  geom_histogram(fill = mycolor[2], bins = 20)+
+  labs(title = "Distribution of Average Sensitivities", x = "Sensitivity", y = "Count")+
+  theme(plot.title = element_text(hjust = 0.5))
+
+ggsave("sensitivityhistogram.png", plot = sensitivity.histogram, width = 175, height = 108, units = "mm")
+```
+
+```{r}
+specificity.histogram <- ggplot(data = sens.spec.data, aes(x = specificity))+
+  geom_histogram(fill = mycolor[2], bins = 20)+
+  labs(title = "Distribution of Average Specificities", x = "Specificity", y = "Count")+
+  theme(plot.title = element_text(hjust = 0.5))
+
+ggsave("specificityhistogram.png", plot = specificity.histogram, width = 175, height = 108, units = "mm")
+```
+
+```{r}
+sens.spec.label.points <- sens.spec.data %>%
+  filter(signature %in% c("SBS38", "SBS4", "SBS89", "SBS3", "SBS33", "SBS44", "SBS31", "SBS40", "SBS42", "SBS23", "SBS19", "SBS11", "SBS30"))
+```
+
+```{r}
+sensitivity.scatter <- ggplot(data = sens.spec.data, mapping = aes(x = cosine_similarity, y = sensitivity))+
+  geom_point()+
+  geom_smooth(method = "lm", se = FALSE, color = mycolor[2])+
+  labs(title = "Average Sensitivity vs. Cosine Similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average sensitivity (10 iterations)")+
+  geom_point(data = sens.spec.label.points, color = mycolor [1])+
+  geom_label_repel(data = sens.spec.label.points, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
+  theme(plot.title = element_text(hjust = 0.5))
+
+sensitivity.scatter
+
+ggsave("sensitivityscatter.png", plot = sensitivity.scatter, width = 175, height = 108, units = "mm")
+```
+
+```{r}
+specificity.scatter <- ggplot(data = sens.spec.data, mapping = aes(x = cosine_similarity, y = specificity))+
+  geom_point()+
+  geom_smooth(method = "lm", se = FALSE, color = mycolor[2])+
+  labs(title = "Average Specificity vs. Cosine Similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average specificity (10 iterations)")+
+  geom_point(data = sens.spec.label.points, color = mycolor [1])+
+  geom_label_repel(data = sens.spec.label.points, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
+  theme(plot.title = element_text(hjust = 0.5))
+
+specificity.scatter
+
+ggsave("specificityscatter.png", plot = specificity.scatter, width = 175, height = 108, units = "mm")
+```

### 2021-07-27 Abstract.Rmd
additions and edits to abstract plots

@@
 date: "7/21/2021"
-output: html_document
+output: html_notebook
+editor_options: 
+  chunk_output_type: console
 ---
@@
 rownames(ex.ffpe.profile.mat) <- mutations
+colnames(ex.ffpe.profile.mat) <- c("Sample 1")
 
-plot_96_profile(ex.ffpe.profile.mat)+
-  labs(title = "Mutational Profile")
+mut.profile <- plot_96_profile(ex.ffpe.profile.mat)+
+  labs(title = "Mutational Profile", x = "Context")+
+  theme(plot.title = element_text(hjust = 0.5))
+
+mut.profile
+
+ggsave("mutprofile.png", plot = mut.profile , width = 175, height = 108, units = "mm")
 ```
@@
 
-plot_contribution(fts.res$contribution, sig.mat, coord_flip = TRUE)+
+contribution.plot <- plot_contribution(fts.res$contribution, sig.mat, coord_flip = TRUE)+
   labs(title = "Relative Contributions of Signatures")+
-  theme(aspect.ratio = 0.3)
+  theme(plot.title = element_text(hjust = 0.5))
+
+contribution.plot
+
+ggsave("contributions.png", plot = contribution.plot , width = 175, height = 54, units = "mm")
 ```
@@
 ```{r}
-plot_96_profile(sig.mat)+
-  labs(title = "Mutational Signatures")
+signatures.plot <- plot_96_profile(sig.mat)+
+  labs(title = "Mutational Signatures", x = "Context")+
+  theme(plot.title = element_text(hjust = 0.5))
+
+signatures.plot
+
+ggsave("signatures.png", plot = signatures.plot , width = 175, height = 108, units = "mm")
 ```
@@
 ```{r}
-plot_96_profile(signature.4.mat)+
-  labs(title = "Signature 4")
+signature4.plot <- plot_96_profile(signature.4.mat)+
+  labs(title = "Signature 4", x = "Context")+
+  theme(plot.title = element_text(hjust = 0.5))
+
+ggsave("signature4.png", plot = signature4.plot , width = 175, height = 81, units = "mm")
 ```
@@
 ```{r}
-plot_96_profile(ffpe.signature)+
-  labs(title = "FFPE Signature")
+ffpe.plot <- plot_96_profile(ffpe.signature)+
+  labs(title = "FFPE Signature", x = "Context")+
+  theme(plot.title = element_text(hjust = 0.5))
+
+ggsave("ffpe.png", plot = ffpe.plot , width = 175, height = 81, units = "mm")
 ```
@@
 ```{r}
-sample.4 <- create_signature_sample_vector(signature.4.mat, 100)
-sample.ffpe <- create_signature_sample_vector(ffpe.signature, 100)
+sample.4 <- create_signature_sample_vector(signature.4.mat, 500)
+sample.ffpe <- create_signature_sample_vector(ffpe.signature, 500)
 
@@
 ex.sim.sample <- vector_to_matrix(ex.sim.sample.vec)
-plot_96_profile(ex.sim.sample)+
-  labs(title = "Simulated Sample Profile")
+
+sampleprof.plot <- plot_96_profile(ex.sim.sample)+
+  labs(title = "Simulated Sample Profile", x = "Context")+
+  theme(plot.title = element_text(hjust = 0.5))
+
+sampleprof.plot
+
+ggsave("sampleprofile.png", plot = sampleprof.plot , width = 175, height = 108, units = "mm")
+```
+
+#Reconstructed Profiles of Signatures 4 and 6
+
+```{r}
+#Create Signature 4 Sample
+sig4.500 <- create_signature_sample_vector(signature.4.mat, 500)
+sig4.profile <- vector_to_matrix(sig4.500)
+
+sig4.profile.plot <- plot_96_profile(sig4.profile)+
+  labs(title = "Signature 4 Simulated Sample", x = "Context")+
+  theme(plot.title = element_text(hjust = 0.5))
+
+sig4.profile.plot
+```
+
+```{r}
+#Add FFPE Noise
+addffpe.500 <- create_signature_sample_vector(ffpe.signature, 500)
+
+sig4.ffpe <- vector_to_matrix(c(sig4.500, addffpe.500))
+
+sig4.ffpe.plot <- plot_96_profile(sig4.ffpe)+
+  labs(title = "Simulated Sample with FFPE Noise", x = "Context")+
+  theme(plot.title = element_text(hjust = 0.5))
+```
+
+```{r}
+#Classify and Filter
+sig4.ffpe.class <- get_classification_df(list(sig4.500, addffpe.500), c("SBS4", "FFPE"), list(signature.4.mat, ffpe.signature))
+
+sig4.ffpe.filtered <- sig4.ffpe.class %>%
+  filter(classify != "FFPE")
+
+sig4.recon.matrix <- vector_to_matrix(sig4.ffpe.filtered$mutations)
+
+sig4.recon.plot <- plot_96_profile(sig4.recon.matrix)+
+  labs(title = "Signature 4 Reconstructed", x = "Context")+
+  theme(plot.title = element_text(hjust = 0.5))
+
+sig4.recon.plot
+```
+
+```{r}
+#Calculate Sensitivity
+
+sig4.sensitivity <- sig4.ffpe.class %>%
+  filter(truth == "SBS4" & classify == "SBS4")
+
+nrow(sig4.sensitivity)/500
+
+sig4.specificity <- sig4.ffpe.class %>%
+  filter(truth == "FFPE" & classify == "FFPE")
+
+nrow(sig4.specificity)/500
+```
+
+```{r}
+#Arrange 3 plots
+sig4.reconstruction <- grid.arrange(sig4.profile.plot, sig4.ffpe.plot, sig4.recon.plot)
+
+ggsave("sig4reconstruction.png", plot = sig4.reconstruction , width = 175, height = 243, units = "mm")
+```
+
+```{r}
+#Create Signature 6 Sample
+sig6.500 <- create_signature_sample_vector(signature.6.mat, 500)
+sig6.profile <- vector_to_matrix(sig6.500)
+
+sig6.profile.plot <- plot_96_profile(sig6.profile)+
+  labs(title = "Signature 6 Simulated Sample", x = "Context")+
+  theme(plot.title = element_text(hjust = 0.5))
+
+sig6.profile.plot
+```
+
+```{r}
+#Add FFPE Noise
+sig6.ffpe <- vector_to_matrix(c(sig6.500, addffpe.500))
+
+sig6.ffpe.plot <- plot_96_profile(sig6.ffpe)+
+  labs(title = "Simulated Sample with FFPE Noise", x = "Context")+
+  theme(plot.title = element_text(hjust = 0.5))
+
+sig6.ffpe.plot
+```
+
+```{r}
+#Classify and Filter
+sig6.ffpe.class <- get_classification_df(list(sig6.500, addffpe.500), c("SBS6", "FFPE"), list(signature.6.mat, ffpe.signature))
+
+sig6.ffpe.filtered <- sig6.ffpe.class %>%
+  filter(classify != "FFPE")
+
+sig6.recon.matrix <- vector_to_matrix(sig6.ffpe.filtered$mutations)
+
+sig6.recon.plot <- plot_96_profile(sig6.recon.matrix)+
+  labs(title = "Signature 6 Reconstructed", x = "Context")+
+  theme(plot.title = element_text(hjust = 0.5))
+
+sig6.recon.plot
+```
+
+```{r}
+#Calculate Sensitivity
+
+sig6.sensitivity <- sig6.ffpe.class %>%
+  filter(truth == "SBS6" & classify == "SBS6")
+
+nrow(sig6.sensitivity)/500
+
+sig6.specificity <- sig6.ffpe.class %>%
+  filter(truth == "FFPE" & classify == "FFPE")
+
+nrow(sig6.specificity)/500
+```
+
+```{r}
+#Arrange 3 plots
+sig6.reconstruction <- grid.arrange(sig6.profile.plot, sig6.ffpe.plot, sig6.recon.plot)
+
+ggsave("sig6reconstruction.png", plot = sig6.reconstruction , width = 175, height = 243, units = "mm")
 ```
@@
 MMR.pred <- prediction(MMR.ROC$MMR, MMR.ROC$truthMMR)
-MMR.perf <- performance(test.pred, measure = "tpr", x.measure = "fpr")
+MMR.perf <- performance(MMR.pred, measure = "tpr", x.measure = "fpr")
 tob.pred <- prediction(tob.ROC$Tobacco, tob.ROC$truthTob)
@@
 abline(a=0, b= 1)
-legend("bottom", c("MMR with FFPE noise", "Tobacco with FFPE noise"),
+legend("bottom", c("Signature 6 with FFPE noise", "Signature 4 with FFPE noise"),
        lty = 2:3)

### 2021-07-27 Abstract.Rmd
Evaluated performance characteristics with varying mutational load.

--- 2021-07-27 Abstract.Rmd
@@
 ```
+
+# Mutational load performance characteristics
+
+## Mutational load predictions
+
+```{r}
+step = seq(50, 1000, by = 50)
+MMR.probabilities <- vector(mode = "list", length = length(step))
+MMR.truth <- vector(mode = "list", length = length(step))
+
+for (x in 1:length(step)){
+  sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2)
+  ffpe.truth <- rep("FFPE", step[x]/2)
+  sample.sig6 <- create_signature_sample_vector(cosmic.sig6, step[x]/2)
+  sig6.truth <- rep("MMR", step[x]/2)
+  test <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"),
+                                list(cosmic.sig6, ffpe.sig))
+  classification <- test %>%
+    mutate(trueBinary = ifelse(truth == "MMR",1,0))
+
+  MMR.probabilities[[x]] <- classification$MMR
+  MMR.truth[[x]] <- classification$trueBinary
+}
+```
+
+```{r}
+step = seq(50, 1000, by = 50)
+tob.probabilities <- vector(mode = "list", length = length(step))
+tob.truth <- vector(mode = "list", length = length(step))
+
+for (x in 1:length(step)){
+  sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2)
+  ffpe.truth <- rep("FFPE", step[x]/2)
+  sample.sig4 <- create_signature_sample_vector(cosmic.sig4, step[x]/2)
+  sig4.truth <- rep("tob", step[x]/2)
+  test <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"),
+                                list(cosmic.sig4, ffpe.sig))
+  classification <- test %>%
+    mutate(trueBinary = ifelse(truth == "Tobacco",1,0))
+
+  tob.probabilities[[x]] <- classification$Tobacco
+  tob.truth[[x]] <- classification$trueBinary
+}
+```
+
+```{r}
+load.MMR <- prediction(MMR.probabilities, MMR.truth)
+load.tob <- prediction(tob.probabilities, tob.truth)
+```
+
+## Sensitivity
+
+```{r}
+AVG.sens.MMR <- vector(mode = "list", length = length(step))
+
+for (iteration in 1:100){
+  sens.MMR <- performance(load.MMR, measure = "sens")
+
+  for(ind in 1:length(sens.MMR@x.values)){
+    thresh = as.vector(slot(sens.MMR, "x.values")[[ind]])
+    sens = as.vector(slot(sens.MMR, "y.values")[[ind]])
+    if (length(thresh) > 2){
+      if (identical(which(thresh<=0.5), integer(0))){
+        AVG.sens.MMR[[ind]][iteration] = NA
+      }
+      else{
+        low.ind = max(which(thresh >=0.5))
+        high.ind = min(which(thresh <=0.5))
+        #point slop formula of a line
+        slope = (sens[high.ind] - sens[low.ind]) / (thresh[high.ind] - thresh[low.ind])
+        AVG.sens.MMR[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + sens[low.ind]
+      }
+    }
+    else{
+      AVG.sens.MMR[[ind]][iteration] = NA
+    }
+  }
+}
+
+load.sens.MMR.vec = c()
+for (avg in 1:length(AVG.sens.MMR)){
+  load.sens.MMR.vec[avg] = mean(AVG.sens.MMR[[avg]], na.rm = TRUE)
+}
+```
+
+```{r}
+AVG.sens.tob <- vector(mode = "list", length = length(step))
+
+for (iteration in 1:100){
+  sens.tob <- performance(load.tob, measure = "sens")
+
+  for(ind in 1:length(sens.tob@x.values)){
+    thresh = as.vector(slot(sens.tob, "x.values")[[ind]])
+    sens = as.vector(slot(sens.tob, "y.values")[[ind]])
+    if (length(thresh) > 2){
+      if (identical(which(thresh<=0.5), integer(0))){
+        AVG.sens.tob[[ind]][iteration] = NA
+      }
+      else{
+        low.ind = max(which(thresh >=0.5))
+        high.ind = min(which(thresh <=0.5))
+        #point slop formula of a line
+        slope = (sens[high.ind] - sens[low.ind]) / (thresh[high.ind] - thresh[low.ind])
+        AVG.sens.tob[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + sens[low.ind]
+      }
+    }
+    else{
+      AVG.sens.tob[[ind]][iteration] = NA
+    }
+  }
+}
+
+load.sens.tob.vec = c()
+for (avg in 1:length(AVG.sens.tob)){
+  load.sens.tob.vec[avg] = mean(AVG.sens.tob[[avg]], na.rm = TRUE)
+}
+```
+
+```{r}
+#plot(step, load.sens.MMR.vec, ylim = c(0,1))
+sensitivity = append(load.sens.MMR.vec, load.sens.tob.vec)
+signature.sens = append(rep("MMR", length(load.sens.MMR.vec)),
+                        rep("Tobacco", length(load.sens.tob.vec)))
+dilution = append(step, step)
+sens.df <- data.frame(dilution, sensitivity, signature.sens)
+
+ggplot(data = sens.df, mapping = aes(x = dilution, 
+                                     y = sensitivity, 
+                                     color = signature.sens)) +
+  geom_point() +
+  geom_smooth() +
+  labs(title = "Specificity as a function of mutational load",
+       x = "Number of mutations in each sample",
+       y = "Specificity value") +
+  coord_cartesian(ylim = c(0,1))
+```
+
+## Specificity
+
+```{r}
+AVG.spec.MMR <- vector(mode = "list", length = length(step))
+
+for (iteration in 1:100){
+  spec.MMR <- performance(load.MMR, measure = "spec")
+
+  for(ind in 1:length(spec.MMR@x.values)){
+    thresh = as.vector(slot(spec.MMR, "x.values")[[ind]])
+    spec = as.vector(slot(spec.MMR, "y.values")[[ind]])
+    if (length(thresh) > 2){
+      if (identical(which(thresh<=0.5), integer(0))){
+        AVG.spec.MMR[[ind]][iteration] = NA
+      }
+      else{
+        low.ind = max(which(thresh >=0.5))
+        high.ind = min(which(thresh <=0.5))
+        #point slop formula of a line
+        slope = (spec[high.ind] - spec[low.ind]) / (thresh[high.ind] - thresh[low.ind])
+        AVG.spec.MMR[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + spec[low.ind]
+      }
+    }
+    else{
+      AVG.spec.MMR[[ind]][iteration] = NA
+    }
+  }
+}
+
+load.spec.MMR.vec = c()
+for (avg in 1:length(AVG.spec.MMR)){
+  load.spec.MMR.vec[avg] = mean(AVG.spec.MMR[[avg]], na.rm = TRUE)
+}
+```
+
+```{r}
+AVG.spec.tob <- vector(mode = "list", length = length(step))
+
+for (iteration in 1:100){
+  spec.tob <- performance(load.tob, measure = "spec")
+
+  for(ind in 1:length(spec.tob@x.values)){
+    thresh = as.vector(slot(spec.tob, "x.values")[[ind]])
+    spec = as.vector(slot(spec.tob, "y.values")[[ind]])
+    if (length(thresh) > 2){
+      if (identical(which(thresh<=0.5), integer(0))){
+        AVG.spec.tob[[ind]][iteration] = NA
+      }
+      else{
+        low.ind = max(which(thresh >=0.5))
+        high.ind = min(which(thresh <=0.5))
+        #point slop formula of a line
+        slope = (spec[high.ind] - spec[low.ind]) / (thresh[high.ind] - thresh[low.ind])
+        AVG.spec.tob[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + spec[low.ind]
+      }
+    }
+    else{
+      AVG.spec.tob[[ind]][iteration] = NA
+    }
+  }
+}
+
+load.spec.tob.vec = c()
+for (avg in 1:length(AVG.spec.tob)){
+  load.spec.tob.vec[avg] = mean(AVG.spec.tob[[avg]], na.rm = TRUE)
+}
+```
+
+```{r}
+#plot(step, load.spec.MMR.vec, ylim = c(0,1))
+
+specificity = append(load.spec.MMR.vec, load.spec.tob.vec)
+signature.spec = append(rep("MMR", length(load.spec.MMR.vec)),
+                        rep("Tobacco", length(load.spec.tob.vec)))
+dilution = append(step, step)
+spec.df <- data.frame(dilution, specificity, signature.spec)
+
+ggplot(data = spec.df, mapping = aes(x = dilution, 
+                                     y = specificity, 
+                                     color = signature.spec)) +
+  geom_point() +
+  geom_smooth() +
+  labs(title = "Specificity as a function of mutational load",
+       x = "Number of mutations in each sample",
+       y = "Specificity value") +
+  coord_cartesian(ylim = c(0,1))
+```

### 2021-07-27 Abstract.Rmd
Tweaked some graphs.

@@
 tob.perf <- performance(tob.pred, measure = "tpr", x.measure = "fpr")
-plot(MMR.perf, colorize=TRUE, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7),
+plot(MMR.perf, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7),
      main = "ROC curve of classifier", lty = 2) #ROC
-plot(tob.perf, colorize=TRUE, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7), 
+plot(tob.perf, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7), 
      add = TRUE, lty = 3)
@@
 ```{r}
-num.mut=1000
-probabilities.MMR <- vector(mode = "list", length = 99)
-truth.MMR <- vector(mode = "list", length = 99)
+num.mut = 1000
+step = seq(50, 950, by = 50)
+```
 
-for (x in seq(10, num.mut-1, by = 10)){
-  sample.ffpe <- create_signature_sample_vector(ffpe.sig, x)
-  ffpe.truth <- rep("FFPE", x)
-  sample.sig6 <- create_signature_sample_vector(cosmic.sig6, num.mut-x)
-  sig6.truth <- rep("MMR", num.mut-x)
+```{r}
+probabilities.MMR <- vector(mode = "list", length = length(step))
+truth.MMR <- vector(mode = "list", length = length (step))
+
+for (x in 1:length(step)){
+  sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x])
+  ffpe.truth <- rep("FFPE", step[x])
+  sample.sig6 <- create_signature_sample_vector(cosmic.sig6, num.mut-step[x])
+  sig6.truth <- rep("MMR", num.mut-step[x])
   test <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"),
@@
 
-  probabilities.MMR[[x/10]] <- classification$MMR
-  truth.MMR[[x/10]] <- classification$trueBinary
+  probabilities.MMR[[x]] <- classification$MMR
+  truth.MMR[[x]] <- classification$trueBinary
 }
@@
 ```{r}
-num.mut=1000
-probabilities.tob <- vector(mode = "list", length = 99)
-truth.tob <- vector(mode = "list", length = 99)
+probabilities.tob <- vector(mode = "list", length = length(step))
+truth.tob <- vector(mode = "list", length = length(step))
 
-for (x in seq(10, num.mut-1, by = 10)){
-  sample.ffpe <- create_signature_sample_vector(ffpe.sig, x)
-  ffpe.truth <- rep("FFPE", x)
-  sample.sig4 <- create_signature_sample_vector(cosmic.sig4, num.mut-x)
-  sig4.truth <- rep("Tobacco", num.mut-x)
+for (x in 1:length(step)){
+  sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x])
+  ffpe.truth <- rep("FFPE", step[x])
+  sample.sig4 <- create_signature_sample_vector(cosmic.sig4, num.mut-step[x])
+  sig4.truth <- rep("Tobacco", num.mut-step[x])
   test <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"),
@@
 
-  probabilities.tob[[x/10]] <- classification$Tobacco
-  truth.tob[[x/10]] <- classification$trueBinary
+  probabilities.tob[[x]] <- classification$Tobacco
+  truth.tob[[x]] <- classification$trueBinary
 }
@@
 ```{r}
-AVG.sens.MMR <- vector(mode = "list", length = 99)
+AVG.sens.MMR <- vector(mode = "list", length =length(step))
 
@@
 ```{r}
-AVG.sens.tob <- vector(mode = "list", length = 99)
+AVG.sens.tob <- vector(mode = "list", length = length(step))
 
@@
                         rep("Tobacco", length(sens.tob.vec)))
-dilution = append(1:99, 1:99)
+dilution = append(step/10, step/10)
 sens.df <- data.frame(dilution, sensitivity, signature.sens)
@@
        color = "True signal") +
-  coord_cartesian(ylim = c(0,1))
+  coord_cartesian(xlim = c(0,100), ylim = c(0,1))
 ```
@@
 ```{r}
-AVG.spec.MMR <- vector(mode = "list", length = 99)
+AVG.spec.MMR <- vector(mode = "list", length = length(step))
 
-for(iteration in 1:20){
+for(iteration in 1:50){
   spec.MMR <- performance(test.MMR, measure = "spec")
@@
 ```{r}
-AVG.spec.tob <- vector(mode = "list", length = 99)
+AVG.spec.tob <- vector(mode = "list", length = length(step))
 
-for (iteration in 1:20){
+for (iteration in 1:50){
   spec.tob <- performance(test.tob, measure = "spec")
@@
                         rep("Tobacco", length(spec.tob.vec)))
-dilution = append(1:99, 1:99)
+dilution = append(step/10, step/10)
 spec.df <- data.frame(dilution, specificity, signature.spec)
@@
   geom_smooth() +
-  labs(title = "Specificity as a function of mutational load",
+  labs(title = "Sensitivity as a function of mutational load",
        x = "Number of mutations in each sample",
-       y = "Specificity value") +
+       y = "Sensitivity value",
+       color = "True signal") +
   coord_cartesian(ylim = c(0,1))
@@
        x = "Number of mutations in each sample",
-       y = "Specificity value") +
+       y = "Specificity value",
+       color = "True signal") +
   coord_cartesian(ylim = c(0,1))

### 2021-07-29 Abstract.Rmd
Created dataframe for all signatures with certain dilutions.

@@
 
-## Sensitivity
+```{r}
+evaluate_perf <- function(num_iter, length, pred_obj, perf_method){
+  AVG.sens <- vector(mode = "list", length = length)
 
-### Sensitivity/MMR
+  for (iteration in 1:num_iter){
+  sens.obj <- performance(pred_obj, measure = perf_method)
 
-```{r}
-AVG.sens.MMR <- vector(mode = "list", length =length(step))
-
-for (iteration in 1:50){
-  sens.MMR <- performance(test.MMR, measure = "sens")
-
-  for(ind in 1:length(sens.MMR@x.values)){
-    thresh = as.vector(slot(sens.MMR, "x.values")[[ind]])
-    sens = as.vector(slot(sens.MMR, "y.values")[[ind]])
+  for(ind in 1:length(sens.obj@x.values)){
+    thresh = as.vector(slot(sens.obj, "x.values")[[ind]])
+    sens = as.vector(slot(sens.obj, "y.values")[[ind]])
     if (length(thresh) > 2){
       if (identical(which(thresh<=0.5), integer(0))){
-        AVG.sens.MMR[[ind]][iteration] = NA
+        AVG.sens[[ind]][iteration] = NA
       }
@@
         slope = (sens[high.ind] - sens[low.ind]) / (thresh[high.ind] - thresh[low.ind])
-        AVG.sens.MMR[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + sens[low.ind]
+        AVG.sens[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + sens[low.ind]
       }
@@
     else{
-      AVG.sens.MMR[[ind]][iteration] = NA
+      AVG.sens[[ind]][iteration] = NA
         }
@@
 
-sens.MMR.vec = c()
-for (avg in 1:length(AVG.sens.MMR)){
-  sens.MMR.vec[avg] = mean(AVG.sens.MMR[[avg]], na.rm = TRUE)
+  sens.vec = c()
+  for (avg in 1:length(AVG.sens)){
+    sens.vec[avg] = mean(AVG.sens[[avg]], na.rm = TRUE)
+  }
+
+  return(sens.vec)
 }
@@
 
-### Sensitivity/Tobacco
+## Sensitivity
 
 ```{r}
-AVG.sens.tob <- vector(mode = "list", length = length(step))
-
-for(iteration in 1:50){
-  sens.tob <- performance(test.tob, measure = "sens")
-
-  for(ind in 1:length(sens.tob@x.values)){
-    thresh = as.vector(slot(sens.tob, "x.values")[[ind]])
-    sens = as.vector(slot(sens.tob, "y.values")[[ind]])
-    if (length(thresh) > 2){
-      if (identical(which(thresh<=0.5), integer(0))){
-        AVG.sens.tob[[ind]][iteration] = NA
-      }
-      else{
-        low.ind = max(which(thresh >=0.5))
-        high.ind = min(which(thresh <=0.5))
-        #point slop formula of a line
-        slope = (sens[high.ind] - sens[low.ind]) / (thresh[high.ind] - thresh[low.ind])
-        AVG.sens.tob[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + sens[low.ind]
-      }
-    }
-    else{
-      AVG.sens.tob[[ind]][iteration] = NA
-    }
-  }
-}
-
-sens.tob.vec = c()
-for (avg in 1:length(AVG.sens.tob)){
-  sens.tob.vec[avg] = mean(AVG.sens.tob[[avg]], na.rm = TRUE)
-}
+sens.MMR.vec = evaluate_perf(100, length(step), test.MMR, "sens")
 ```
 
-### Sensitivity across signatures
+```{r}
+sens.tob.vec = evaluate_perf(100, length(step), test.tob, "sens")
+```
+
+### Sensitivity plot
 
@@
 sensitivity = append(sens.MMR.vec, sens.tob.vec)
-signature.sens = append(rep("MMR", length(sens.MMR.vec)),
-                        rep("Tobacco", length(sens.tob.vec)))
+signature.sens = append(rep("Signature 6", length(sens.MMR.vec)),
+                        rep("Signature 4", length(sens.tob.vec)))
 dilution = append(step/10, step/10)
@@
 
-### Specificity/MMR
-
 ```{r}
-AVG.spec.MMR <- vector(mode = "list", length = length(step))
-
-for(iteration in 1:50){
-  spec.MMR <- performance(test.MMR, measure = "spec")
-
-  for(ind in 1:length(spec.MMR@x.values)){
-    thresh = as.vector(slot(spec.MMR, "x.values")[[ind]])
-    spec = as.vector(slot(spec.MMR, "y.values")[[ind]])
-    if (length(thresh) > 2){
-      if (identical(which(thresh<=0.5), integer(0))){
-        AVG.spec.MMR[[ind]][iteration] = NA
-      }
-      else{
-        low.ind = max(which(thresh >=0.5))
-        high.ind = min(which(thresh <=0.5))
-        #point slop formula of a line
-        slope = (spec[high.ind] - spec[low.ind]) / (thresh[high.ind] - thresh[low.ind])
-        AVG.spec.MMR[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + spec[low.ind]
-      }
-    }
-    else{
-      AVG.spec.MMR[[ind]][iteration] = NA
-    }
-  }
-}
-
-spec.MMR.vec = c()
-for (avg in 1:length(AVG.spec.MMR)){
-  spec.MMR.vec[avg] = mean(AVG.spec.MMR[[avg]], na.rm = TRUE)
-}
+spec.MMR.vec = evaluate_perf(100, length(step), test.MMR, "spec")
+spec.tob.vec = evaluate_perf(100, length(step), test.tob, "spec")
 ```
 
-### Specificity/Tobacco
-
-```{r}
-AVG.spec.tob <- vector(mode = "list", length = length(step))
-
-for (iteration in 1:50){
-  spec.tob <- performance(test.tob, measure = "spec")
-
-  for(ind in 1:length(spec.tob@x.values)){
-    thresh = as.vector(slot(spec.tob, "x.values")[[ind]])
-    spec = as.vector(slot(spec.tob, "y.values")[[ind]])
-    if (length(thresh) > 2){
-      if (identical(which(thresh<=0.5), integer(0))){
-        AVG.spec.tob[[ind]][iteration] = NA
-      }
-      else{
-        low.ind = max(which(thresh >=0.5))
-        high.ind = min(which(thresh <=0.5))
-        #point slop formula of a line
-        slope = (spec[high.ind] - spec[low.ind]) / (thresh[high.ind] - thresh[low.ind])
-        AVG.spec.tob[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + spec[low.ind]
-      }
-    }
-    else{
-      AVG.spec.tob[[ind]][iteration] = NA
-    }
-  }
-}
-
-spec.tob.vec = c()
-for (avg in 1:length(AVG.spec.tob)){
-  spec.tob.vec[avg] = mean(AVG.spec.tob[[avg]], na.rm = TRUE)
-}
-```
-
-### Specificity across all signatures
+### Specificity plot
 
@@
 specificity = append(spec.MMR.vec, spec.tob.vec)
-signature.spec = append(rep("MMR", length(sens.MMR.vec)),
-                        rep("Tobacco", length(spec.tob.vec)))
+signature.spec = append(rep("Signature 6", length(sens.MMR.vec)),
+                        rep("Signature 4", length(spec.tob.vec)))
 dilution = append(step/10, step/10)
@@
 
+## File of all signatures diluted with FFPE noise
+
+```{r}
+#columns in data frame
+sens = vector(mode = "list", length = length(signature.names))
+spec = vector(mode = "list", length = length(signature.names))
+signature = vector(mode = "list", length = length(signature.names))
+
+#create data frame values 10 times
+iteration.num = 3
+for (iteration in 1:iteration.num){
+  #iterate through each signature
+  for (sig.num in 1:length(signature.names)){
+    #store name of signature (once for every dilution)
+    signature[[sig.num]] = append(signature[[sig.num]],
+                                  rep(signature.names[sig.num], 5))
+    #store values generated by get_classification_df for each dilution
+    probabilities <- vector(mode = "list", length = 5)
+    truth <- vector(mode = "list", length = 5)
+    #initialize accumulator
+    i = 1
+      #each signature is diluted with 10%, 30%, 50%, 70%, and 90% FFPE noise
+      # (1000 mutations total each time)
+      for (x in c(100, 300, 500, 700, 900)){
+        #create some percent of FFPE noise
+        sample.ffpe <- create_signature_sample_vector(ffpe.signature, x)
+        ffpe.truth <- rep("FFPE", x)
+        #fill the rest with true signal
+        sample <- create_signature_sample_vector(signature.matrix.list[[sig.num]], 1000-x)
+        sample.truth <- rep(signature.names[sig.num], 1000-x)
+        #evaluate using Bayes
+        test <- get_classification_df(list(sample, sample.ffpe), 
+                                      c(signature.names[sig.num],"FFPE"),
+                                      list(signature.matrix.list[[sig.num]],
+                                           ffpe.signature))
+        #truth column turned into binary for ROCR package
+        classification <- test %>%
+          mutate(trueBinary = ifelse(truth == signature.names[sig.num],1,0))
+        #store probabilities and truths for particular dilution
+        probabilities[[i]] <- classification[[3]]
+        truth[[i]] <- classification$trueBinary
+        i = i + 1
+      }
+    #evaluate all 5 dilution samples
+    pred = prediction(probabilities, truth)
+    sens[[sig.num]] = append(sens[[sig.num]], evaluate_perf(100, 5, pred, "sens"))
+    spec[[sig.num]] = append(spec[[sig.num]], evaluate_perf(100, 5, pred, "spec"))
+  }
+}
+
+#vector indicating dilution amount
+percent.FFPE = rep(c("10%", "30%", "50%", "70%", "90%"),
+                   length(signature.names)*iteration.num)
+#vectorize lists
+signature = unlist(signature)
+sens = unlist(sens)
+spec = unlist(spec)
+
+#convert into data frame
+table = data.frame(percent.FFPE, signature, sens, spec)
+
+#avg the multiple sensitivity/specificity values for each dilution and each signature
+table <- table %>%
+  group_by(percent.FFPE, signature) %>%
+  summarize(sens = mean(sens, na.rm = TRUE),
+            spec = mean(spec, na.rm = TRUE))
+```
+
 # Mutational load performance characteristics
@@
 ```{r}
-step = seq(50, 1000, by = 50)
+step_load = seq(50, 1000, by = 50)
 MMR.probabilities <- vector(mode = "list", length = length(step))
@@
 ```{r}
-step = seq(50, 1000, by = 50)
 tob.probabilities <- vector(mode = "list", length = length(step))
@@
 ```{r}
-AVG.sens.MMR <- vector(mode = "list", length = length(step))
-
-for (iteration in 1:100){
-  sens.MMR <- performance(load.MMR, measure = "sens")
-
-  for(ind in 1:length(sens.MMR@x.values)){
-    thresh = as.vector(slot(sens.MMR, "x.values")[[ind]])
-    sens = as.vector(slot(sens.MMR, "y.values")[[ind]])
-    if (length(thresh) > 2){
-      if (identical(which(thresh<=0.5), integer(0))){
-        AVG.sens.MMR[[ind]][iteration] = NA
-      }
-      else{
-        low.ind = max(which(thresh >=0.5))
-        high.ind = min(which(thresh <=0.5))
-        #point slop formula of a line
-        slope = (sens[high.ind] - sens[low.ind]) / (thresh[high.ind] - thresh[low.ind])
-        AVG.sens.MMR[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + sens[low.ind]
-      }
-    }
-    else{
-      AVG.sens.MMR[[ind]][iteration] = NA
-    }
-  }
-}
-
-load.sens.MMR.vec = c()
-for (avg in 1:length(AVG.sens.MMR)){
-  load.sens.MMR.vec[avg] = mean(AVG.sens.MMR[[avg]], na.rm = TRUE)
-}
-```
-
-```{r}
-AVG.sens.tob <- vector(mode = "list", length = length(step))
-
-for (iteration in 1:100){
-  sens.tob <- performance(load.tob, measure = "sens")
-
-  for(ind in 1:length(sens.tob@x.values)){
-    thresh = as.vector(slot(sens.tob, "x.values")[[ind]])
-    sens = as.vector(slot(sens.tob, "y.values")[[ind]])
-    if (length(thresh) > 2){
-      if (identical(which(thresh<=0.5), integer(0))){
-        AVG.sens.tob[[ind]][iteration] = NA
-      }
-      else{
-        low.ind = max(which(thresh >=0.5))
-        high.ind = min(which(thresh <=0.5))
-        #point slop formula of a line
-        slope = (sens[high.ind] - sens[low.ind]) / (thresh[high.ind] - thresh[low.ind])
-        AVG.sens.tob[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + sens[low.ind]
-      }
-    }
-    else{
-      AVG.sens.tob[[ind]][iteration] = NA
-    }
-  }
-}
-
-load.sens.tob.vec = c()
-for (avg in 1:length(AVG.sens.tob)){
-  load.sens.tob.vec[avg] = mean(AVG.sens.tob[[avg]], na.rm = TRUE)
-}
+load.sens.MMR.vec = evaluate_perf(100, length(step_load), load.MMR, "sens")
+load.sens.tob.vec = evaluate_perf(100, length(step_load), load.tob, "sens")
 ```
@@
 ```{r}
-AVG.spec.MMR <- vector(mode = "list", length = length(step))
-
-for (iteration in 1:100){
-  spec.MMR <- performance(load.MMR, measure = "spec")
-
-  for(ind in 1:length(spec.MMR@x.values)){
-    thresh = as.vector(slot(spec.MMR, "x.values")[[ind]])
-    spec = as.vector(slot(spec.MMR, "y.values")[[ind]])
-    if (length(thresh) > 2){
-      if (identical(which(thresh<=0.5), integer(0))){
-        AVG.spec.MMR[[ind]][iteration] = NA
-      }
-      else{
-        low.ind = max(which(thresh >=0.5))
-        high.ind = min(which(thresh <=0.5))
-        #point slop formula of a line
-        slope = (spec[high.ind] - spec[low.ind]) / (thresh[high.ind] - thresh[low.ind])
-        AVG.spec.MMR[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + spec[low.ind]
-      }
-    }
-    else{
-      AVG.spec.MMR[[ind]][iteration] = NA
-    }
-  }
-}
-
-load.spec.MMR.vec = c()
-for (avg in 1:length(AVG.spec.MMR)){
-  load.spec.MMR.vec[avg] = mean(AVG.spec.MMR[[avg]], na.rm = TRUE)
-}
-```
-
-```{r}
-AVG.spec.tob <- vector(mode = "list", length = length(step))
-
-for (iteration in 1:100){
-  spec.tob <- performance(load.tob, measure = "spec")
-
-  for(ind in 1:length(spec.tob@x.values)){
-    thresh = as.vector(slot(spec.tob, "x.values")[[ind]])
-    spec = as.vector(slot(spec.tob, "y.values")[[ind]])
-    if (length(thresh) > 2){
-      if (identical(which(thresh<=0.5), integer(0))){
-        AVG.spec.tob[[ind]][iteration] = NA
-      }
-      else{
-        low.ind = max(which(thresh >=0.5))
-        high.ind = min(which(thresh <=0.5))
-        #point slop formula of a line
-        slope = (spec[high.ind] - spec[low.ind]) / (thresh[high.ind] - thresh[low.ind])
-        AVG.spec.tob[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + spec[low.ind]
-      }
-    }
-    else{
-      AVG.spec.tob[[ind]][iteration] = NA
-    }
-  }
-}
-
-load.spec.tob.vec = c()
-for (avg in 1:length(AVG.spec.tob)){
-  load.spec.tob.vec[avg] = mean(AVG.spec.tob[[avg]], na.rm = TRUE)
-}
+load.spec.MMR.vec = evaluate_perf(100, length(step.load), load.MMR, "spec")
+load.spec.tob.vec = evaluate_perf(100, length(step.load), load.tob, "spec")
 ```

### 2021-07-29 Abstract.Rmd
started coding function to classify using fit_to_signatures

--- 2021-07-27 Abstract.Rmd
@@
 date: "7/21/2021"
-output: html_notebook
+output: html_document
 editor_options: 
@@
 ```{r}
-#Calculate Sensitivity
+#Calculate Sensitivity, Specificity, and Misclassification rate
 
@@
 nrow(sig4.specificity)/500
+
+mean(sig4.ffpe.class$misclassification)
 ```
@@
 ```{r}
-#Calculate Sensitivity
+#Calculate Sensitivity, specificity, and misclassification rate
 
@@
 nrow(sig6.specificity)/500
+
+mean(sig6.ffpe.class$misclassification)
 ```
@@
     #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
-    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
+    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],500)
     #Create FFPE sample vector of 100 mutations
@@
 ```
-
-# Mutational load performance characteristics
-
-## Mutational load predictions
-
-```{r}
-step = seq(50, 1000, by = 50)
-MMR.probabilities <- vector(mode = "list", length = length(step))
-MMR.truth <- vector(mode = "list", length = length(step))
-
-for (x in 1:length(step)){
-  sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2)
-  ffpe.truth <- rep("FFPE", step[x]/2)
-  sample.sig6 <- create_signature_sample_vector(cosmic.sig6, step[x]/2)
-  sig6.truth <- rep("MMR", step[x]/2)
-  test <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"),
-                                list(cosmic.sig6, ffpe.sig))
-  classification <- test %>%
-    mutate(trueBinary = ifelse(truth == "MMR",1,0))
-
-  MMR.probabilities[[x]] <- classification$MMR
-  MMR.truth[[x]] <- classification$trueBinary
-}
-```
-
-```{r}
-step = seq(50, 1000, by = 50)
-tob.probabilities <- vector(mode = "list", length = length(step))
-tob.truth <- vector(mode = "list", length = length(step))
-
-for (x in 1:length(step)){
-  sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2)
-  ffpe.truth <- rep("FFPE", step[x]/2)
-  sample.sig4 <- create_signature_sample_vector(cosmic.sig4, step[x]/2)
-  sig4.truth <- rep("tob", step[x]/2)
-  test <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"),
-                                list(cosmic.sig4, ffpe.sig))
-  classification <- test %>%
-    mutate(trueBinary = ifelse(truth == "Tobacco",1,0))
-
-  tob.probabilities[[x]] <- classification$Tobacco
-  tob.truth[[x]] <- classification$trueBinary
-}
-```
-
-```{r}
-load.MMR <- prediction(MMR.probabilities, MMR.truth)
-load.tob <- prediction(tob.probabilities, tob.truth)
-```
-
-## Sensitivity
-
-```{r}
-AVG.sens.MMR <- vector(mode = "list", length = length(step))
-
-for (iteration in 1:100){
-  sens.MMR <- performance(load.MMR, measure = "sens")
-
-  for(ind in 1:length(sens.MMR@x.values)){
-    thresh = as.vector(slot(sens.MMR, "x.values")[[ind]])
-    sens = as.vector(slot(sens.MMR, "y.values")[[ind]])
-    if (length(thresh) > 2){
-      if (identical(which(thresh<=0.5), integer(0))){
-        AVG.sens.MMR[[ind]][iteration] = NA
-      }
-      else{
-        low.ind = max(which(thresh >=0.5))
-        high.ind = min(which(thresh <=0.5))
-        #point slop formula of a line
-        slope = (sens[high.ind] - sens[low.ind]) / (thresh[high.ind] - thresh[low.ind])
-        AVG.sens.MMR[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + sens[low.ind]
-      }
-    }
-    else{
-      AVG.sens.MMR[[ind]][iteration] = NA
-    }
-  }
-}
-
-load.sens.MMR.vec = c()
-for (avg in 1:length(AVG.sens.MMR)){
-  load.sens.MMR.vec[avg] = mean(AVG.sens.MMR[[avg]], na.rm = TRUE)
-}
-```
-
-```{r}
-AVG.sens.tob <- vector(mode = "list", length = length(step))
-
-for (iteration in 1:100){
-  sens.tob <- performance(load.tob, measure = "sens")
-
-  for(ind in 1:length(sens.tob@x.values)){
-    thresh = as.vector(slot(sens.tob, "x.values")[[ind]])
-    sens = as.vector(slot(sens.tob, "y.values")[[ind]])
-    if (length(thresh) > 2){
-      if (identical(which(thresh<=0.5), integer(0))){
-        AVG.sens.tob[[ind]][iteration] = NA
-      }
-      else{
-        low.ind = max(which(thresh >=0.5))
-        high.ind = min(which(thresh <=0.5))
-        #point slop formula of a line
-        slope = (sens[high.ind] - sens[low.ind]) / (thresh[high.ind] - thresh[low.ind])
-        AVG.sens.tob[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + sens[low.ind]
-      }
-    }
-    else{
-      AVG.sens.tob[[ind]][iteration] = NA
-    }
-  }
-}
-
-load.sens.tob.vec = c()
-for (avg in 1:length(AVG.sens.tob)){
-  load.sens.tob.vec[avg] = mean(AVG.sens.tob[[avg]], na.rm = TRUE)
-}
-```
-
-```{r}
-#plot(step, load.sens.MMR.vec, ylim = c(0,1))
-sensitivity = append(load.sens.MMR.vec, load.sens.tob.vec)
-signature.sens = append(rep("MMR", length(load.sens.MMR.vec)),
-                        rep("Tobacco", length(load.sens.tob.vec)))
-dilution = append(step, step)
-sens.df <- data.frame(dilution, sensitivity, signature.sens)
-
-ggplot(data = sens.df, mapping = aes(x = dilution, 
-                                     y = sensitivity, 
-                                     color = signature.sens)) +
-  geom_point() +
-  geom_smooth() +
-  labs(title = "Specificity as a function of mutational load",
-       x = "Number of mutations in each sample",
-       y = "Specificity value") +
-  coord_cartesian(ylim = c(0,1))
-```
-
-## Specificity
-
-```{r}
-AVG.spec.MMR <- vector(mode = "list", length = length(step))
-
-for (iteration in 1:100){
-  spec.MMR <- performance(load.MMR, measure = "spec")
-
-  for(ind in 1:length(spec.MMR@x.values)){
-    thresh = as.vector(slot(spec.MMR, "x.values")[[ind]])
-    spec = as.vector(slot(spec.MMR, "y.values")[[ind]])
-    if (length(thresh) > 2){
-      if (identical(which(thresh<=0.5), integer(0))){
-        AVG.spec.MMR[[ind]][iteration] = NA
-      }
-      else{
-        low.ind = max(which(thresh >=0.5))
-        high.ind = min(which(thresh <=0.5))
-        #point slop formula of a line
-        slope = (spec[high.ind] - spec[low.ind]) / (thresh[high.ind] - thresh[low.ind])
-        AVG.spec.MMR[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + spec[low.ind]
-      }
-    }
-    else{
-      AVG.spec.MMR[[ind]][iteration] = NA
-    }
-  }
-}
-
-load.spec.MMR.vec = c()
-for (avg in 1:length(AVG.spec.MMR)){
-  load.spec.MMR.vec[avg] = mean(AVG.spec.MMR[[avg]], na.rm = TRUE)
-}
-```
-
-```{r}
-AVG.spec.tob <- vector(mode = "list", length = length(step))
-
-for (iteration in 1:100){
-  spec.tob <- performance(load.tob, measure = "spec")
-
-  for(ind in 1:length(spec.tob@x.values)){
-    thresh = as.vector(slot(spec.tob, "x.values")[[ind]])
-    spec = as.vector(slot(spec.tob, "y.values")[[ind]])
-    if (length(thresh) > 2){
-      if (identical(which(thresh<=0.5), integer(0))){
-        AVG.spec.tob[[ind]][iteration] = NA
-      }
-      else{
-        low.ind = max(which(thresh >=0.5))
-        high.ind = min(which(thresh <=0.5))
-        #point slop formula of a line
-        slope = (spec[high.ind] - spec[low.ind]) / (thresh[high.ind] - thresh[low.ind])
-        AVG.spec.tob[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + spec[low.ind]
-      }
-    }
-    else{
-      AVG.spec.tob[[ind]][iteration] = NA
-    }
-  }
-}
-
-load.spec.tob.vec = c()
-for (avg in 1:length(AVG.spec.tob)){
-  load.spec.tob.vec[avg] = mean(AVG.spec.tob[[avg]], na.rm = TRUE)
-}
-```
-
-```{r}
-#plot(step, load.spec.MMR.vec, ylim = c(0,1))
-
-specificity = append(load.spec.MMR.vec, load.spec.tob.vec)
-signature.spec = append(rep("MMR", length(load.spec.MMR.vec)),
-                        rep("Tobacco", length(load.spec.tob.vec)))
-dilution = append(step, step)
-spec.df <- data.frame(dilution, specificity, signature.spec)
-
-ggplot(data = spec.df, mapping = aes(x = dilution, 
-                                     y = specificity, 
-                                     color = signature.spec)) +
-  geom_point() +
-  geom_smooth() +
-  labs(title = "Specificity as a function of mutational load",
-       x = "Number of mutations in each sample",
-       y = "Specificity value") +
-  coord_cartesian(ylim = c(0,1))
-```

### 2021-07-29 Abstract.Rmd
Added accuracy to table with sensitivity/specificity.

--- 2021-07-29 Abstract.Rmd
@@
 spec = vector(mode = "list", length = length(signature.names))
+acc = vector(mode = "list", length = length(signature.names))
 signature = vector(mode = "list", length = length(signature.names))
@@
     pred = prediction(probabilities, truth)
-    sens[[sig.num]] = append(sens[[sig.num]], evaluate_perf(100, 5, pred, "sens"))
-    spec[[sig.num]] = append(spec[[sig.num]], evaluate_perf(100, 5, pred, "spec"))
+    acc[[sig.num]] = append(acc[[sig.num]], evaluate_perf(50, 5, pred, "acc"))
+    sens[[sig.num]] = append(sens[[sig.num]], evaluate_perf(50, 5, pred, "sens"))
+    spec[[sig.num]] = append(spec[[sig.num]], evaluate_perf(50, 5, pred, "spec"))
   }
@@
   summarize(sens = mean(sens, na.rm = TRUE),
-            spec = mean(spec, na.rm = TRUE))
+            spec = mean(spec, na.rm = TRUE),
+            acc = mean(acc, na.rm = TRUE))
 ```

### 2021-07-29 Abstract.Rmd
Final changes to supplementary table.

@@
 signature = unlist(signature)
+acc = unlist(acc)
 sens = unlist(sens)
@@
 #convert into data frame
-table = data.frame(percent.FFPE, signature, sens, spec)
+table = data.frame(percent.FFPE, signature, sens, spec, acc)
 
@@
             spec = mean(spec, na.rm = TRUE),
-            acc = mean(acc, na.rm = TRUE))
+            acc = mean(acc, na.rm = TRUE)) %>%
+  mutate(signature = factor(signature, signature.names)) %>%
+  arrange(signature)
 ```

### 2021-08-01 Abstract.Rmd
Formated values in supplementary table.
High resolution plots added.

@@
 editor_options: 
-  chunk_output_type: console
+  chunk_output_type: inline
 ---
@@
 library(ROCR)
+library(gridExtra)
 source("bayes_function.R")
@@
 MMR.ROC <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"),
-                              list(signature.matrix.list[[6]], ffpe.sig))
+                              list(signature.matrix.list[[6]], ffpe.signature))
 
@@
 tob.ROC <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"),
-                              list(signature.matrix.list[[4]], ffpe.sig))
+                              list(signature.matrix.list[[4]], ffpe.signature))
 ```
@@
 tob.perf <- performance(tob.pred, measure = "tpr", x.measure = "fpr")
+
+png("MMRtobROC.png", res = 300, width = 131.25, height = 108, units = "mm")
+
 plot(MMR.perf, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7),
@@
 abline(a=0, b= 1)
-legend("bottom", c("Signature 6 with FFPE noise", "Signature 4 with FFPE noise"),
+legend("bottomright", c("Signature 6 with FFPE noise", "Signature 4 with FFPE noise"),
        lty = 2:3)
+
+dev.off()
 ```
@@
 
-ggplot(data = sens.df, mapping = aes(x = dilution, 
+MMR.tob.sens <- ggplot(data = sens.df, mapping = aes(x = dilution, 
                                      y = sensitivity, 
@@
   coord_cartesian(xlim = c(0,100), ylim = c(0,1))
+
+ggsave("MMRtobSens.png", plot = MMR.tob.sens, dpi = 300, width = 175, height = 108, units = "mm")
 ```
@@
 
-ggplot(data = spec.df, mapping = aes(x = dilution, 
+MMR.tob.spec <- ggplot(data = spec.df, mapping = aes(x = dilution, 
                                      y = specificity, 
@@
   coord_cartesian(xlim = c(0,100), ylim = c(0,1))
+
+ggsave("MMRtobSpec.png", plot = MMR.tob.spec, dpi = 300, width = 175, height = 108, units = "mm")
+```
+
+## Reconstructed profiles of diluted MMR
+
+```{r}
+dilution.sig = signature.6.mat
+dilution.sig.name = "SBS6"
+total.mut = 2000
+dilute = c(0.1,0.3,0.5,0.7,0.9)*total.mut
+dilution.list <- vector(mode = "list", length = 5)
+i = 1
+
+for (mut in dilute){
+  #create samples with some percent of noise
+  add.sig = create_signature_sample_vector(dilution.sig, total.mut-mut)
+  sig.profile = vector_to_matrix(add.sig)
+  ffpe.sample = create_signature_sample_vector(ffpe.signature, mut)
+  ffpe.sig = vector_to_matrix(c(ffpe.sample, add.sig))
+  #classify mutations
+  dilution.df <- get_classification_df(list(add.sig, ffpe.sample), c("SBS6", "FFPE"),
+                                       list(dilution.sig, ffpe.signature))
+  #filter FFPE mutations
+  dilution.df.filtered <- dilution.df %>%
+      filter(classify != "FFPE")
+  recon.sig <- vector_to_matrix(dilution.df.filtered$mutations)
+
+  #values for performance characteristics
+  performance.df <- dilution.df %>%
+    mutate(TP = ifelse(classify == dilution.sig.name & truth == dilution.sig.name, 1,0),
+           FP = ifelse(classify == dilution.sig.name & truth == "FFPE", 1,0),
+           TN = ifelse(classify == "FFPE" & truth == "FFPE", 1,0),
+           FN = ifelse(classify == "FFPE" & truth == dilution.sig.name, 1,0))
+  TN = performance.df$TN
+  FN = performance.df$FN
+  TP = performance.df$TP
+  FP = performance.df$FP
+  acc.vec = TP + TN
+
+  #calculate performance characterisitics
+  sens.value = round(mean(TP)/(mean(TP) + mean(FN)), digits = 2)
+  spec.value = round(mean(TN)/(mean(TN) + mean(FP)), digits = 2)
+  acc.value = round(mean(acc.vec), digits = 2)
+
+  sig.plot <- plot_96_profile(sig.profile)+labs(title = "Signature 6 simulated sample", 
+                                                x = "Context")
+  ffpe.sig.plot <- plot_96_profile(ffpe.sig)+labs(title = "Simulated Sample with FFPE noise",
+                                                  x = "Context")
+  recon.plot <- plot_96_profile(recon.sig)+labs(title = "Signature 6 Reconstructed", 
+                                                x = "Context", 
+                                                caption = paste0("Sensitivity: ", sens.value,
+                                                                 " | Specificity: ",
+                                                                 spec.value, " | Accuracy: ",
+                                                                 acc.value))
+  dilution.list[[i]] <- grid.arrange(sig.plot, ffpe.sig.plot, recon.plot)
+
+  i = i+1
+}
+
+ggsave("sig6reconstruction_1.png", plot = dilution.list[[1]], dpi = 300, width = 175, height = 243, units = "mm")
+ggsave("sig6reconstruction_2.png", plot = dilution.list[[2]], dpi = 300, width = 175, height = 243, units = "mm")
+ggsave("sig6reconstruction_3.png", plot = dilution.list[[3]], dpi = 300, width = 175, height = 243, units = "mm")
+ggsave("sig6reconstruction_4.png", plot = dilution.list[[4]], dpi = 300, width = 175, height = 243, units = "mm")
+ggsave("sig6reconstruction_5.png", plot = dilution.list[[5]], dpi = 300, width = 175, height = 243, units = "mm")
 ```
@@
 #convert into data frame
-table = data.frame(percent.FFPE, signature, sens, spec, acc)
+table = data.frame(percent.FFPE, signature, round(sens, digits = 2), 
+                   round(spec, digits = 2), round(acc, digits = 2))
 

### 2021-08-19 Abstract.Rmd
Updated fig 2a and package versions

@@
 editor_options: 
-  chunk_output_type: inline
+  chunk_output_type: console
 ---
@@
 source("simulation.R")
+sessionInfo()
 ```
@@
 ```{r}
+set.seed()
 #Create Signature 4 Sample
@@
 
-plot(MMR.perf, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7),
-     main = "ROC curve of classifier", lty = 2) #ROC
+plot(MMR.perf, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7),lty = 3) #ROC
 plot(tob.perf, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7), 
-     add = TRUE, lty = 3)
+     add = TRUE, lty = 2)
 abline(a=0, b= 1)
-legend("bottomright", c("Signature 6 with FFPE noise", "Signature 4 with FFPE noise"),
+legend("bottomright", c("SBS4", "SBS6"),
        lty = 2:3)
@@
 
-# Mutational load performance characteristics
+<!-- # Mutational load performance characteristics -->
 
-## Mutational load predictions
+<!-- ## Mutational load predictions -->
 
-```{r}
-step_load = seq(50, 1000, by = 50)
-MMR.probabilities <- vector(mode = "list", length = length(step))
-MMR.truth <- vector(mode = "list", length = length(step))
+<!-- ```{r} -->
+<!-- step_load = seq(50, 1000, by = 50) -->
+<!-- MMR.probabilities <- vector(mode = "list", length = length(step)) -->
+<!-- MMR.truth <- vector(mode = "list", length = length(step)) -->
 
-for (x in 1:length(step)){
-  sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2)
-  ffpe.truth <- rep("FFPE", step[x]/2)
-  sample.sig6 <- create_signature_sample_vector(cosmic.sig6, step[x]/2)
-  sig6.truth <- rep("MMR", step[x]/2)
-  test <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"),
-                                list(cosmic.sig6, ffpe.sig))
-  classification <- test %>%
-    mutate(trueBinary = ifelse(truth == "MMR",1,0))
+<!-- for (x in 1:length(step)){ -->
+<!--   sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2) -->
+<!--   ffpe.truth <- rep("FFPE", step[x]/2) -->
+<!--   sample.sig6 <- create_signature_sample_vector(cosmic.sig6, step[x]/2) -->
+<!--   sig6.truth <- rep("MMR", step[x]/2) -->
+<!--   test <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"), -->
+<!--                                 list(cosmic.sig6, ffpe.sig)) -->
+<!--   classification <- test %>% -->
+<!--     mutate(trueBinary = ifelse(truth == "MMR",1,0)) -->
 
-  MMR.probabilities[[x]] <- classification$MMR
-  MMR.truth[[x]] <- classification$trueBinary
-}
-```
+<!--   MMR.probabilities[[x]] <- classification$MMR -->
+<!--   MMR.truth[[x]] <- classification$trueBinary -->
+<!-- } -->
+<!-- ``` -->
 
-```{r}
-tob.probabilities <- vector(mode = "list", length = length(step))
-tob.truth <- vector(mode = "list", length = length(step))
+<!-- ```{r} -->
+<!-- tob.probabilities <- vector(mode = "list", length = length(step)) -->
+<!-- tob.truth <- vector(mode = "list", length = length(step)) -->
 
-for (x in 1:length(step)){
-  sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2)
-  ffpe.truth <- rep("FFPE", step[x]/2)
-  sample.sig4 <- create_signature_sample_vector(cosmic.sig4, step[x]/2)
-  sig4.truth <- rep("tob", step[x]/2)
-  test <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"),
-                                list(cosmic.sig4, ffpe.sig))
-  classification <- test %>%
-    mutate(trueBinary = ifelse(truth == "Tobacco",1,0))
+<!-- for (x in 1:length(step)){ -->
+<!--   sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2) -->
+<!--   ffpe.truth <- rep("FFPE", step[x]/2) -->
+<!--   sample.sig4 <- create_signature_sample_vector(cosmic.sig4, step[x]/2) -->
+<!--   sig4.truth <- rep("tob", step[x]/2) -->
+<!--   test <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"), -->
+<!--                                 list(cosmic.sig4, ffpe.sig)) -->
+<!--   classification <- test %>% -->
+<!--     mutate(trueBinary = ifelse(truth == "Tobacco",1,0)) -->
 
-  tob.probabilities[[x]] <- classification$Tobacco
-  tob.truth[[x]] <- classification$trueBinary
-}
-```
+<!--   tob.probabilities[[x]] <- classification$Tobacco -->
+<!--   tob.truth[[x]] <- classification$trueBinary -->
+<!-- } -->
+<!-- ``` -->
 
-```{r}
-load.MMR <- prediction(MMR.probabilities, MMR.truth)
-load.tob <- prediction(tob.probabilities, tob.truth)
-```
+<!-- ```{r} -->
+<!-- load.MMR <- prediction(MMR.probabilities, MMR.truth) -->
+<!-- load.tob <- prediction(tob.probabilities, tob.truth) -->
+<!-- ``` -->
 
-## Sensitivity
+<!-- ## Sensitivity -->
 
-```{r}
-load.sens.MMR.vec = evaluate_perf(100, length(step_load), load.MMR, "sens")
-load.sens.tob.vec = evaluate_perf(100, length(step_load), load.tob, "sens")
-```
+<!-- ```{r} -->
+<!-- load.sens.MMR.vec = evaluate_perf(100, length(step_load), load.MMR, "sens") -->
+<!-- load.sens.tob.vec = evaluate_perf(100, length(step_load), load.tob, "sens") -->
+<!-- ``` -->
 
-```{r}
-#plot(step, load.sens.MMR.vec, ylim = c(0,1))
-sensitivity = append(load.sens.MMR.vec, load.sens.tob.vec)
-signature.sens = append(rep("MMR", length(load.sens.MMR.vec)),
-                        rep("Tobacco", length(load.sens.tob.vec)))
-dilution = append(step, step)
-sens.df <- data.frame(dilution, sensitivity, signature.sens)
+<!-- ```{r} -->
+<!-- #plot(step, load.sens.MMR.vec, ylim = c(0,1)) -->
+<!-- sensitivity = append(load.sens.MMR.vec, load.sens.tob.vec) -->
+<!-- signature.sens = append(rep("MMR", length(load.sens.MMR.vec)), -->
+<!--                         rep("Tobacco", length(load.sens.tob.vec))) -->
+<!-- dilution = append(step, step) -->
+<!-- sens.df <- data.frame(dilution, sensitivity, signature.sens) -->
 
-ggplot(data = sens.df, mapping = aes(x = dilution, 
-                                     y = sensitivity, 
-                                     color = signature.sens)) +
-  geom_point() +
-  geom_smooth() +
-  labs(title = "Sensitivity as a function of mutational load",
-       x = "Number of mutations in each sample",
-       y = "Sensitivity value",
-       color = "True signal") +
-  coord_cartesian(ylim = c(0,1))
-```
+<!-- ggplot(data = sens.df, mapping = aes(x = dilution,  -->
+<!--                                      y = sensitivity,  -->
+<!--                                      color = signature.sens)) + -->
+<!--   geom_point() + -->
+<!--   geom_smooth() + -->
+<!--   labs(title = "Sensitivity as a function of mutational load", -->
+<!--        x = "Number of mutations in each sample", -->
+<!--        y = "Sensitivity value", -->
+<!--        color = "True signal") + -->
+<!--   coord_cartesian(ylim = c(0,1)) -->
+<!-- ``` -->
 
-## Specificity
+<!-- ## Specificity -->
 
-```{r}
-load.spec.MMR.vec = evaluate_perf(100, length(step.load), load.MMR, "spec")
-load.spec.tob.vec = evaluate_perf(100, length(step.load), load.tob, "spec")
-```
+<!-- ```{r} -->
+<!-- load.spec.MMR.vec = evaluate_perf(100, length(step.load), load.MMR, "spec") -->
+<!-- load.spec.tob.vec = evaluate_perf(100, length(step.load), load.tob, "spec") -->
+<!-- ``` -->
 
-```{r}
-#plot(step, load.spec.MMR.vec, ylim = c(0,1))
+<!-- ```{r} -->
+<!-- #plot(step, load.spec.MMR.vec, ylim = c(0,1)) -->
 
-specificity = append(load.spec.MMR.vec, load.spec.tob.vec)
-signature.spec = append(rep("MMR", length(load.spec.MMR.vec)),
-                        rep("Tobacco", length(load.spec.tob.vec)))
-dilution = append(step, step)
-spec.df <- data.frame(dilution, specificity, signature.spec)
+<!-- specificity = append(load.spec.MMR.vec, load.spec.tob.vec) -->
+<!-- signature.spec = append(rep("MMR", length(load.spec.MMR.vec)), -->
+<!--                         rep("Tobacco", length(load.spec.tob.vec))) -->
+<!-- dilution = append(step, step) -->
+<!-- spec.df <- data.frame(dilution, specificity, signature.spec) -->
 
-ggplot(data = spec.df, mapping = aes(x = dilution, 
-                                     y = specificity, 
-                                     color = signature.spec)) +
-  geom_point() +
-  geom_smooth() +
-  labs(title = "Specificity as a function of mutational load",
-       x = "Number of mutations in each sample",
-       y = "Specificity value",
-       color = "True signal") +
-  coord_cartesian(ylim = c(0,1))
-```
+<!-- ggplot(data = spec.df, mapping = aes(x = dilution,  -->
+<!--                                      y = specificity,  -->
+<!--                                      color = signature.spec)) + -->
+<!--   geom_point() + -->
+<!--   geom_smooth() + -->
+<!--   labs(title = "Specificity as a function of mutational load", -->
+<!--        x = "Number of mutations in each sample", -->
+<!--        y = "Specificity value", -->
+<!--        color = "True signal") + -->
+<!--   coord_cartesian(ylim = c(0,1)) -->
+<!-- ``` -->

### 2021-08-19 Abstract.Rmd
Changes in figures and information required for manuscript

@@
 
-#Diagram of Methods
+# Diagram of Methods
 
@@
 
-#Diagram of Simulating Data
+# Diagram of Simulating Data
 ```{r}
@@
 
-#Reconstructed Profiles of Signatures 4 and 6
+# Reconstructed Profiles of Signatures 4 and 6
 
@@
 MMR.pred <- prediction(MMR.ROC$MMR, MMR.ROC$truthMMR)
+mmr.auc <-  performance(MMR.pred, measure = "auc")
+mmr.auc@y.values[[1]][1] #AUC for SBS6
 MMR.perf <- performance(MMR.pred, measure = "tpr", x.measure = "fpr")
+
 tob.pred <- prediction(tob.ROC$Tobacco, tob.ROC$truthTob)
+tob.auc <-  performance(tob.pred, measure = "auc")
+tob.auc@y.values[[1]][1] #AUC for SBS4
+
 tob.perf <- performance(tob.pred, measure = "tpr", x.measure = "fpr")
@@
 
-#Scatterplot of Average AUC vs. Cosine Similarity to FFPE
+# Scatterplot of Average AUC vs. Cosine Similarity to FFPE
 ```{r}
+set.seed(500)
 #outer for loop: loop through each COSMIC v3 signature
@@
 ```{r}
+
+ggplot(cos.sim.auc.data, aes(x=average_auc)) +
+  geom_histogram(color="black", fill="lightblue")+
+  labs( x = "Average AUC", y = "Count")
+ggsave("supp.fig.5.png", dpi = 300, width = 175, height = 108, units = "mm")
+
+median(cos.sim.auc.data$average_auc)
+IQR(cos.sim.auc.data$average_auc)
+
 model2 <- lm(average_auc ~ cosine_similarity, data = cos.sim.auc.data)
 
-#summary(model2)
-
-model2_summary <- summary(model2)
+summary(model2)
 
@@
 cos.sim.auc.outliers <- cos.sim.auc.data %>%
-  filter(residuals >= quantile(cos.sim.auc.data$residuals, 0.90) | 
-         residuals <= quantile(cos.sim.auc.data$residuals, 0.10) |
-           signature %in% c("SBS4"))
+  filter(average_auc<0.8)
+
+cos.sim.auc.outliers
+
+cos.sim.auc.outliers2 <- cos.sim.auc.data %>%
+  filter(average_auc>=0.8 & average_auc<0.9)
+
+cos.sim.sbs4 <- cos.sim.auc.data %>%
+  filter(signature=="SBS4")
+
 ```
@@
   geom_point()+
-  geom_smooth(method = "lm")+
-  labs(title = "Average area under curve vs. cosine similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average area under curve (10 iterations)")+
+  geom_smooth(method = "lm", se=FALSE)+
+  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average AUC")+
   geom_point(data = cos.sim.auc.outliers, color = "red")+
-  geom_label_repel(data = cos.sim.auc.outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)
+  geom_point(data = cos.sim.auc.outliers2, color = "purple")+
+  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
+  geom_hline(yintercept=0.9, linetype="dashed", color = "purple")+
+  geom_label_repel(data = cos.sim.auc.outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
+  geom_label_repel(data = cos.sim.auc.outliers2, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
+  geom_label_repel(data = cos.sim.sbs4, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
+
+ggsave("Fig2b.png", dpi = 300, width = 175, height = 108, units = "mm")
 
@@
 
-#Scatterplot of Average Sensitivity and Specificity vs. Cosine Similarity to FFPE
+# Scatterplot of Average Sensitivity and Specificity vs. Cosine Similarity to FFPE
 
 ```{r}
+set.seed(500)
 #outer for loop: loop through each COSMIC v3 signature
@@
 
-sens.spec.data.2 <- sens.spec.data %>%
-  pivot_longer(c(sensitivity, specificity), names_to = "measure", values_to = "value")
+ggplot(sens.spec.data, aes(x=sensitivity)) +
+  geom_histogram(color="black", fill="lightblue")+
+  labs( x = "Average sensitivity", y = "Count")
+ggsave("supp.fig.3.png", dpi = 300, width = 175, height = 108, units = "mm")
+
+median(sens.spec.data$sensitivity)
+IQR(sens.spec.data$sensitivity)
+
+outliers <- sens.spec.data %>%
+  filter(sensitivity<0.8 | signature=="SBS4" | signature=="SBS6")
+
+ref.signatures <- sens.spec.data %>%
+  filter(signature=="SBS4" )
+
+ggplot(sens.spec.data, aes(x = cosine_similarity, y = sensitivity))+
+  geom_point()+
+  geom_smooth(method = "lm", se=FALSE)+
+  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average sensitivity")+
+  geom_point(data = outliers, color = "red")+
+  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
+  geom_label_repel(data = outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
+
+ggsave("supp.fig6.png", dpi = 300, width = 175, height = 108, units = "mm")
+
+model.sens <- lm(sensitivity ~ cosine_similarity, data = sens.spec.data)
+
+summary(model.sens)
+
+ggplot(sens.spec.data, aes(x=specificity)) +
+  geom_histogram(color="black", fill="lightblue")+
+  labs( x = "Average specificity", y = "Count")
+ggsave("supp.fig.4.png", dpi = 300, width = 175, height = 108, units = "mm")
+
+median(sens.spec.data$specificity)
+IQR(sens.spec.data$specificity)
+
+outliers <- sens.spec.data %>%
+  filter(specificity<0.8 | signature=="SBS4" | signature=="SBS6")
+
+ggplot(sens.spec.data, aes(x = cosine_similarity, y = specificity))+
+  geom_point()+
+  geom_smooth(method = "lm", se=FALSE)+
+  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average specificity")+
+  geom_point(data = outliers, color = "red")+
+  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
+  geom_label_repel(data = outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
+
+ggsave("supp.fig7.png", dpi = 300, width = 175, height = 108, units = "mm")
+
+model.spec <- lm(specificity ~ cosine_similarity, data = sens.spec.data)
+
+summary(model.spec)
+
 ```
@@
 sensitivity = append(sens.MMR.vec, sens.tob.vec)
-signature.sens = append(rep("Signature 6", length(sens.MMR.vec)),
-                        rep("Signature 4", length(sens.tob.vec)))
+signature.sens = append(rep("SBS6", length(sens.MMR.vec)),
+                        rep("SBS4", length(sens.tob.vec)))
 dilution = append(step/10, step/10)
@@
   geom_point() +
-  geom_smooth() +
-  labs(title = "Sensitivity as a function of FFPE noise",
-       x = "Percent of FFPE noise in samples",
-       y = "Sensitivity value",
-       color = "True signal") +
-  coord_cartesian(xlim = c(0,100), ylim = c(0,1))
+  geom_smooth(se=FALSE) +
+  labs(x = "Percent of FFPE signature",
+       y = "Sensitivity",
+       color = "Baseline signature") +
+  coord_cartesian(xlim = c(0,100), ylim = c(0,1))+
+  scale_y_continuous(breaks=seq(from=0, to=1, by=.2))+
+  geom_hline(yintercept=0.8, linetype="dashed", color = "red")
 
-ggsave("MMRtobSens.png", plot = MMR.tob.sens, dpi = 300, width = 175, height = 108, units = "mm")
+ggsave("fig3a.png", plot = MMR.tob.sens, dpi = 300, width = 175, height = 108, units = "mm")
+
+MMR.tob.sens
+
 ```
@@
 specificity = append(spec.MMR.vec, spec.tob.vec)
-signature.spec = append(rep("Signature 6", length(sens.MMR.vec)),
-                        rep("Signature 4", length(spec.tob.vec)))
+signature.spec = append(rep("SBS6", length(sens.MMR.vec)),
+                        rep("SBS4", length(spec.tob.vec)))
 dilution = append(step/10, step/10)
@@
   geom_smooth() +
-  labs(title = "Specificity as a function of FFPE noise",
-       x = "Percent of FFPE noise in samples",
-       y = "Specificity value",
-       color = "True signal") +
-  coord_cartesian(xlim = c(0,100), ylim = c(0,1))
+  labs(x = "Percent of FFPE signature",
+       y = "Specificity",
+       color = "Baseline signature") +
+  coord_cartesian(xlim = c(0,100), ylim = c(0,1))+
+    scale_y_continuous(breaks=seq(from=0, to=1, by=.2))+
+  geom_hline(yintercept=0.8, linetype="dashed", color = "red")
 
-ggsave("MMRtobSpec.png", plot = MMR.tob.spec, dpi = 300, width = 175, height = 108, units = "mm")
+ggsave("fig3b.png", plot = MMR.tob.spec, dpi = 300, width = 175, height = 108, units = "mm")
+
+MMR.tob.spec
 ```

### 2021-08-19 Abstract.Rmd
Adding set seed for reproducibility

@@
 ```{r}
-set.seed()
+set.seed(500)
 #Create Signature 4 Sample

### 2021-08-19 Abstract.Rmd
small changes to generate markdown

@@
 
-summary(model2)
+model2_summary <- summary(model2)
 
@@
 
-```
-
-```{r}
-ggplot(data = sens.spec.data.2, aes(x = cosine_similarity, y = value, color = measure))+
-  geom_point()+
-  geom_smooth(method = lm, se = FALSE)+
-  labs(title = "Sensitivity & Specificity vs. Cosine Similarity to FFPE Signature", x = "Cosine Similarity of COSMIC Signature to FFPE", y = "Average Sensitivity and Specificity (10 iterations)")
 ```

### 2022-02-01 Abstract.Rmd
Evaluate max/min sensitivity, specificity, and AUC among diluted samples at which classifier fails for a fixed optimal sensitivity/specificity/AUC threshold and for all COSMIC signatures. Each evaluation is plotted and includes a linear model with distributions of optimal sensitivity/specificity/AUC.

@@
 date: "7/21/2021"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 editor_options:
@@
 
-```{r}
+```{r, include = FALSE}
 vector_to_matrix <- function(mutations.vector){
@@
 
-```{r}
+```{r, include=FALSE}
 #load FFPE signature as a matrix
@@
 
-```{r}
+```{r, include=FALSE}
 mmr.sample <- create_signature_sample_vector(signature.6.mat, 200)
@@
 
-ggsave("mutprofile.png", plot = mut.profile , width = 175, height = 108, units = "mm")
+#ggsave("mutprofile.png", plot = mut.profile , width = 175, height = 108, units = "mm")
 ```
 
-```{r}
+```{r, include=FALSE}
 sig.mat <- cbind(signature.4.mat, signature.6.mat, ffpe.signature)
@@
 
-ggsave("contributions.png", plot = contribution.plot , width = 175, height = 54, units = "mm")
+#ggsave("contributions.png", plot = contribution.plot , width = 175, height = 54, units = "mm")
 ```
 
-```{r}
+```{r, include=FALSE}
 signatures.plot <- plot_96_profile(sig.mat)+
@@
 
-ggsave("signatures.png", plot = signatures.plot , width = 175, height = 108, units = "mm")
+#ggsave("signatures.png", plot = signatures.plot , width = 175, height = 108, units = "mm")
 ```
@@
 # Diagram of Simulating Data
-```{r}
+```{r, include=FALSE}
 signature4.plot <- plot_96_profile(signature.4.mat)+
@@
 
-ggsave("signature4.png", plot = signature4.plot , width = 175, height = 81, units = "mm")
+#ggsave("signature4.png", plot = signature4.plot , width = 175, height = 81, units = "mm")
 ```
 
-```{r}
+```{r, include=FALSE}
 ffpe.plot <- plot_96_profile(ffpe.signature)+
@@
 
-ggsave("ffpe.png", plot = ffpe.plot , width = 175, height = 81, units = "mm")
+#ggsave("ffpe.png", plot = ffpe.plot , width = 175, height = 81, units = "mm")
 ```
 
-```{r}
+```{r, include=FALSE}
 sample.4 <- create_signature_sample_vector(signature.4.mat, 500)
@@
 
-ggsave("sampleprofile.png", plot = sampleprof.plot , width = 175, height = 108, units = "mm")
+#ggsave("sampleprofile.png", plot = sampleprof.plot , width = 175, height = 108, units = "mm")
 ```
@@
 
-```{r}
+```{r, include=FALSE}
 set.seed(500)
@@
 
-```{r}
+```{r, include=FALSE}
 #Add FFPE Noise
@@
 
-```{r}
+```{r, include=FALSE}
 #Classify and Filter
@@
 
-```{r}
+```{r, include=FALSE}
 #Calculate Sensitivity, Specificity, and Misclassification rate
@@
 
-```{r}
+```{r, include=FALSE}
 #Arrange 3 plots
@@
 
-ggsave("sig4reconstruction.png", plot = sig4.reconstruction , width = 175, height = 243, units = "mm")
+#ggsave("sig4reconstruction.png", plot = sig4.reconstruction , width = 175, height = 243, units = "mm")
 ```
 
-```{r}
+```{r, include=FALSE}
 #Create Signature 6 Sample
@@
 
-```{r}
+```{r, include=FALSE}
 #Calculate Sensitivity, specificity, and misclassification rate
@@
 
-ggsave("sig6reconstruction.png", plot = sig6.reconstruction , width = 175, height = 243, units = "mm")
+#ggsave("sig6reconstruction.png", plot = sig6.reconstruction , width = 175, height = 243, units = "mm")
 ```
@@
 
-```{r}
+```{r, include=FALSE}
 sample.sig6 <- create_signature_sample_vector(signature.6.mat, 500)
@@
 
-```{r}
+```{r, include=FALSE}
 MMR.ROC <- MMR.ROC %>%
@@
 # Scatterplot of Average AUC vs. Cosine Similarity to FFPE
-```{r}
+```{r, include=FALSE}
 set.seed(500)
@@
 
-```{r}
+```{r, include=FALSE}
 cos.sim.auc.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.auc.values)
@@
 
-```{r}
+```{r, include=FALSE}
 
@@
   labs( x = "Average AUC", y = "Count")
-ggsave("supp.fig.5.png", dpi = 300, width = 175, height = 108, units = "mm")
+#ggsave("supp.fig.5.png", dpi = 300, width = 175, height = 108, units = "mm")
 
@@
 
-```{r}
+```{r, include=FALSE}
 ggplot(cos.sim.auc.data, aes(x = cosine_similarity, y = average_auc))+
@@
   geom_label_repel(data = cos.sim.auc.outliers2, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
-  geom_label_repel(data = cos.sim.sbs4, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
+  geom_label_repel(data = cos.sim.sbs4, aes(label = signature), max.overlaps = 10, label.size =   0.1)
 
-ggsave("Fig2b.png", dpi = 300, width = 175, height = 108, units = "mm")
+#ggsave("Fig2b.png", dpi = 300, width = 175, height = 108, units = "mm")
 
@@
 
-```{r}
+```{r, include=FALSE}
 sens.spec.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.sensitivity.values, average.specificity.values)
@@
   labs( x = "Average sensitivity", y = "Count")
-ggsave("supp.fig.3.png", dpi = 300, width = 175, height = 108, units = "mm")
+#ggsave("supp.fig.3.png", dpi = 300, width = 175, height = 108, units = "mm")
 
@@
   geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
-  geom_label_repel(data = outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
+  geom_label_repel(data = outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)
 
-ggsave("supp.fig6.png", dpi = 300, width = 175, height = 108, units = "mm")
+#ggsave("supp.fig6.png", dpi = 300, width = 175, height = 108, units = "mm")
 
@@
   labs( x = "Average specificity", y = "Count")
-ggsave("supp.fig.4.png", dpi = 300, width = 175, height = 108, units = "mm")
+#ggsave("supp.fig.4.png", dpi = 300, width = 175, height = 108, units = "mm")
 
@@
   geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
-  geom_label_repel(data = outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
+  geom_label_repel(data = outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)
 
-ggsave("supp.fig7.png", dpi = 300, width = 175, height = 108, units = "mm")
+#ggsave("supp.fig7.png", dpi = 300, width = 175, height = 108, units = "mm")
 
@@
 num.mut = 1000
-step = seq(50, 950, by = 50)
+step = seq(10, 990, by = 10)
 ```
@@
 sens.MMR.vec = evaluate_perf(100, length(step), test.MMR, "sens")
-```
-
-```{r}
 sens.tob.vec = evaluate_perf(100, length(step), test.tob, "sens")
@@
                                      y = sensitivity, 
-                                     color = signature.sens)) +
-  geom_point() +
+                                     linetype = signature.sens)) +
+  geom_point(aes(shape = signature.sens)) +
   geom_smooth(se=FALSE) +
@@
        y = "Sensitivity",
-       color = "Baseline signature") +
+       linetype = "Baseline signature",
+       shape = "Baseline signature") +
   coord_cartesian(xlim = c(0,100), ylim = c(0,1))+
   scale_y_continuous(breaks=seq(from=0, to=1, by=.2))+
-  geom_hline(yintercept=0.8, linetype="dashed", color = "red")
+  geom_hline(yintercept=0.8, linetype=4, color = "red")
 
-ggsave("fig3a.png", plot = MMR.tob.sens, dpi = 300, width = 175, height = 108, units = "mm")
+#ggsave("fig3a.png", plot = MMR.tob.sens, dpi = 300, width = 175, height = 108, units = "mm")
 
@@
 
+```
+
+### Sensitivity threshold for all signatures
+
+```{r}
+#set threshold for a decent sensitivity for our classifier
+optimal.sens = 0.8
+max.sens.noise = c()
+#loop through each signature
+for (sig in 1:length(signature.matrix.list)){
+  #empty lists to hold classifier outputs and ground truths
+  probabilities.sig <- vector(mode = "list", length = length(step))
+  truth.sig <- vector(mode = "list", length = length(step))
+
+  #dilute each signature with some percent of FFPE
+  for (x in 1:length(step)){
+    sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x])
+    ffpe.truth <- rep("FFPE", step[x])
+    sample.sig <- create_signature_sample_vector(signature.matrix.list[[sig]],
+                                                 num.mut-step[x])
+    sig.truth <- rep("Biological", num.mut-step[x])
+    #run classifier
+    test <- get_classification_df(list(sample.sig, sample.ffpe), c("Biological","FFPE"),
+                                  list(signature.matrix.list[[sig]], ffpe.sig))
+    classification <- test %>%
+      mutate(trueBinary = ifelse(truth == "Biological",1,0))
+    #extract classifications and ground truth
+    probabilities.sig[[x]] <- classification$Biological
+    truth.sig[[x]] <- classification$trueBinary
+  }
+
+  #create ROC prediction object and evaluate
+  test.sig <- prediction(probabilities.sig, truth.sig)
+  sens.sig.vec = evaluate_perf(100, length(step), test.sig, "sens")
+
+  #cut off noise past our threshold
+  if(identical(which(sens.sig.vec <= optimal.sens), integer(0))){
+    max.sens.noise[sig] = 100
+    }
+  else{
+    max.sens.noise[sig] = min(which(sens.sig.vec <= optimal.sens))
+    }
+}
+```
+
+```{r}
+#visualize results
+
+all.sens.df <- data.frame(signature.names, max.sens.noise, ffpe.cosmic.cos.sims)
+cos.sim.sbs4.sbs6 <- all.sens.df %>%
+  filter(signature.names %in% c("SBS4", "SBS6"))
+cos.sim.sens.outliers <- all.sens.df %>%
+  filter(max.sens.noise <= 50)
+
+all.sens <- ggplot(data = all.sens.df, mapping = aes(x = ffpe.cosmic.cos.sims,
+                                         y = max.sens.noise))+
+  geom_point()+
+  geom_smooth(method = "lm", se = FALSE)+
+  labs(x = "Cosine Similarity to FFPE",
+       y = "Maximum Percent of FFPE Noise") +
+  coord_cartesian(ylim = c(0,100), xlim = c(0,1)) +
+  geom_hline(yintercept=50, linetype="dashed", color = "red")+
+  geom_label_repel(data = cos.sim.sbs4.sbs6, aes(label = signature.names), 
+                   max.overlaps = 10,
+                   label.size =   0.1)+
+  geom_label_repel(data = cos.sim.sens.outliers, aes(label = signature.names), 
+                   max.overlaps = 10,
+                   label.size =   0.1)
+
+#ggsave("fig3c_sens.png", plot = all.sens, dpi = 300, width = 175, height = 108, units = "mm")
+all.sens
+```
+
+```{r}
+#linear relationship between cosine similarity and max FFPE noise
+model.all.sens <- lm(max.sens.noise ~ ffpe.cosmic.cos.sims, data = all.sens.df)
+summary(model.all.sens)
 ```
@@
 specificity = append(spec.MMR.vec, spec.tob.vec)
-signature.spec = append(rep("SBS6", length(sens.MMR.vec)),
+signature.spec = append(rep("SBS6", length(spec.MMR.vec)),
                         rep("SBS4", length(spec.tob.vec)))
@@
                                      y = specificity, 
-                                     color = signature.sens)) +
-  geom_point() +
+                                     lty = signature.spec)) +
+  geom_point(aes(shape = signature.spec)) +
   geom_smooth() +
@@
        y = "Specificity",
-       color = "Baseline signature") +
+       lty = "Baseline signature",
+       shape = "Baseline signature") +
   coord_cartesian(xlim = c(0,100), ylim = c(0,1))+
     scale_y_continuous(breaks=seq(from=0, to=1, by=.2))+
-  geom_hline(yintercept=0.8, linetype="dashed", color = "red")
+  geom_hline(yintercept=0.8, linetype=4, color = "red")
 
-ggsave("fig3b.png", plot = MMR.tob.spec, dpi = 300, width = 175, height = 108, units = "mm")
+#ggsave("fig3b.png", plot = MMR.tob.spec, dpi = 300, width = 175, height = 108, units = "mm")
 
@@
 
+### Specificity threshold for all signatures
+
+```{r}
+#set threshold for a decent specificity for our classifier
+optimal.spec = 0.8
+min.spec.noise = c()
+#loop through each signature
+for (sig in 1:length(signature.matrix.list)){
+  #empty lists to hold classifier outputs and ground truths
+  probabilities.sig <- vector(mode = "list", length = length(step))
+  truth.sig <- vector(mode = "list", length = length(step))
+
+  #dilute each signature with some percent of FFPE
+  for (x in 1:length(step)){
+    sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x])
+    ffpe.truth <- rep("FFPE", step[x])
+    sample.sig <- create_signature_sample_vector(signature.matrix.list[[sig]],
+                                                 num.mut-step[x])
+    sig.truth <- rep("Biological", num.mut-step[x])
+    #run classifier
+    test <- get_classification_df(list(sample.sig, sample.ffpe), c("Biological","FFPE"),
+                                  list(signature.matrix.list[[sig]], ffpe.sig))
+    classification <- test %>%
+      mutate(trueBinary = ifelse(truth == "Biological",1,0))
+    #extract classifications and ground truth
+    probabilities.sig[[x]] <- classification$Biological
+    truth.sig[[x]] <- classification$trueBinary
+  }
+
+  #create ROC prediction object and evaluate
+  test.sig <- prediction(probabilities.sig, truth.sig)
+  spec.sig.vec = evaluate_perf(100, length(step), test.sig, "spec")
+
+  #cut off noise past our threshold
+  if(identical(which(spec.sig.vec <= optimal.spec), integer(0))){
+    min.spec.noise[sig] = 1
+    }
+  else{
+    min.spec.noise[sig] = max(which(spec.sig.vec <= optimal.spec))
+    }
+}
+```
+
+```{r}
+#visualize results
+
+all.spec.df <- data.frame(signature.names, min.spec.noise, ffpe.cosmic.cos.sims)
+cos.sim.sbs4.sbs6 <- all.spec.df %>%
+  filter(signature.names == "SBS4")
+cos.sim.spec.outliers <- all.spec.df %>%
+  filter(min.spec.noise >= 50)
+
+all.spec <- ggplot(data = all.spec.df, mapping = aes(x = ffpe.cosmic.cos.sims,
+                                         y = min.spec.noise))+
+  geom_point()+
+  geom_smooth(method = "lm", se = FALSE)+
+  labs(x = "Cosine Similarity to FFPE",
+       y = "Minimum Percent of FFPE Noise") +
+  coord_cartesian(ylim = c(0,100), xlim = c(0,1)) +
+  geom_hline(yintercept=50, linetype="dashed", color = "red")+
+  geom_label_repel(data = cos.sim.sbs4.sbs6, aes(label = signature.names), 
+                   max.overlaps = 10,
+                   label.size =   0.1)+
+  geom_label_repel(data = cos.sim.spec.outliers, aes(label = signature.names), 
+                   max.overlaps = 10,
+                   label.size =   0.1)
+
+#ggsave("fig3d_spec.png", plot = all.spec, dpi = 300, width = 175, height = 108, units = "mm")
+all.spec
+```
+
+```{r}
+model.all.spec <- lm(min.spec.noise ~ ffpe.cosmic.cos.sims, data = all.spec.df)
+summary(model.all.spec)
+```
+
 ## Reconstructed profiles of diluted MMR
 
-```{r}
+```{r, include=FALSE}
 dilution.sig = signature.6.mat
@@
 
-ggsave("sig6reconstruction_1.png", plot = dilution.list[[1]], dpi = 300, width = 175, height = 243, units = "mm")
-ggsave("sig6reconstruction_2.png", plot = dilution.list[[2]], dpi = 300, width = 175, height = 243, units = "mm")
-ggsave("sig6reconstruction_3.png", plot = dilution.list[[3]], dpi = 300, width = 175, height = 243, units = "mm")
-ggsave("sig6reconstruction_4.png", plot = dilution.list[[4]], dpi = 300, width = 175, height = 243, units = "mm")
-ggsave("sig6reconstruction_5.png", plot = dilution.list[[5]], dpi = 300, width = 175, height = 243, units = "mm")
+#ggsave("sig6reconstruction_1.png", plot = dilution.list[[1]], dpi = 300, width = 175, height = 243, units = "mm")
+#ggsave("sig6reconstruction_2.png", plot = dilution.list[[2]], dpi = 300, width = 175, height = 243, units = "mm")
+#ggsave("sig6reconstruction_3.png", plot = dilution.list[[3]], dpi = 300, width = 175, height = 243, units = "mm")
+#ggsave("sig6reconstruction_4.png", plot = dilution.list[[4]], dpi = 300, width = 175, height = 243, units = "mm")
+#ggsave("sig6reconstruction_5.png", plot = dilution.list[[5]], dpi = 300, width = 175, height = 243, units = "mm")
 ```
@@
 
-```{r}
+```{r, include=FALSE}
 #columns in data frame

### 2022-02-09 Mayo Data Analysis/Mayo Data Analysis.Rmd
File to analyze Mayo data. An error is produced when running classifier function because the first argument is typically a list of length 2. This list has two character vectors of point mutations mutation where one is (usually) randomly generated from a COSMIC signature and the other is randomly generated from the FFPE signature. The idea is that the classifier will combine simulated samples into a single sample. Here, we send a list where each element is a vector of point mutations for a whole sample. The error is probably coming from the list not being two parts of a single sample.

---
title: "Mayo Data Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(NMF)
library(MutationalPatterns)
library(tidyverse)
library(ggrepel)
library(ROCR)
library(gridExtra)
library(mosaic)
library(readr)
source("bayes_function.R")
source("get_classification_df.R")
source("simulation.R")
sessionInfo()
```

```{r}
#load Mayo data
mayo.samples <- read.csv("sample.final.csv")
#load FFPE signature as a matrix
ffpe.signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))
mutations <- rownames(ffpe.signature)
#Load COSMIC signatures (version 3)
cosmic.signatures <- get_known_signatures(muttype = "snv")
mmr.sig <- cosmic.signatures[,4]
mmr.matrix <- as.matrix(mmr.sig)
rownames(mmr.matrix) <- mutations

#mayo sample names
sample.names <- unique(mayo.samples$sample)

```

```{r}
#convert context column into list
#each element is a sample, each subelement is a point mutation
mutations.list <- vector(mode = "list", length = length(sample.names))
for (i in 1:length(sample.names)){
  temp.sample.df <- mayo.samples %>%
    filter(sample == sample.names[i])
  mutations.list[[i]] = temp.sample.df$context
}
```

```{r}
#run classifier
classification.df <- get_classification_df(mutations.list, c("SBS6", "FFFPE"),
                                           list(mmr.matrix, ffpe.signature))
```

### 2022-02-09 Mayo Data Analysis/get_classification_df.R

@@ -1,0 +1,129 @@
+
+#Arguments:
+#1) list where each element is a sample generated from a certain signature and
+#   each subelement is a trinucleotide mutation type (as a string)
+#2) vector of strings with names of signatures in the same order as the elements
+#   in the list from argument 1
+#3) list where each element is the true mutational signature vector
+
+
+get_classification_df <- function(samples, sig.names, signatures) {
+  
+  #Empty vectors for names of mutation sources (signatures) and randomly generated samples
+  names <- vector(mode = "character")
+  signature.sample <- vector(mode = "character")
+  
+  for (i in 1:length(samples)) {
+    #Adds each signature name for every element/profile in the sample to empty vector
+    names <- append(names, rep(sig.names[i], length(samples[[i]])))
+    #Adds each element/profile in the sample to empty vector
+    signature.sample <- append(signature.sample, samples[[i]])
+  }
+  
+  #Puts the long generated mutations and their source into a data frame and names columns
+  df.final <- data.frame(signature.sample, names)
+  colnames(df.final) <- c("mutations", "truth")
+  
+  
+  #converts sample vector into a table of probabilities
+  tab <- table(signature.sample)/sum(table(signature.sample))
+
+  #converts table of probabilities to a data frame and defines column names
+  sample.probs.df <- data.frame(tab)
+  colnames(sample.probs.df) <- c("mutations", "frequencies")
+
+  #defines a function that joins two data frames and replaces NA with 0
+  left_join_NA <- function(x, y, by) {
+    left_join(x = x, y = y, by = by) %>% 
+      mutate_each(funs(replace(., which(is.na(.)), 0)))
+  }
+  
+  #creates a vector of mutation types
+  mutations <- rownames(signatures[[1]])
+  
+  #creates a data.frame of mutation types to join with sample probabilities
+  df <- data.frame(mutations)
+
+  #joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
+  sample.df <- left_join_NA(df, sample.probs.df, by = "mutations")
+
+  #Add pseudocount to column of probabilities 
+  sample.matrix<- as.matrix(sample.df$frequencies) + 0.0001
+  rownames(sample.matrix)= mutations
+  
+  #Computes exposure matrix using sample matrix and true signature matrix
+  nmf_res <- fit_to_signatures(sample.matrix, do.call(cbind, signatures))
+  #Add element with original signatures so that Bayes function accepts this object
+  nmf_res$signatures <-do.call(cbind, signatures)
+  
[…]

### 2022-08-02 Abstract.Rmd
Reformatting of figure 2, replacing AUC for accuracy

@@
 # Diagram of Simulating Data
+
 ```{r, include=FALSE}
@@
 
+# Accuracy vs Cosine similarity to FFPE
+
+We first create a function that mixes mutations from FFPE and another given mutational signature and calls the bayes classifier `get_classification_df`
+
+```{r}
+test_mix_mutations <- function (cosmic_idx, num_cosmic, num_ffpe) {
+
+  # Create baseline mutations
+  baseline_sig <- signature.matrix.list[[cosmic_idx]]
+  baseline_name <- signature.names[cosmic_idx]
+  baseline_mut_vec <- create_signature_sample_vector(baseline_sig, num_cosmic)
+
+  # Create FFPE mutations
+  ffpe_mut_vec <- create_signature_sample_vector(ffpe.signature, num_ffpe)
+
+  # Call excerno classification and format as a tibble
+  get_classification_df (list(baseline_mut_vec, ffpe_mut_vec), 
+                          c(baseline_name,"FFPE"), 
+                          list(baseline_sig, ffpe.signature)) %>% 
+    as_tibble()
+}
+```
+
+And we try `test_mix_mutations` and plot the results to check it is working:
+
+```{r}
+set.seed(12345)
+test_tbl <- test_mix_mutations(1,500,500)
+
+test_tbl %>%
+  select (mutations, truth, SBS1, FFPE) %>%
+  pivot_longer(c(SBS1, FFPE)) %>%
+  ggplot(aes(x=truth, y=value, fill=name)) +
+    geom_boxplot()
+
+```
+
+Now let's create a function that calculates the performance metrics (sensitivity, specificity, accuracy) on the output of `test_mix_mutations`
+
+```{r}
+calc_metrics_mix <- function (cosmic_idx, num_cosmic, num_ffpe) {
+
+  baseline_name <- signature.names[cosmic_idx]
+  test_tbl <- test_mix_mutations(cosmic_idx,num_cosmic,num_ffpe)
+
+  tp <- test_tbl %>%
+    filter(truth == baseline_name & classify == baseline_name) %>%
+    count() %>%
+    pull(n)
+
+  fp <- test_tbl %>%
+    filter(truth != baseline_name & classify == baseline_name) %>%
+    count() %>%
+    pull(n)
+
+  fn <- test_tbl %>%
+    filter(truth == baseline_name & classify == "FFPE") %>%
+    count() %>%
+    pull(n)
+
+  tn <- test_tbl %>%
+    filter(truth == "FFPE" & classify == "FFPE") %>%
+    count() %>%
+    pull(n)
+
+  tibble(
+    Sensitivity = tp / (tp +fn),
+    Specificity = tn / (tp + fn),
+    Accuracy = (tp +tn) /(tp+tn+fp+fn))
+}
+
+```
+
+And we test our method:
+
+```{r}
+set.seed(12345)
+
+calc_metrics_mix(1,500,500)
+calc_metrics_mix(4,500,500)
+calc_metrics_mix(6,500,500)
+```
+
+We create a function that calculates the average metrics over a fixed number of simulations `num_iter` and test it
+
+```{r}
+calc_metrics_mix_num <- function (cosmic_idx, num_iter, num_cosmic, num_ffpe) {
+
+  baseline_name <- signature.names[cosmic_idx]
+
+  1:num_iter %>% 
+    map_dfr(~calc_metrics_mix(cosmic_idx,num_cosmic,num_ffpe))  %>%
+    pivot_longer(1:3) %>% 
+    group_by(name) %>%
+    summarize(mean = mean(value)) %>%
+    mutate(sbs = baseline_name)
+
+}
+
+calc_metrics_mix_num(1,10,500,500)
+calc_metrics_mix_num(4,10,500,500)
+calc_metrics_mix_num(6,10,500,500)
+```
+
+Finally we are ready to generate the data over all of COSMIC mutational signatures, using 10 iterations and 500 mutations for FFPE and 500 mutations from the baseline COSMIC signature
+
+```{r}
+sbs_num = length(signature.names)
+base_num = 500
+ffpe_num = 500
+reps = 10
+
+set.seed(123456)
+sim_tbl <- 1:sbs_num %>%
+  map_dfr(~calc_metrics_mix_num(.x, reps, base_num, ffpe_num))
+```
+
+First we summarize the metrics across all the COSMIC mutational signatures
+
+```{r}
+library(kableExtra)
+
+sim_tbl <- sim_tbl %>%
+  mutate (name = factor (name, 
+                         levels = c("Sensitivity", "Specificity","Accuracy"))) 
+sim_tbl %>%
+  ggplot(aes(mean))+
+  geom_histogram(binwidth = 0.025)+
+  facet_grid(.~name) +
+  labs(x="Metric across all COSMIC signatures")
+
+ggsave("paper_figures/supp3.png", dpi = 300, width = 175, height = 75, units = "mm")
+
+summary_tbl <- sim_tbl %>%
+  group_by(name) %>%
+  summarize (median = median (mean),
+           IQR = IQR(mean)) %>%
+  mutate_if(is.numeric, format, digits=2)
+
+summary_tbl %>%
+  kbl() %>%
+  kable_styling()
+```
+
+Now we explore the relationship of the performance metrics against the similarity between the baseline signature and FFPE using a figure
+
+```{r}
+library(broom)
+sim_tbl <- sim_tbl %>%
+  mutate (ffpe_cos = rep(ffpe.cosmic.cos.sims, each=3))
+
+sim_tbl %>%
+  ggplot(aes(x=ffpe_cos, mean)) +
+  geom_point()+
+  geom_smooth(method="lm", formula = "y~x")+
+  facet_grid(~name)
+
+```
+
+And using a linear model
+
+```{r}
+lm_metric<- lm(mean ~ name:ffpe_cos, data = sim_tbl)
+
+tidy(lm_metric) %>%
+  kbl() %>%
+  kable_styling()
+
+glance(lm_metric) %>%
+  kbl() %>%
+  kable_styling()
+```
+
+We define a function to represent the mutational signatures that fall below a threshold for a particular metric.
+
+```{r}
+plot_metrics_excerno <- function (tbl, metric, threshold) {
+
+  # Subset the table to the specific metric  
+  new_tbl <- filter(tbl, name==metric) 
+  # Identify signature that fall below the threshold
+  outliers <- filter( new_tbl, mean < threshold)
+  # Extract signatures SBS4 and SBS6 as a fixe reference in the graph
+  fixed_sbs <- filter( new_tbl, sbs == "SBS4" | sbs == "SBS6")
+
+  ggplot(new_tbl, aes(x = ffpe_cos, y = mean))+
+    geom_point()+
+    geom_smooth(method = "lm", se=FALSE)+
+    geom_point(data = outliers, color = "red")+
+    geom_point(data = fixed_sbs, color = "red")+
+    geom_hline(yintercept=threshold, linetype="dashed", color = "red")+
+    geom_label_repel(data=outliers, 
+                   aes(label = sbs),
+                   max.overlaps = 10, 
+                   label.size = .1)+
+    geom_label_repel(data=fixed_sbs, 
+                   aes(label = sbs),
+                   max.overlaps = 10, 
+                   label.size =   0.1) +
+   labs(x = "Cosine similarity against FFPE signature", 
+         y = metric )
+}
+```
+
+And we plot all the metrics using a threshold of 80%
+
+```{r}
+plot_metrics_excerno(sim_tbl, "Accuracy", 0.8)
+ggsave("paper_figures/fig2b.png", dpi = 300, width = 175, height = 108, units = "mm")
+
+plot_metrics_excerno(sim_tbl, "Sensitivity", 0.8)
+ggsave("paper_figures/supp6.png", dpi = 300, width = 175, height = 108, units = "mm")
+
+plot_metrics_excerno(sim_tbl, "Specificity", 0.8)
+ggsave("paper_figures/supp7.png", dpi = 300, width = 175, height = 108, units = "mm")
+```
+
+And we obtain the linear model for each metric separately and print their information
+
+```{r}
+print_lm_coefs <- function (tbl, metric) {
+  new_tbl <-  filter (tbl, name==metric)
+  lm_metrics<- lm(mean ~ ffpe_cos, data = new_tbl)
+
+  tidy(lm_metrics) %>%
+    mutate_if(is.numeric, format, digits=2)%>%
+    kbl() %>%
+    kable_styling()
+}
+
+print_lm_fit <- function (tbl, metric) {
+  new_tbl <-  filter (tbl, name==metric)
+  lm_metrics<- lm(mean ~ ffpe_cos, data = new_tbl)
+
+  glance(lm_metrics) %>%
+    mutate_if(is.numeric, format, digits=2)%>%
+    kbl() %>%
+    kable_styling()
+}
+
+print_lm_coefs(sim_tbl, "Accuracy")
+print_lm_fit(sim_tbl, "Accuracy")
+
+print_lm_coefs(sim_tbl, "Sensitivity")
+print_lm_fit(sim_tbl, "Sensitivity")
+
+print_lm_coefs(sim_tbl, "Specificity")
+print_lm_fit(sim_tbl, "Specificity")
+
+```
+
 # Scatterplot of Average AUC vs. Cosine Similarity to FFPE
+
 ```{r, include=FALSE}
@@
     #Create FFPE sample vector of 100 mutations
-    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
-    classification.df.1 <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
+    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 500)
+    classification.df.1 <- 
+      get_classification_df(list(sample.vec, ffpe.sample.vec), 
+                            c(signature.names[x],"FFPE"), 
+                            list(signature.matrix.list[[x]], ffpe.signature))
+
     class.df.colnames <- colnames(classification.df.1)
@@
 <!-- ```{r} -->
+
 <!-- step_load = seq(50, 1000, by = 50) -->
+
 <!-- MMR.probabilities <- vector(mode = "list", length = length(step)) -->
+
 <!-- MMR.truth <- vector(mode = "list", length = length(step)) -->
@@
 <!-- for (x in 1:length(step)){ -->
+
 <!--   sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2) -->
+
 <!--   ffpe.truth <- rep("FFPE", step[x]/2) -->
+
 <!--   sample.sig6 <- create_signature_sample_vector(cosmic.sig6, step[x]/2) -->
+
 <!--   sig6.truth <- rep("MMR", step[x]/2) -->
+
 <!--   test <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"), -->
+
 <!--                                 list(cosmic.sig6, ffpe.sig)) -->
+
 <!--   classification <- test %>% -->
+
 <!--     mutate(trueBinary = ifelse(truth == "MMR",1,0)) -->
@@
 <!--   MMR.probabilities[[x]] <- classification$MMR -->
+
 <!--   MMR.truth[[x]] <- classification$trueBinary -->
+
 <!-- } -->
+
 <!-- ``` -->
@@
 <!-- ```{r} -->
+
 <!-- tob.probabilities <- vector(mode = "list", length = length(step)) -->
+
 <!-- tob.truth <- vector(mode = "list", length = length(step)) -->
@@
 <!-- for (x in 1:length(step)){ -->
+
 <!--   sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2) -->
+
 <!--   ffpe.truth <- rep("FFPE", step[x]/2) -->
+
 <!--   sample.sig4 <- create_signature_sample_vector(cosmic.sig4, step[x]/2) -->
+
 <!--   sig4.truth <- rep("tob", step[x]/2) -->
+
 <!--   test <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"), -->
+
 <!--                                 list(cosmic.sig4, ffpe.sig)) -->
+
 <!--   classification <- test %>% -->
+
 <!--     mutate(trueBinary = ifelse(truth == "Tobacco",1,0)) -->
@@
 <!--   tob.probabilities[[x]] <- classification$Tobacco -->
+
 <!--   tob.truth[[x]] <- classification$trueBinary -->
+
 <!-- } -->
+
 <!-- ``` -->
@@
 <!-- ```{r} -->
+
 <!-- load.MMR <- prediction(MMR.probabilities, MMR.truth) -->
+
 <!-- load.tob <- prediction(tob.probabilities, tob.truth) -->
+
 <!-- ``` -->
@@
 <!-- ```{r} -->
+
 <!-- load.sens.MMR.vec = evaluate_perf(100, length(step_load), load.MMR, "sens") -->
+
 <!-- load.sens.tob.vec = evaluate_perf(100, length(step_load), load.tob, "sens") -->
+
 <!-- ``` -->
@@
 <!-- ```{r} -->
+
 <!-- #plot(step, load.sens.MMR.vec, ylim = c(0,1)) -->
+
 <!-- sensitivity = append(load.sens.MMR.vec, load.sens.tob.vec) -->
+
 <!-- signature.sens = append(rep("MMR", length(load.sens.MMR.vec)), -->
+
 <!--                         rep("Tobacco", length(load.sens.tob.vec))) -->
+
 <!-- dilution = append(step, step) -->
+
 <!-- sens.df <- data.frame(dilution, sensitivity, signature.sens) -->
@@
 <!-- ggplot(data = sens.df, mapping = aes(x = dilution,  -->
+
 <!--                                      y = sensitivity,  -->
+
 <!--                                      color = signature.sens)) + -->
+
 <!--   geom_point() + -->
+
 <!--   geom_smooth() + -->
+
 <!--   labs(title = "Sensitivity as a function of mutational load", -->
+
 <!--        x = "Number of mutations in each sample", -->
+
 <!--        y = "Sensitivity value", -->
+
 <!--        color = "True signal") + -->
+
 <!--   coord_cartesian(ylim = c(0,1)) -->
+
 <!-- ``` -->
@@
 <!-- ```{r} -->
+
 <!-- load.spec.MMR.vec = evaluate_perf(100, length(step.load), load.MMR, "spec") -->
+
 <!-- load.spec.tob.vec = evaluate_perf(100, length(step.load), load.tob, "spec") -->
+
 <!-- ``` -->
@@
 <!-- ```{r} -->
+
 <!-- #plot(step, load.spec.MMR.vec, ylim = c(0,1)) -->
@@
 <!-- specificity = append(load.spec.MMR.vec, load.spec.tob.vec) -->
+
 <!-- signature.spec = append(rep("MMR", length(load.spec.MMR.vec)), -->
+
 <!--                         rep("Tobacco", length(load.spec.tob.vec))) -->
+
 <!-- dilution = append(step, step) -->
+
 <!-- spec.df <- data.frame(dilution, specificity, signature.spec) -->
@@
 <!-- ggplot(data = spec.df, mapping = aes(x = dilution,  -->
+
 <!--                                      y = specificity,  -->
+
 <!--                                      color = signature.spec)) + -->
+
 <!--   geom_point() + -->
+
 <!--   geom_smooth() + -->
+
 <!--   labs(title = "Specificity as a function of mutational load", -->
+
 <!--        x = "Number of mutations in each sample", -->
+
 <!--        y = "Specificity value", -->
+
 <!--        color = "True signal") + -->
+
 <!--   coord_cartesian(ylim = c(0,1)) -->
+
 <!-- ``` -->

### 2022-08-03 paper_analysis.Rmd
Report and figures for the paper

---
title: "Excerno analysis"
author: "Audrey Mitchell, Marco Ruiz, Soua Yang, Jaime Davila"
date: "8/3/2022"
output:
  html_document: 
    keep_md: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache = TRUE, 
                      warning=FALSE, message = FALSE)
```

# Set up

We start our analysis by loading our libraries and functions from other files

```{r}
library(NMF)
library(MutationalPatterns)
library(tidyverse)
library(ggrepel)
library(ROCR)
library(kableExtra)
library(gridExtra)
library(broom)
library(plotROC)
source("bayes_function.R")
source("get_classification_df.R")
source("simulation.R")
```

We also define the function `vector_to_matrix` which allows to switch between representations of mutational profiles, in particular to use `plot_96_profile` from `MutationalPatterns`

```{r}
vector_to_matrix <- function(mutations.vector){
  tab <- table(mutations.vector)/sum(table(mutations.vector))
  #transform table to a data frame and assign column names to mutations and frequencies
  mutations.df <- data.frame(tab)
  colnames(mutations.df) <- c("mutations", "frequencies")

  #defines a function that joins two data frames and replaces NA with 0
  left_join_NA <- function(x, y, by) {
    left_join(x = x, y = y, by = by) %>% 
      mutate_each(funs(replace(., which(is.na(.)), 0)))
  }

  #converts vector of 96 mutation types to a data frame
  mutation.types.df <- data.frame(mutations)
  #joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
  complete.mutations.df <- left_join_NA(mutation.types.df, mutations.df, by = "mutations")

  #converts data frame of mutation types and frequencies to a data frame and assigns rownames
  mutations.matrix <- as.matrix(complete.mutations.df$frequencies)
  rownames(mutations.matrix) = mutations

  return(mutations.matrix)
}
```

We start by loading the FFPE mutational signatures and the SBS mutational signatures from COSMIC:

```{r}
#load FFPE signature as a matrix
ffpe_signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))

#create vector of the 96 mutation types
mutations <- rownames(ffpe_signature)

#Load COSMIC signatures (version 3)
cosmic_signatures <- get_known_signatures(muttype = "snv")

#Create vector of the signature names
signature_names <- colnames(cosmic_signatures)
```

And we create a list with all the COSMIC mutational signatures

```{r}
#Create list of signature matrices with assigned rownames
signature_matrix_list <- vector ("list", length(signature_names))
for (i in 1:length(signature_names)){
  sig <- cosmic_signatures[,i]
  sig_mat <- as.matrix(sig)
  rownames(sig_mat) <- mutations
  colnames(sig_mat) <- signature_names[i]
  signature_matrix_list[[i]] <- sig_mat
}
```

# Initial testing

We created two simulated datasets as follows:

* We combine 500 mutations from SBS4 and 500 mutations from FFPE

* We combine 500 mutations from SBS6 and 500 mutations from FFPE

## First dataset: SBS4

Let's create and visualize our first dataset by simulating mutations from SBS4

```{r}
set.seed(500)
signature_4_mat <- as.matrix(cosmic_signatures[,"SBS4"])
rownames(signature_4_mat) <- mutations

sig4_500 <- create_signature_sample_vector(signature_4_mat, 500)
sig4_profile <- vector_to_matrix(sig4_500)

plot_96_profile(sig4_profile)+
  labs(title = "Signature 4 Simulated Sample", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))
```

Then, let's add the FFPE noise and visualize the resulting mutational profile

```{r}
addffpe_500 <- create_signature_sample_vector(ffpe_signature, 500)
sig4_ffpe <- vector_to_matrix(c(sig4_500, addffpe_500))

plot_96_profile(sig4_ffpe)+
  labs(title = "Simulated Sample with FFPE Noise", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))
```

Now, let's apply our method on the simulated dataset:

```{r}
sig4_ffpe_tbl <- get_classification_df(
  list(sig4_500, addffpe_500),
  c("SBS4", "FFPE"), 
  list(signature_4_mat, ffpe_signature)) %>%
  as_tibble()
```

And calculate the confusion matrix

```{r}
(conf_mat <- table(sig4_ffpe_tbl$truth, sig4_ffpe_tbl$classify))
```

Our method is very good at identifying mutations that originated from the FFPE process, although some mutations originated from SBS4 gets missclassified as FFPE. Let's calculate the performance metrics:

```{r}
(sig4_sensitivity <- (conf_mat[2,2] / (conf_mat[2,2]+ conf_mat[2,1])))
(sig4_specificity <- (conf_mat[1,1] / (conf_mat[1,1]+ conf_mat[1,2])))
```

## Second dataset: SBS6

Let's repeat the same experiment, this time using SBS6 as our baseline signature

```{r}
signature_6_mat <- as.matrix(cosmic_signatures[,"SBS6"])
rownames(signature_6_mat) <- mutations
sig6_500 <- create_signature_sample_vector(signature_6_mat, 500)

sig6_ffpe_tbl <- get_classification_df(
  list(sig6_500, addffpe_500),
  c("SBS6", "FFPE"),
  list(signature_6_mat, ffpe_signature))%>%
  as_tibble()
```

And let's calculate our confusion matrix and our performance characteristics

```{r}
(conf_mat <- table(sig6_ffpe_tbl$truth, sig6_ffpe_tbl$classify))
(sig6_sensitivity <- (conf_mat[2,2] / (conf_mat[2,2]+ conf_mat[2,1])))
(sig6_specificity <- (conf_mat[1,1] / (conf_mat[1,1]+ conf_mat[1,2])))
```

It seems this time our method did not perform as well. Let's create mutational profile plots for:

* The original profile (SBS6)
* The original profile (SBS6) plus the FFPE noise
* The profile obtained after filtering FFPE using `excerno`

```{r}
# Original profile
sig6_profile <- vector_to_matrix(sig6_500)
sig6_profile_plot <- plot_96_profile(sig6_profile)+
  labs(title = "Signature 6 Simulated Sample", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))
# Profile + FFPE noise
sig6_ffpe <- vector_to_matrix(c(sig6_500, addffpe_500))
sig6_ffpe_plot <- plot_96_profile(sig6_ffpe)+
  labs(title = "Simulated Sample with FFPE Noise", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))
# Profile after excerno
sig6_ffpe_filtered <- sig6_ffpe_tbl %>%
  filter(classify != "FFPE")
sig6_recon_matrix <- vector_to_matrix(sig6_ffpe_filtered$mutations)

sig6_recon_plot <- plot_96_profile(sig6_recon_matrix)+
  labs(title = "Signature 6 Reconstructed", x = "Context")+
  theme(plot.title = element_text(hjust = 0.5))

(sig6_grid <- grid.arrange(sig6_profile_plot, sig6_ffpe_plot, sig6_recon_plot))
ggsave("paper_figures/sup1.png", sig6_grid, width = 175, height = 243, units = "mm")

```

It seems filtered SBS6 dataset differs from the original signature, particularly in C>T mutations preceded by an A or a T

## ROC curve for both datasets

We would like to combine the information from both datasets (SBS4 and SBS6) in order to plot them together.

```{r}
sig4_tbl <- sig4_ffpe_tbl %>%
  select(truth, SBS4) %>%
  rename (bayes = SBS4) %>%
  mutate (dataset = "SBS4",
          truth = if_else (truth=="SBS4",1, 0))

sig6_tbl <- sig6_ffpe_tbl %>%
  select(truth, SBS6) %>%
  rename (bayes = SBS6) %>%
  mutate (dataset = "SBS6",
          truth = if_else (truth=="SBS6",1, 0))

allsim_tbl <- rbind (sig4_tbl, sig6_tbl)
```

And we generate the ROC curve for both datasets and calculate their AUC

```{r}
(roc_plot <- 
  ggplot(allsim_tbl, aes(d=truth, m=bayes, color=dataset))+
  geom_roc(cutoffs.at = c(0.5), cutoff.labels = c(0.5)) +
  labs(x="False Positive Rate",
       y="True Positive Rate",
       color="Signature"))

ggsave("paper_figures/fig2a.png", roc_plot, width = 131.25, height = 108, units = "mm")

calc_auc(roc_plot)
```

# When is `excerno` a good classifier?

We hypothesize that the performance of our classifier depends on how different the base mutational signature is from the FFPE signature.

## Initial Setup

A straightforward way of calculating the similarity across signatures the cosine similarity. So we first generate the cosine similarity of all the COSMIC signatures against the FFPE signature:

```{r}
ffpe_cosmic_cos_sims <- vector(length=length(signature_names))
for (i in 1:length(signature_names)){
  cos_sim <- cos_sim_matrix(signature_matrix_list[[i]], ffpe_signature)
  ffpe_cosmic_cos_sims[i] <- cos_sim[1,1]
}
```

Second, we will create a function that mixes mutations from FFPE and another given mutational signature and calls `excerno`

```{r}
test_mix_mutations <- function (cosmic_idx, num_cosmic, num_ffpe) {

  # Create baseline mutations
  baseline_sig <- signature_matrix_list[[cosmic_idx]]
  baseline_name <- signature_names[cosmic_idx]
  baseline_mut_vec <- create_signature_sample_vector(baseline_sig, num_cosmic)

  # Create FFPE mutations
  ffpe_mut_vec <- create_signature_sample_vector(ffpe_signature, num_ffpe)

  # Call excerno classification and format as a tibble
  get_classification_df (list(baseline_mut_vec, ffpe_mut_vec), 
                          c(baseline_name,"FFPE"), 
                          list(baseline_sig, ffpe_signature)) %>% 
    as_tibble()
}
```

And we try `test_mix_mutations` and plot the results to check it is working:

```{r}
set.seed(12345)
test_tbl <- test_mix_mutations(1,500,500)

test_tbl %>%
  select (mutations, truth, SBS1, FFPE) %>%
  pivot_longer(c(SBS1, FFPE)) %>%
  ggplot(aes(x=truth, y=value, fill=name)) +
    geom_boxplot()

```

Now let's create a function that calculates the performance metrics (sensitivity, specificity, accuracy) on the output of `test_mix_mutations`

```{r}
calc_metrics_mix <- function (cosmic_idx, num_cosmic, num_ffpe) {

  baseline_name <- signature_names[cosmic_idx]
  test_tbl <- test_mix_mutations(cosmic_idx,num_cosmic,num_ffpe)

  tp <- test_tbl %>%
    filter(truth == baseline_name & classify == baseline_name) %>%
    count() %>%
    pull(n)

  fp <- test_tbl %>%
    filter(truth != baseline_name & classify == baseline_name) %>%
    count() %>%
    pull(n)

  fn <- test_tbl %>%
    filter(truth == baseline_name & classify == "FFPE") %>%
    count() %>%
    pull(n)

  tn <- test_tbl %>%
    filter(truth == "FFPE" & classify == "FFPE") %>%
    count() %>%
    pull(n)

  tibble(
    Sensitivity = tp / (tp +fn),
    Specificity = tn / (tp + fn),
    Accuracy = (tp +tn) /(tp+tn+fp+fn))
}

```

And we test our method:

```{r}
set.seed(12345)

calc_metrics_mix(1,500,500)
calc_metrics_mix(4,500,500)
calc_metrics_mix(6,500,500)
```

We create a function that calculates the average metrics over a fixed number of simulations `num_iter` and test it

```{r}
calc_metrics_mix_num <- function (cosmic_idx, num_iter, num_cosmic, num_ffpe) {

  baseline_name <- signature_names[cosmic_idx]

  1:num_iter %>% 
    map_dfr(~calc_metrics_mix(cosmic_idx,num_cosmic,num_ffpe))  %>%
    pivot_longer(1:3) %>% 
    group_by(name) %>%
    summarize(mean = mean(value)) %>%
    mutate(sbs = baseline_name)
}

calc_metrics_mix_num(1,10,500,500)
calc_metrics_mix_num(4,10,500,500)
calc_metrics_mix_num(6,10,500,500)
```

## Simulations across all COSMIC signatures

We are ready to generate the data over all of COSMIC mutational signatures, using 10 iterations and 500 mutations for FFPE and 500 mutations from the baseline COSMIC signature

```{r}
sbs_num = length(signature_names)
base_num = 500
ffpe_num = 500
reps = 10

set.seed(123456)
sim_tbl <- 1:sbs_num %>%
  map_dfr(~calc_metrics_mix_num(.x, reps, base_num, ffpe_num))
```

First we summarize the metrics across all the COSMIC mutational signatures

```{r}
sim_tbl <- sim_tbl %>%
  mutate (name = factor (name, 
                         levels = c("Sensitivity", "Specificity","Accuracy"))) 
sim_tbl %>%
  ggplot(aes(mean))+
  geom_histogram(binwidth = 0.025)+
  facet_grid(.~name) +
  labs(x="Metric across all COSMIC signatures")

ggsave("paper_figures/supp2.png", dpi = 300, width = 175, height = 75, units = "mm")

summary_tbl <- sim_tbl %>%
  group_by(name) %>%
  summarize (median = median (mean),
           IQR = IQR(mean)) %>%
  mutate_if(is.numeric, format, digits=2)

summary_tbl %>%
  kbl() %>%
  kable_styling()
```

## Performance characteristics vs similarity against FFPE signature

Now we explore the relationship of the performance metrics against the similarity between the baseline signature and FFPE using a figure

```{r}
sim_tbl <- sim_tbl %>%
  mutate (ffpe_cos = rep(ffpe_cosmic_cos_sims, each=3))

sim_tbl %>%
  ggplot(aes(x=ffpe_cos, mean)) +
  geom_point()+
  geom_smooth(method="lm", formula = "y~x")+
  facet_grid(~name)
```

We define a function to represent the mutational signatures that fall below a threshold for a particular metric.

```{r}
plot_metrics_excerno <- function (tbl, metric, threshold) {

  # Subset the table to the specific metric  
  new_tbl <- filter(tbl, name==metric) 
  # Identify signature that fall below the threshold
  outliers <- filter( new_tbl, (mean < threshold) | sbs == "SBS4" | sbs == "SBS6")

  ggplot(new_tbl, aes(x = ffpe_cos, y = mean))+
    geom_point()+
    geom_smooth(method = "lm", se=FALSE)+
    geom_point(data = outliers, color = "red")+
    geom_hline(yintercept=threshold, linetype="dashed", color = "red")+
    geom_label_repel(data=outliers, 
                   aes(label = sbs),
                   max.overlaps = 10, 
                   label.size = .1)+
    labs(x = "Cosine similarity against FFPE signature", 
         y = metric )
}
```

And we plot all the metrics using a threshold of 80%

```{r}
plot_metrics_excerno(sim_tbl, "Accuracy", 0.8)
ggsave("paper_figures/fig2b.png", dpi = 300, width = 175, height = 108, units = "mm")

plot_metrics_excerno(sim_tbl, "Sensitivity", 0.8)
ggsave("paper_figures/supp3.png", dpi = 300, width = 175, height = 108, units = "mm")

plot_metrics_excerno(sim_tbl, "Specificity", 0.8)
ggsave("paper_figures/supp4.png", dpi = 300, width = 175, height = 108, units = "mm")
```

And we create auxiliar function for getting the coefficients of the linear model and their fit

```{r}
print_lm_coefs <- function (tbl, metric) {
  new_tbl <-  filter (tbl, name==metric)
  lm_metrics<- lm(mean ~ ffpe_cos, data = new_tbl)

  tidy(lm_metrics) %>%
    mutate_if(is.numeric, format, digits=2)%>%
    kbl() %>%
    kable_styling()
}

print_lm_fit <- function (tbl, metric) {
  new_tbl <-  filter (tbl, name==metric)
  lm_metrics<- lm(mean ~ ffpe_cos, data = new_tbl)

  glance(lm_metrics) %>%
    mutate_if(is.numeric, format, digits=2)%>%
    kbl() %>%
    kable_styling()
}
```

And print the information of the linear model for

* Accuracy

```{r}
print_lm_coefs(sim_tbl, "Accuracy")
print_lm_fit(sim_tbl, "Accuracy")
```

* Sensitivity

```{r}
print_lm_coefs(sim_tbl, "Sensitivity")
print_lm_fit(sim_tbl, "Sensitivity")
```

* Specificity
```{r}
print_lm_coefs(sim_tbl, "Specificity")
print_lm_fit(sim_tbl, "Specificity")
```

```{r}
sessionInfo()
```

### 2022-08-05 paper_analysis.Rmd
Updated figure 3 for dillution analysis

@@
 ```{r}
+scale = seq (0, 1, by=0.2)
+
 (roc_plot <- 
@@
        y="True Positive Rate",
-       color="Signature"))
+       color="Signature")+
+  scale_x_continuous(labels=scale, breaks=scale)+
+  scale_y_continuous(labels=scale, breaks=scale))
 
-ggsave("paper_figures/fig2a.png", roc_plot, width = 131.25, height = 108, units = "mm")
+ggsave("paper_figures/fig2a.png", roc_plot, width = 150, height = 90, units = "mm")
 
@@
     Sensitivity = tp / (tp +fn),
-    Specificity = tn / (tp + fn),
+    Specificity = tn / (tn + fp),
     Accuracy = (tp +tn) /(tp+tn+fp+fn))
@@
 ```{r}
-plot_metrics_excerno(sim_tbl, "Accuracy", 0.8)
+
+plot_metrics_excerno(sim_tbl, "Accuracy", 0.8) +
+  scale_x_continuous(labels=scale, breaks=scale)
 ggsave("paper_figures/fig2b.png", dpi = 300, width = 175, height = 108, units = "mm")
 
-plot_metrics_excerno(sim_tbl, "Sensitivity", 0.8)
+plot_metrics_excerno(sim_tbl, "Sensitivity", 0.8)+
+    scale_x_continuous(labels=scale, breaks=scale)
 ggsave("paper_figures/supp3.png", dpi = 300, width = 175, height = 108, units = "mm")
 
-plot_metrics_excerno(sim_tbl, "Specificity", 0.8)
+plot_metrics_excerno(sim_tbl, "Specificity", 0.8)+
+  scale_x_continuous(labels=scale, breaks=scale)
 ggsave("paper_figures/supp4.png", dpi = 300, width = 175, height = 108, units = "mm")
@@
 
+# What is the influence of FFPE percentage in the performance of `excerno`?
+
+We are interested in the performance characteristics of excerno under different percentages of FFPE signature.
+
+## An initial simulation
+
+So we will mimick our original scenario scenario of SBS4 and SBS6, however this time we will introduce varying proportions going from 5% to 100% every 5%.
+
+Let's start with SBS4
+
+```{r}
+set.seed(654321)
+pct_ffpe = seq (10,90, by=5)
+metrics_order = c("Sensitivity","Specificity", "Accuracy")
+
+sbs4_dilution_tbl <- pct_ffpe %>%
+  map_dfr(~calc_metrics_mix_num(4,10, (100-.x)*10, .x*10)) %>%
+  mutate(pct_ffpe = rep(pct_ffpe, each=3),
+         name = factor (name, levels = metrics_order))
+```
+
+And plot it
+
+```{r}
+ggplot(sbs4_dilution_tbl, 
+       aes(x=pct_ffpe, y=mean)) +
+  geom_point()+
+  geom_smooth(se=FALSE)+
+  facet_grid(name~.) + 
+  geom_hline(yintercept=0.8, linetype="dashed",color = "red")
+```
+
+Seems the metrics stay above 80%, except for specificity at 10% which slightly dips below 80%
+
+Let's do the same analysis with SBS6:
+
+```{r}
+sbs6_dilution_tbl <- pct_ffpe %>%
+  map_dfr(~calc_metrics_mix_num(6,10, (100-.x)*10, .x*10)) %>%
+  mutate(pct_ffpe = rep(pct_ffpe, each=3),
+         name = factor (name, levels = metrics_order))
+
+ggplot(sbs6_dilution_tbl, 
+       aes(x=pct_ffpe, y=mean)) +
+  geom_point()+
+  geom_smooth(se=FALSE)+
+  facet_grid(name~.) + 
+  geom_hline(yintercept=0.8, linetype="dashed",color = "red")
+```
+
+Here we see how sensitivity dips below our threshold for all ffpe percentages *above* 60% ffpe and specificity drops below 80% for all the % ffpe *below* 50%.
+Accuracy stays mostly above 80% and it is best when there is only one class of mutations (either FFPE or a single SBS)
+
+Let's combine both datasets in a single graph by using color and focus on sensitivity and specificity.
+
+```{r}
+dilution_tbl <- rbind (sbs4_dilution_tbl, sbs6_dilution_tbl)
+
+scale_x = seq (10,90, by=10)
+scale_y = seq (0.2, 1, by=0.2)
+
+dilution_tbl %>%
+  filter (name!="Accuracy") %>%
+  ggplot(aes(x=pct_ffpe, y=mean, color=sbs)) +
+    geom_point()+
+    geom_smooth(se=FALSE)+
+    facet_grid(name~.) + 
+    geom_hline(yintercept=0.8, linetype="dashed",color = "red")+
+    labs(x= "Percent of FFPE signature",
+         y="",
+         color = "Signature")+
+    scale_x_continuous(breaks = scale_x, labels = scale_x)+
+    scale_y_continuous(breaks = scale_y, labels = scale_y)
+
+ggsave("paper_figures/fig3a.png", dpi = 300, width = 150, height = 90, units = "mm")
+```
+
+Notice the trend is that for increasing percent of FFPE signature, the sensitivity decreases, while the specificity increases.
+
+## A general approach
+
+We are interested in assessing the trends in performance metrics across *all* of the signatures in COSMIC.
+In order to do that we need to create an auxiliary function `calc_dillution_metrics` that generates a tibble with this information for a particular mutational signature.
+We also create a function `plot_dillution_metrics` that plots this information
+
+```{r}
+calc_dillution_metrics <- function (ffpe_grid, cosmic_idx) {
+  metrics_order = c("Sensitivity","Specificity", "Accuracy")
+  sig_name=factor(signature_names[cosmic_idx], signature_names)
+
+  ffpe_grid %>%
+    map_dfr(~calc_metrics_mix_num(cosmic_idx,10, (100-.x)*10, .x*10)) %>%
+    mutate(pct_ffpe = rep(ffpe_grid, each=3),
+         sbs = sig_name,
+         metric = factor (name, levels = metrics_order)) %>%
+    select(-c(name))
+}
+
+plot_dillution_metrics <- function (sim_tbl) {
+  scale_x = seq (10,90, by=10)
+  scale_y = seq (0.2, 1, by=0.2)
+
+  sim_tbl %>%
+  ggplot(aes(x=pct_ffpe, y=mean, color=sbs)) +
+    geom_point()+
+    geom_smooth(se=FALSE)+
+    facet_grid(metric~.) + 
+    geom_hline(yintercept=0.8, linetype="dashed",color = "red")+
+    labs(x= "Percent of FFPE signature",
+         y="",
+         color = "Signature")+
+    scale_x_continuous(breaks = scale_x, labels = scale_x)+
+    scale_y_continuous(breaks = scale_y, labels = scale_y)
+}
+```
+
+We test our functions for SBS4 and SBS6:
+
+```{r}
+calc_dillution_metrics(pct_ffpe, 4) %>%
+  plot_dillution_metrics()
+
+calc_dillution_metrics(pct_ffpe, 6) %>%
+  plot_dillution_metrics()
+```
+
+Now we are ready to calculate the dillution metrics across all COSMIC mutational signatures (this takes 20min to run)
+
+```{r}
+set.seed(123456)
+# Full dataset run
+sim_dillution_tbl <- 1:sbs_num %>%
+  map_dfr(~calc_dillution_metrics(pct_ffpe, .x))
+```
+
+And we plot the sensitivity and specificity across all levels of FFPE contamination:
+
+```{r}
+scale_x = seq (10,90, by=10)
+scale_y = seq (0.2, 1, by=0.2)
+
+sim_dillution_tbl %>%
+  filter(metric != "Accuracy") %>%
+  ggplot (aes(x=pct_ffpe, y=mean, color=metric)) +
+    geom_point (alpha=0.3) +
+    geom_smooth(method="lm", se=FALSE)+
+    facet_grid(metric~.)+
+    geom_hline(yintercept=0.8, linetype="dashed",color = "red")+
+    scale_x_continuous(breaks = scale_x, labels = scale_x)+
+    scale_y_continuous(breaks = scale_y, labels = scale_y)+
+    labs(x= "Percent of FFPE signature",
+         y="")+
+    theme(legend.position = "none")
+
+ggsave("paper_figures/supp5.png", dpi = 300, width = 150, height = 90, units = "mm")
+```
+
 ```{r}

### 2022-08-09 paper_analysis.Rmd
Changes to the paper figure flow.

@@
   ggplot(aes(mean))+
-  geom_histogram(binwidth = 0.025)+
+  geom_histogram(binwidth = 0.025, color="black", fill="lightblue")+
   facet_grid(.~name) +
@@
 ```{r}
-plot_metrics_excerno <- function (tbl, metric, threshold) {
+plot_metrics_excerno <- function (tbl, metrics, threshold) {
+
+  fixed_sbs <- c("SBS4", "SBS6") # Fixed signatures
 
   # Subset the table to the specific metric  
-  new_tbl <- filter(tbl, name==metric) 
+  new_tbl <- filter(tbl, name %in% metrics) 
   # Identify signature that fall below the threshold
-  outliers <- filter( new_tbl, (mean < threshold) | sbs == "SBS4" | sbs == "SBS6")
+  fixed_below <- filter( new_tbl, (mean < threshold) & (sbs%in% fixed_sbs))
+  fixed_above <- filter( new_tbl, (mean > threshold) & (sbs%in% fixed_sbs))
 
@@
     geom_smooth(method = "lm", se=FALSE)+
-    geom_point(data = outliers, color = "red")+
+    geom_point(data = fixed_below, color = "red")+
+    geom_point(data = fixed_above, color = "deepskyblue")+
     geom_hline(yintercept=threshold, linetype="dashed", color = "red")+
-    geom_label_repel(data=outliers, 
-                   aes(label = sbs),
-                   max.overlaps = 10, 
-                   label.size = .1)+
+    geom_label_repel(data=fixed_below, aes(label=sbs))+
+    geom_label_repel(data=fixed_above, aes(label=sbs))+
     labs(x = "Cosine similarity against FFPE signature", 
-         y = metric )
+         y="")+
+    facet_wrap(.~name, strip.position="left")+
+    theme(strip.text.y = element_text(size = 15),
+          axis.title.x = element_text (size=15))
 }
@@
 ```{r}
+plot_metrics_excerno(sim_tbl, c("Sensitivity", "Specificity"),0.8)+
+  scale_x_continuous(labels=scale, breaks=scale)
 
-plot_metrics_excerno(sim_tbl, "Accuracy", 0.8) +
-  scale_x_continuous(labels=scale, breaks=scale)
-ggsave("paper_figures/fig2b.png", dpi = 300, width = 175, height = 108, units = "mm")
+ggsave("paper_figures/fig2b.png", dpi = 300, width = 180, height = 90, units = "mm")
 
-plot_metrics_excerno(sim_tbl, "Sensitivity", 0.8)+
-    scale_x_continuous(labels=scale, breaks=scale)
-ggsave("paper_figures/supp3.png", dpi = 300, width = 175, height = 108, units = "mm")
-
-plot_metrics_excerno(sim_tbl, "Specificity", 0.8)+
-  scale_x_continuous(labels=scale, breaks=scale)
-ggsave("paper_figures/supp4.png", dpi = 300, width = 175, height = 108, units = "mm")
 ```
@@
 
-Let's start with SBS4
+Let's start with SBS4:
 
@@
   geom_hline(yintercept=0.8, linetype="dashed",color = "red")
+
 ```
@@
 
-Let's combine both datasets in a single graph by using color and focus on sensitivity and specificity.
+Let's combine both datasets in a single graph, coloring by signature
 
@@
 dilution_tbl %>%
-  filter (name!="Accuracy") %>%
+  filter(name != "Accuracy") %>%
   ggplot(aes(x=pct_ffpe, y=mean, color=sbs)) +
     geom_point()+
-    geom_smooth(se=FALSE)+
-    facet_grid(name~.) + 
+    geom_smooth(se=FALSE,formula=y~x, method=loess)+
+    facet_wrap(.~name, strip.position="left")+
     geom_hline(yintercept=0.8, linetype="dashed",color = "red")+
@@
     scale_x_continuous(breaks = scale_x, labels = scale_x)+
-    scale_y_continuous(breaks = scale_y, labels = scale_y)
+    scale_y_continuous(breaks = scale_y, labels = scale_y)+
+    theme(strip.text.y = element_text(size = 15),
+          axis.title.x = element_text (size=15))
 
-ggsave("paper_figures/fig3a.png", dpi = 300, width = 150, height = 90, units = "mm")
+ggsave("paper_figures/fig3a.png", dpi = 300, width = 180, height = 90, units = "mm")
 ```
@@
   map_dfr(~calc_dillution_metrics(pct_ffpe, .x))
+
+rep_ffpe_cos <- rep(ffpe_cosmic_cos_sims, each=length(pct_ffpe)*3)
+
+sim_dillution_tbl <-  sim_dillution_tbl %>%
+  mutate (ffpe_cos = rep_ffpe_cos)
 ```
 
-And we plot the sensitivity and specificity across all levels of FFPE contamination:
+And we plot the sensitivity and specificity across all levels of FFPE contamination as a scatter plot:
 
@@
 sim_dillution_tbl %>%
-  filter(metric != "Accuracy") %>%
-  ggplot (aes(x=pct_ffpe, y=mean, color=metric)) +
-    geom_point (alpha=0.3) +
-    geom_smooth(method="lm", se=FALSE)+
-    facet_grid(metric~.)+
-    geom_hline(yintercept=0.8, linetype="dashed",color = "red")+
+  filter (metric!="Accuracy") %>%
+  ggplot (aes(x=pct_ffpe, y=mean)) +
+    geom_smooth(se=FALSE, method="loess", formula="y~x", size=.75)+
+    geom_point (alpha=0.2) +
+    facet_wrap(.~metric, strip.position="left")+
+    geom_hline(yintercept=0.8, linetype="dashed",color = "red", size=0.5)+
     scale_x_continuous(breaks = scale_x, labels = scale_x)+
@@
          y="")+
-    theme(legend.position = "none")
+    theme(strip.text.y = element_text(size = 15),
+          axis.title.x = element_text (size=15))
 
-ggsave("paper_figures/supp5.png", dpi = 300, width = 150, height = 90, units = "mm")
+ggsave("paper_figures/supp2.png", dpi = 300, width = 180, height = 90, units = "mm")
+```
+
+We are interested in plotting the performance characteristics across multiple dillution levels so we define a function to do that:
+
+```{r}
+plot_metrics_grid_excerno <- function (tbl, grid, fixed_sbs, metrics, threshold) {
+
+  # Subset the table to the specific metric  
+  new_tbl <- filter(tbl, metric %in% metrics & pct_ffpe %in% grid) 
+  fixed_below <- filter( new_tbl, (mean < threshold) & (sbs%in% fixed_sbs))
+  fixed_above <- filter( new_tbl, (mean > threshold) & (sbs%in% fixed_sbs))
+
+  grid.labs <- str_c(grid,"% FFPE")
+  names(grid.labs) <- grid
+
+  ggplot(new_tbl, aes(x = ffpe_cos, y = mean))+
+    geom_point()+
+    geom_smooth(method = "lm", se=FALSE, formula="y~x")+
+    geom_hline(yintercept=threshold, linetype="dashed", color = "red")+
+    geom_point(data = fixed_below, color = "red")+
+    geom_point(data = fixed_above, color = "deepskyblue")+
+    geom_label_repel(data=fixed_below, aes(label=sbs))+
+    geom_label_repel(data=fixed_above, aes(label=sbs))+
+    facet_grid(pct_ffpe~metric, labeller=labeller(pct_ffpe=grid.labs))+
+    labs(x = "Cosine similarity against FFPE signature", 
+         y = "")+
+    theme(strip.text.x = element_text(size = 15),
+          strip.text.y = element_text(size = 12),
+          axis.title.x = element_text (size=15))
+}
+```
+
+```{r}
+perf_grid = c(10,30,50,70,90)
+plot_metrics_grid_excerno(sim_dillution_tbl, perf_grid, 
+                          fixed_sbs, c("Sensitivity","Specificity"),0.8)
+ggsave("paper_figures/fig3b.png", dpi = 300, width = 150, height = 150, units = "mm")
+```
+
+From these graphs we observe that
+
+* In the sensitivity plot, the slope of the trend line decreases as the percent FFPE increases.
+* In the specificity plot, the slope of the trend line increases as the percent FFPE increases.
+
+In order to quantify and make these observations more precise we will build linear models with an interaction term between the cosine similarity against the FFPE signature and the percent FFPE.
+In order to do that we need the following functions:
+
+```{r}
+print_lm_coefs <- function (tbl, metric_name, formula) {
+  new_tbl <-  filter (tbl, metric==metric_name)
+  lm_metrics<- lm(formula, data = new_tbl)
+
+  tidy(lm_metrics) %>%
+    mutate_if(is.numeric, format, digits=1)%>%
+    kbl() %>%
+    kable_styling()
+}
+
+print_lm_fit <- function (tbl, metric_name, formula) {
+  new_tbl <-  filter (tbl, metric==metric_name)
+  lm_metrics<- lm(formula, data = new_tbl)
+
+  glance(lm_metrics) %>%
+    mutate_if(is.numeric, format, digits=2)%>%
+    kbl() %>%
+    kable_styling()
+}
+
+```
+
+Here is the information for each of the performance characteristics:
+
+```{r}
+sim_dillution_tbl <- sim_dillution_tbl %>%
+  mutate(ffpe = pct_ffpe/100)
+```
+
+* Sensitivity:
+
+```{r}
+print_lm_coefs(sim_dillution_tbl, "Sensitivity", "mean~ffpe*ffpe_cos")
+print_lm_fit(sim_dillution_tbl, "Sensitivity", "mean~ffpe*ffpe_cos")
+```
+
+* Specificity
+
+```{r}
+print_lm_coefs(sim_dillution_tbl, "Specificity", "mean~ffpe*ffpe_cos")
+print_lm_fit(sim_dillution_tbl, "Specificity", "mean~ffpe*ffpe_cos")
+```
+
+## Another way to look at it (Omit)
+
+We know that
+
+* Sensitivity is a decreasing function of percent FFPE.
+* Specificity is an increasing function of percent FFPE.
+
+We would like to summarize this characteristics more succintly and establish the relationship with the similarity against the FFPE signature. To do that we will we fix a threshold value in 80% and:
+
+* We will record the maximum ffpe where sensitivity drops below this threshold. We will call this **maximum permissible sensitivy FFPE* 
+* We will record the first percent ffpe where specificity rises above this threshold **minimum permissible specificity FFPE*
+
+```{r}
+threshold <- 0.8
+
+sens_thres_tbl <- sim_dillution_tbl %>%
+  filter (metric == "Sensitivity") %>%
+  filter(mean>threshold) %>%
+  group_by (sbs) %>%
+  slice (which.max(pct_ffpe)) %>%
+  ungroup() %>%
+  mutate(cos = ffpe_cosmic_cos_sims)
+
+spec_thres_tbl <- sim_dillution_tbl %>%
+  filter (metric == "Specificity" ) %>%
+  filter (mean > threshold) %>%
+  group_by (sbs) %>%
+  slice (which.min(pct_ffpe)) %>%
+  ungroup() %>%
+  mutate(cos = ffpe_cosmic_cos_sims)
+```
+
+Let's look at the distribution of those thresholds
+
+```{r}
+sens_thres_tbl %>%
+  summarize (median = median (mean),
+           IQR = IQR(mean)) %>%
+  mutate_if(is.numeric, format, digits=2) %>%
+  kbl() %>%
+  kable_styling()
+
+spec_thres_tbl %>%
+  summarize (median = median (mean),
+           IQR = IQR(mean)) %>%
+  mutate_if(is.numeric, format, digits=2) %>%
+  kbl() %>%
+  kable_styling()
+
+ggplot(sens_thres_tbl, aes(x=pct_ffpe))+
+  geom_histogram(binwidth=5, color="black", fill="lightblue")+
+  labs(x= "Percent of FFPE signature",
+       y="Number")
+
+ggplot(spec_thres_tbl, aes(x=pct_ffpe))+
+  geom_histogram(binwidth=5, color="black", fill="lightblue")+
+  labs(x= "Percent of FFPE signature",
+       y="Number")
+
+```
+
+And let's plot the relationship with similarity against the FFPE signature:
+
+```{r}
+sens_thres_tbl %>%
+  ggplot (aes(cos, pct_ffpe)) +
+  geom_point() + 
+  geom_smooth(method="lm", formula = "y~x")+
+  scale_x_continuous(breaks = scale_y, labels = scale_y)+
+  scale_y_continuous(breaks = scale_x, labels = scale_x)+
+  labs(x= "Cosine similarity against FFPE",
+         y="Extreme percent FFPE")
+
+spec_thres_tbl %>%
+  ggplot (aes(cos, pct_ffpe)) +
+  geom_point() + 
+  geom_smooth(method="lm", formula = "y~x")+
+  scale_x_continuous(breaks = scale_y, labels = scale_y)+
+  scale_y_continuous(breaks = scale_x, labels = scale_x)+
+  labs(x= "Cosine similarity against FFPE",
+         y="Extreme percent FFPE")
+
+```
+
+And the linear model metrics for each case:
+
+```{r}
+sens_thres_lm <- lm (pct_ffpe ~ cos, sens_thres_tbl)
+
+spec_thres_lm <- lm (pct_ffpe ~ cos, spec_thres_tbl)
+
+summary(sens_thres_lm)
+summary(spec_thres_lm)
 ```

### 2022-08-10 paper_analysis.Rmd
Changes to figures including combined effect of FFPE percent and cosine similarity. Cleanup of previous markdowns and files

@@
 
-# When is `excerno` a good classifier?
+# The effect of mutational signatures in `excerno`
 
-We hypothesize that the performance of our classifier depends on how different the base mutational signature is from the FFPE signature.
+We hypothesize that the performance of our classifier depends on how the similarity of the base mutational signature against the FFPE signature.
 
@@
 
-A straightforward way of calculating the similarity across signatures the cosine similarity. So we first generate the cosine similarity of all the COSMIC signatures against the FFPE signature:
+A straightforward way of calculating the similarity across signatures is the cosine similarity. So we first generate the cosine similarity of all the COSMIC signatures against the FFPE signature:
 
@@
 
-And we try `test_mix_mutations` and plot the results to check it is working:
+We try `test_mix_mutations` and plot the results to check it is working:
 
@@
 
-We create a function that calculates the average metrics over a fixed number of simulations `num_iter` and test it
+Finally we create a function `calc_metrics_mix_num` that calculates the average metrics across a fixed number of simulations `num_iter`
 
@@
 }
+```
 
+And we test it using our to-go signatures `SBS1`, `SBS4`, and `SBS6`
+
+```{r}
 calc_metrics_mix_num(1,10,500,500)
@@
 
-## Simulations across all COSMIC signatures
+## Testing across all COSMIC signatures
 
@@
 
-First we summarize the metrics across all the COSMIC mutational signatures
+We summarize the performance metrics across all the COSMIC mutational signatures
 
@@
   labs(x="Metric across all COSMIC signatures")
-
-ggsave("paper_figures/supp2.png", dpi = 300, width = 175, height = 75, units = "mm")
 
@@
 
-## Performance characteristics vs similarity against FFPE signature
+## Performance metrics vs similarity against FFPE signature
 
-Now we explore the relationship of the performance metrics against the similarity between the baseline signature and FFPE using a figure
+Now we explore the relationship of the performance metrics against the similarity between the baseline signature and FFPE :
 
@@
 
-We define a function to represent the mutational signatures that fall below a threshold for a particular metric.
+To a varying extent it seems there is a linear trend across all the mutational signatures.
+
+Now we define a function that polishes the previous graph and plots two signatures on it (`SBS4` and `SBS6`).
+The two fixed signatues will have different shape and color depending on whether or not they meet a threshold of 80%
 
@@
     geom_smooth(method = "lm", se=FALSE)+
-    geom_point(data = fixed_below, color = "red")+
-    geom_point(data = fixed_above, color = "deepskyblue")+
+    geom_point(data = fixed_below, fill = "red", shape=24, size=3)+
+    geom_point(data = fixed_above, fill = "blue",shape=22, size=3)+
     geom_hline(yintercept=threshold, linetype="dashed", color = "red")+
@@
 
-And we create auxiliar function for getting the coefficients of the linear model and their fit
+And we create a function for getting the coefficients of the linear model and their fit
 
@@
 
-# What is the influence of FFPE percentage in the performance of `excerno`?
+# The effect of FFPE percent in the performance of `excerno`.
 
@@
 
-So we will mimick our original scenario scenario of SBS4 and SBS6, however this time we will introduce varying proportions going from 5% to 100% every 5%.
+So we will mimic our original scenario scenario of `SBS4` and `SBS6`, however this time we will introduce varying proportions going from 10% to 90% every 5%.
 
@@
   geom_point()+
-  geom_smooth(se=FALSE)+
+  geom_smooth(se=FALSE, method="loess", formula=y~x)+
   facet_grid(name~.) + 
@@
 
-Let's combine both datasets in a single graph, coloring by signature
+Let's combine both datasets in a single graph, focusing on sensitivity and specificity and coloring by signature
 
@@
   filter(name != "Accuracy") %>%
-  ggplot(aes(x=pct_ffpe, y=mean, color=sbs)) +
+  ggplot(aes(x=pct_ffpe, y=mean, color=sbs, shape=sbs)) +
     geom_point()+
@@
          y="",
-         color = "Signature")+
+         color = "Signature",
+         shape="")+
     scale_x_continuous(breaks = scale_x, labels = scale_x)+
@@
     theme(strip.text.y = element_text(size = 15),
-          axis.title.x = element_text (size=15))
+          axis.title.x = element_text (size=15))+
+    guides(shape= "none")
 
@@
 
-## A general approach
+## Simulating across all COSMIC signatures
 
@@
 
-Now we are ready to calculate the dillution metrics across all COSMIC mutational signatures (this takes 20min to run)
+Now we are ready to calculate the dilution metrics across all COSMIC mutational signatures (this takes 20min to run)
 
@@
 set.seed(123456)
+# sbs_num <- 3
 # Full dataset run
@@
 
-rep_ffpe_cos <- rep(ffpe_cosmic_cos_sims, each=length(pct_ffpe)*3)
+rep_ffpe_cos <- rep(ffpe_cosmic_cos_sims[1:sbs_num], each=length(pct_ffpe)*3)
 
@@
     theme(strip.text.y = element_text(size = 15),
-          axis.title.x = element_text (size=15))
+          axis.title.x = element_text (size = 15))
 
@@
 
-We are interested in plotting the performance characteristics across multiple dillution levels so we define a function to do that:
+We are interested in plotting the performance characteristics across multiple dilution levels so we define a function to do that:
 
@@
     geom_hline(yintercept=threshold, linetype="dashed", color = "red")+
-    geom_point(data = fixed_below, color = "red")+
-    geom_point(data = fixed_above, color = "deepskyblue")+
+    geom_point(data = fixed_below, fill = "red", shape=24, size=3)+
+    geom_point(data = fixed_above, fill = "blue",shape=22, size=3)+
     geom_label_repel(data=fixed_below, aes(label=sbs))+
@@
 ```{r}
+
 perf_grid = c(10,30,50,70,90)
+fixed_sbs = c("SBS4", "SBS6")
 plot_metrics_grid_excerno(sim_dillution_tbl, perf_grid, 
@@
 
-Here is the information for each of the performance characteristics:
+Finally we make sure that the ffpe content is on the decimal scale (to be consistent) with the cosine similarity, which is a number from 0 to 1.
 
@@
 ```
+
+Here is the information of the models for sensitivity and specificity
 
@@
 
-## Another way to look at it (Omit)
+# Conclusion
 
-We know that
+We tested our method across varying levels of FFPE content and across different mutational signatures.
+We found that the performance characteristics of our method decrease as the cosine similarity against the FFPE signature increases.
+We also establish that FFPE content increases, our sensitivity decreases and our specificity increases and establish linear models with good fit.
 
-* Sensitivity is a decreasing function of percent FFPE.
-* Specificity is an increasing function of percent FFPE.
-
-We would like to summarize this characteristics more succintly and establish the relationship with the similarity against the FFPE signature. To do that we will we fix a threshold value in 80% and:
-
-* We will record the maximum ffpe where sensitivity drops below this threshold. We will call this **maximum permissible sensitivy FFPE* 
-* We will record the first percent ffpe where specificity rises above this threshold **minimum permissible specificity FFPE*
-
-```{r}
-threshold <- 0.8
-
-sens_thres_tbl <- sim_dillution_tbl %>%
-  filter (metric == "Sensitivity") %>%
-  filter(mean>threshold) %>%
-  group_by (sbs) %>%
-  slice (which.max(pct_ffpe)) %>%
-  ungroup() %>%
-  mutate(cos = ffpe_cosmic_cos_sims)
-
-spec_thres_tbl <- sim_dillution_tbl %>%
-  filter (metric == "Specificity" ) %>%
-  filter (mean > threshold) %>%
-  group_by (sbs) %>%
-  slice (which.min(pct_ffpe)) %>%
-  ungroup() %>%
-  mutate(cos = ffpe_cosmic_cos_sims)
-```
-
-Let's look at the distribution of those thresholds
-
-```{r}
-sens_thres_tbl %>%
-  summarize (median = median (mean),
-           IQR = IQR(mean)) %>%
-  mutate_if(is.numeric, format, digits=2) %>%
-  kbl() %>%
-  kable_styling()
-
-spec_thres_tbl %>%
-  summarize (median = median (mean),
-           IQR = IQR(mean)) %>%
-  mutate_if(is.numeric, format, digits=2) %>%
-  kbl() %>%
-  kable_styling()
-
-ggplot(sens_thres_tbl, aes(x=pct_ffpe))+
-  geom_histogram(binwidth=5, color="black", fill="lightblue")+
-  labs(x= "Percent of FFPE signature",
-       y="Number")
-
-ggplot(spec_thres_tbl, aes(x=pct_ffpe))+
-  geom_histogram(binwidth=5, color="black", fill="lightblue")+
-  labs(x= "Percent of FFPE signature",
-       y="Number")
-
-```
-
-And let's plot the relationship with similarity against the FFPE signature:
-
-```{r}
-sens_thres_tbl %>%
-  ggplot (aes(cos, pct_ffpe)) +
-  geom_point() + 
-  geom_smooth(method="lm", formula = "y~x")+
-  scale_x_continuous(breaks = scale_y, labels = scale_y)+
-  scale_y_continuous(breaks = scale_x, labels = scale_x)+
-  labs(x= "Cosine similarity against FFPE",
-         y="Extreme percent FFPE")
-
-spec_thres_tbl %>%
-  ggplot (aes(cos, pct_ffpe)) +
-  geom_point() + 
-  geom_smooth(method="lm", formula = "y~x")+
-  scale_x_continuous(breaks = scale_y, labels = scale_y)+
-  scale_y_continuous(breaks = scale_x, labels = scale_x)+
-  labs(x= "Cosine similarity against FFPE",
-         y="Extreme percent FFPE")
-
-```
-
-And the linear model metrics for each case:
-
-```{r}
-sens_thres_lm <- lm (pct_ffpe ~ cos, sens_thres_tbl)
-
-spec_thres_lm <- lm (pct_ffpe ~ cos, spec_thres_tbl)
-
-summary(sens_thres_lm)
-summary(spec_thres_lm)
-```
+This analysis was done using the following environment:
 

### 2022-08-10 sandbox/Abstract.Rmd
Changes to figures including combined effect of FFPE percent and cosine similarity. Cleanup of previous markdowns and files

--- 2022-08-02 Abstract.Rmd
@@
 
-# Scatterplot of Average AUC vs. Cosine Similarity to FFPE
-
-```{r, include=FALSE}
-set.seed(500)
-#outer for loop: loop through each COSMIC v3 signature
-average.auc.values <- c()
-for (x in 1:length(signature.names)){
-  signature.auc.vec <- c()
-  #inner for loop: calculate misclassification rate 10x for each signature
-  for (y in 1:10){
-    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
-    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],500)
-    #Create FFPE sample vector of 100 mutations
-    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 500)
-    classification.df.1 <- 
-      get_classification_df(list(sample.vec, ffpe.sample.vec), 
-                            c(signature.names[x],"FFPE"), 
-                            list(signature.matrix.list[[x]], ffpe.signature))
-
-    class.df.colnames <- colnames(classification.df.1)
-    sig.name <- class.df.colnames[3]
-    classification.df.2 <- classification.df.1 %>%
-      mutate(signature.indicator = ifelse(truth == sig.name, 1, 0))
-    predictions <- classification.df.2[3]
-    labels <- classification.df.2$signature.indicator
-    pred <- prediction(predictions, labels)
-    auc.perf <- performance(pred, measure = "auc")
-    signature.auc.vec[y] <- auc.perf@y.values[[1]][1]
-  }
-  average.signature.auc <- mean(signature.auc.vec)
-  average.auc.values[x] <- average.signature.auc
-}
-```
-
-```{r, include=FALSE}
-cos.sim.auc.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.auc.values)
-colnames(cos.sim.auc.data) <- c("signature", "cosine_similarity", "average_auc")
-```
-
-```{r, include=FALSE}
-
-ggplot(cos.sim.auc.data, aes(x=average_auc)) +
-  geom_histogram(color="black", fill="lightblue")+
-  labs( x = "Average AUC", y = "Count")
-#ggsave("supp.fig.5.png", dpi = 300, width = 175, height = 108, units = "mm")
-
-median(cos.sim.auc.data$average_auc)
-IQR(cos.sim.auc.data$average_auc)
-
-model2 <- lm(average_auc ~ cosine_similarity, data = cos.sim.auc.data)
-
-model2_summary <- summary(model2)
-
-cos.sim.auc.data$residuals <- model2_summary$residuals
-
-cos.sim.auc.outliers <- cos.sim.auc.data %>%
-  filter(average_auc<0.8)
-
-cos.sim.auc.outliers
-
-cos.sim.auc.outliers2 <- cos.sim.auc.data %>%
-  filter(average_auc>=0.8 & average_auc<0.9)
-
-cos.sim.sbs4 <- cos.sim.auc.data %>%
-  filter(signature=="SBS4")
-
-```
-
-```{r, include=FALSE}
-ggplot(cos.sim.auc.data, aes(x = cosine_similarity, y = average_auc))+
-  geom_point()+
-  geom_smooth(method = "lm", se=FALSE)+
-  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average AUC")+
-  geom_point(data = cos.sim.auc.outliers, color = "red")+
-  geom_point(data = cos.sim.auc.outliers2, color = "purple")+
-  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
-  geom_hline(yintercept=0.9, linetype="dashed", color = "purple")+
-  geom_label_repel(data = cos.sim.auc.outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
-  geom_label_repel(data = cos.sim.auc.outliers2, aes(label = signature), max.overlaps = 10, label.size =   0.1)+
-  geom_label_repel(data = cos.sim.sbs4, aes(label = signature), max.overlaps = 10, label.size =   0.1)
-
-#ggsave("Fig2b.png", dpi = 300, width = 175, height = 108, units = "mm")
-
-```
-
-# Scatterplot of Average Sensitivity and Specificity vs. Cosine Similarity to FFPE
-
-```{r, include=FALSE}
-set.seed(500)
-#outer for loop: loop through each COSMIC v3 signature
-average.sensitivity.values <- c()
-average.specificity.values <- c()
-for (x in 1:length(signature.names)){
-  signature.sensitivity.vec <- c()
-  signature.specificity.vec <- c()
-  #inner for loop: calculate misclassification rate 10x for each signature
-  for (y in 1:10){
-    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
-    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
-    #Create FFPE sample vector of 100 mutations
-    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
-    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
-    class.df.colnames <- colnames(classification.df)
-    sig.name <- class.df.colnames[3]
-    #Count number of true positives (for sensititivity)
-    sensitivity.df <- classification.df %>%
-      filter(truth == sig.name & classify == sig.name)
-    #Calculate sensitivity
-    sensitivity.val = nrow(sensitivity.df)/100
-    #Add to the vector of 10 sensitivity values
-    signature.sensitivity.vec[y] = sensitivity.val
-    #Count number of true negatives (for specificity)
-    specificity.df <- classification.df %>%
-      filter(truth == "FFPE" & classify == "FFPE")
-    #Calculate specificity
-    specificity.val = nrow(specificity.df)/100
-    #Add to the vector of 10 specificity values
-    signature.specificity.vec[y] = specificity.val
-  }
-  average.signature.sensitivity <- mean(signature.sensitivity.vec)
-  average.sensitivity.values[x] <- average.signature.sensitivity
-  average.signature.specificity <- mean(signature.specificity.vec)
-  average.specificity.values[x] <- average.signature.specificity
-}
-```
-
-```{r, include=FALSE}
-sens.spec.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.sensitivity.values, average.specificity.values)
-
-colnames(sens.spec.data) <- c("signature", "cosine_similarity", "sensitivity", "specificity")
-
-ggplot(sens.spec.data, aes(x=sensitivity)) +
-  geom_histogram(color="black", fill="lightblue")+
-  labs( x = "Average sensitivity", y = "Count")
-#ggsave("supp.fig.3.png", dpi = 300, width = 175, height = 108, units = "mm")
-
-median(sens.spec.data$sensitivity)
-IQR(sens.spec.data$sensitivity)
-
-outliers <- sens.spec.data %>%
-  filter(sensitivity<0.8 | signature=="SBS4" | signature=="SBS6")
-
-ref.signatures <- sens.spec.data %>%
-  filter(signature=="SBS4" )
-
-ggplot(sens.spec.data, aes(x = cosine_similarity, y = sensitivity))+
-  geom_point()+
-  geom_smooth(method = "lm", se=FALSE)+
-  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average sensitivity")+
-  geom_point(data = outliers, color = "red")+
-  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
-  geom_label_repel(data = outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)
-
-#ggsave("supp.fig6.png", dpi = 300, width = 175, height = 108, units = "mm")
-
-model.sens <- lm(sensitivity ~ cosine_similarity, data = sens.spec.data)
-
-summary(model.sens)
-
-ggplot(sens.spec.data, aes(x=specificity)) +
-  geom_histogram(color="black", fill="lightblue")+
-  labs( x = "Average specificity", y = "Count")
-#ggsave("supp.fig.4.png", dpi = 300, width = 175, height = 108, units = "mm")
-
-median(sens.spec.data$specificity)
-IQR(sens.spec.data$specificity)
-
-outliers <- sens.spec.data %>%
-  filter(specificity<0.8 | signature=="SBS4" | signature=="SBS6")
-
-ggplot(sens.spec.data, aes(x = cosine_similarity, y = specificity))+
-  geom_point()+
-  geom_smooth(method = "lm", se=FALSE)+
-  labs( x = "Cosine similarity between baseline COSMIC and FFPE signatures", y = "Average specificity")+
-  geom_point(data = outliers, color = "red")+
-  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
-  geom_label_repel(data = outliers, aes(label = signature), max.overlaps = 10, label.size =   0.1)
-
-#ggsave("supp.fig7.png", dpi = 300, width = 175, height = 108, units = "mm")
-
-model.spec <- lm(specificity ~ cosine_similarity, data = sens.spec.data)
-
-summary(model.spec)
-
-```
-
 # Dilution performance characteristics
@@
 ```{r}
-cosmic.sig4 = signature.matrix.list[[4]]
-cosmic.sig6 = signature.matrix.list[[6]]
-ffpe.sig = ffpe.signature
+cosmic.sig4 = signature_matrix_list[[4]]
+cosmic.sig6 = signature_matrix_list[[6]]
+ffpe.sig = ffpe_signature
 ```
@@
 ```
-
-<!-- # Mutational load performance characteristics -->
-
-<!-- ## Mutational load predictions -->
-
-<!-- ```{r} -->
-
-<!-- step_load = seq(50, 1000, by = 50) -->
-
-<!-- MMR.probabilities <- vector(mode = "list", length = length(step)) -->
-
-<!-- MMR.truth <- vector(mode = "list", length = length(step)) -->
-
-<!-- for (x in 1:length(step)){ -->
-
-<!--   sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2) -->
-
-<!--   ffpe.truth <- rep("FFPE", step[x]/2) -->
-
-<!--   sample.sig6 <- create_signature_sample_vector(cosmic.sig6, step[x]/2) -->
-
-<!--   sig6.truth <- rep("MMR", step[x]/2) -->
-
-<!--   test <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"), -->
-
-<!--                                 list(cosmic.sig6, ffpe.sig)) -->
-
-<!--   classification <- test %>% -->
-
-<!--     mutate(trueBinary = ifelse(truth == "MMR",1,0)) -->
-
-<!--   MMR.probabilities[[x]] <- classification$MMR -->
-
-<!--   MMR.truth[[x]] <- classification$trueBinary -->
-
-<!-- } -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- tob.probabilities <- vector(mode = "list", length = length(step)) -->
-
-<!-- tob.truth <- vector(mode = "list", length = length(step)) -->
-
-<!-- for (x in 1:length(step)){ -->
-
-<!--   sample.ffpe <- create_signature_sample_vector(ffpe.sig, step[x]/2) -->
-
-<!--   ffpe.truth <- rep("FFPE", step[x]/2) -->
-
-<!--   sample.sig4 <- create_signature_sample_vector(cosmic.sig4, step[x]/2) -->
-
-<!--   sig4.truth <- rep("tob", step[x]/2) -->
-
-<!--   test <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"), -->
-
-<!--                                 list(cosmic.sig4, ffpe.sig)) -->
-
-<!--   classification <- test %>% -->
-
-<!--     mutate(trueBinary = ifelse(truth == "Tobacco",1,0)) -->
-
-<!--   tob.probabilities[[x]] <- classification$Tobacco -->
-
-<!--   tob.truth[[x]] <- classification$trueBinary -->
-
-<!-- } -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- load.MMR <- prediction(MMR.probabilities, MMR.truth) -->
-
-<!-- load.tob <- prediction(tob.probabilities, tob.truth) -->
-
-<!-- ``` -->
-
-<!-- ## Sensitivity -->
-
-<!-- ```{r} -->
-
-<!-- load.sens.MMR.vec = evaluate_perf(100, length(step_load), load.MMR, "sens") -->
-
-<!-- load.sens.tob.vec = evaluate_perf(100, length(step_load), load.tob, "sens") -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- #plot(step, load.sens.MMR.vec, ylim = c(0,1)) -->
-
-<!-- sensitivity = append(load.sens.MMR.vec, load.sens.tob.vec) -->
-
-<!-- signature.sens = append(rep("MMR", length(load.sens.MMR.vec)), -->
-
-<!--                         rep("Tobacco", length(load.sens.tob.vec))) -->
-
-<!-- dilution = append(step, step) -->
-
-<!-- sens.df <- data.frame(dilution, sensitivity, signature.sens) -->
-
-<!-- ggplot(data = sens.df, mapping = aes(x = dilution,  -->
-
-<!--                                      y = sensitivity,  -->
-
-<!--                                      color = signature.sens)) + -->
-
-<!--   geom_point() + -->
-
-<!--   geom_smooth() + -->
-
-<!--   labs(title = "Sensitivity as a function of mutational load", -->
-
-<!--        x = "Number of mutations in each sample", -->
-
-<!--        y = "Sensitivity value", -->
-
-<!--        color = "True signal") + -->
-
-<!--   coord_cartesian(ylim = c(0,1)) -->
-
-<!-- ``` -->
-
-<!-- ## Specificity -->
-
-<!-- ```{r} -->
-
-<!-- load.spec.MMR.vec = evaluate_perf(100, length(step.load), load.MMR, "spec") -->
-
-<!-- load.spec.tob.vec = evaluate_perf(100, length(step.load), load.tob, "spec") -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-
-<!-- #plot(step, load.spec.MMR.vec, ylim = c(0,1)) -->
-
-<!-- specificity = append(load.spec.MMR.vec, load.spec.tob.vec) -->
-
-<!-- signature.spec = append(rep("MMR", length(load.spec.MMR.vec)), -->
-
-<!--                         rep("Tobacco", length(load.spec.tob.vec))) -->
-
-<!-- dilution = append(step, step) -->
-
-<!-- spec.df <- data.frame(dilution, specificity, signature.spec) -->
-
-<!-- ggplot(data = spec.df, mapping = aes(x = dilution,  -->
-
-<!--                                      y = specificity,  -->
-
-<!--                                      color = signature.spec)) + -->
-
-<!--   geom_point() + -->
-
-<!--   geom_smooth() + -->
-
-<!--   labs(title = "Specificity as a function of mutational load", -->
-
-<!--        x = "Number of mutations in each sample", -->
-
-<!--        y = "Specificity value", -->
-
-<!--        color = "True signal") + -->
-
-<!--   coord_cartesian(ylim = c(0,1)) -->
-
-<!-- ``` -->

### 2022-08-10 sandbox/Bayes Mutation Example.Rmd
Changes to figures including combined effect of FFPE percent and cosine similarity. Cleanup of previous markdowns and files

---
title: "Bayes Question"
author: "Audrey Mitchell"
date: "6/7/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MutationalPatterns)
library(BSgenome)
library(BiocManager)
install("BSgenome.Hsapiens.UCSC.hg19")
ref_genome <- "BSgenome.Hsapiens.UCSC.hg19"
library(ref_genome, character.only = TRUE)
vcf_files <- list.files(system.file("extdata", package="MutationalPatterns"), pattern = ".vcf", full.names = TRUE)
vcf_files <- vcf_files[-7]
sample_names <- c("blood-AC", "blood-ACC55", "blood-BCH", "colon1", "colon2", "colon3", "intestine1", "intestine2", "intestine3", "liver1", "liver2", "liver3")
vcfs <- read_vcfs_as_granges(vcf_files, sample_names, ref_genome)
tissue <- c(rep("blood", 3), rep("colon",3), rep("intestine",3), rep("liver",3))
type_occurrences <- mut_type_occurrences(vcfs, ref_genome)
mut_mat <- mut_matrix(vcf_list = vcfs, ref_genome = ref_genome)
```

```{r, include = FALSE}
nmf_res <- extract_signatures(mut_mat, rank = 2, nrun = 10)
colnames(nmf_res$signatures) <- c("Signature A", "Signature B")
rownames(nmf_res$contribution) <- c("Signature A", "Signature B")
```

# Given that we observe a A(C>T)A mutation in the blood-AC sample, what is the probability that it came from Signature A?

## Method 1: Splitting Pr(A(C>T)A) into A(C>T)A mutations from Signature A and A(C>T)A mutations from Signature B
$Pr(SigA|A(C>T)A) = \frac{Pr(SigA)Pr(A(C>T)A|SigA)}{Pr(SigA)Pr(A(C>T)A|SigA)+Pr(SigB)Pr(A(C>T)A|SigB)}$

```{r, echo = FALSE}
plot_96_profile(nmf_res$signatures, condensed = TRUE)
```
From the plot above we observe that $Pr(A(C>T)A|SigA)= 0.025$ and $Pr(A(C>T)A|SigB)=0.05$.
```{r, echo= FALSE}
plot_contribution(nmf_res$contribution, nmf_res$signature, mode = "relative", coord_flip = TRUE)
nmf_res$contribution
```

```{r}
0.1093118/(0.2800362+0.1093118)
0.2800362/(0.2800362+0.1093118)
```

From the plot above and the contributions of Signature A and B, we see that for the blood-AC sample $Pr(SigA) \approx 0.28$ and $Pr(SigB) \approx 0.72$.\
Then $Pr(SigA|A(C>T)A) = \frac{Pr(SigA)Pr(A(C>T)A|SigA)}{Pr(SigA)Pr(A(C>T)A|SigA)+Pr(SigB)Pr(A(C>T)A|SigB)}=\frac{0.28(0.025)}{0.28(0.025)+0.72(0.05)}=\frac{0.007}{0.043}=0.1628$

## Method 2: Using Total Pr(A(C>T)A) from the mutational profile of the sample
$Pr(SigA|A(C>T)A)=\frac{Pr(SigA)Pr(A(C>T)A|SigA)}{Pr(A(C>T)A)}$

```{r,echo=FALSE}
plot_96_profile(mut_mat[,c(1,7)])
```
In the blood-AC sample, A(C>T)A mutations account for approximately 0.05 of the total mutations.\
$Pr(SigA|A(C>T)A)=\frac{Pr(SigA)Pr(A(C>T)A|SigA)}{Pr(A(C>T)A)}=\frac{0.28(0.025)}{0.05}=\frac{0.007}{0.05}=0.14$

```{r,echo=FALSE}
plot_contribution(nmf_res$contribution, nmf_res$signature, mode = "absolute", coord_flip = TRUE)
```

### 2022-08-10 sandbox/June 25 Presentation Demo.Rmd
Changes to figures including combined effect of FFPE percent and cosine similarity. Cleanup of previous markdowns and files

--- 2021-06-23 June 25 Presentation Demo.Rmd
@@
 #Show Bayes function, explain inputs and outputs
+function(mutation, sample, nmf_res)
 
@@
 rownames(test_nmf$contribution) <- c("Signature 18", "Signature 21")
+
+plot_96_profile(test_nmf$signatures)
 ```
@@
 #269 times more likely in signature 21 than 18
-#Sample 6 is ~100% Signature 21
-extract_all_prob("G[T>C]A", 6, test_nmf)
-#Sample 5 is ~90% Signature 21
-extract_all_prob("G[T>C]A", 5, test_nmf)
-#Sample 4 is ~70% Signature 21
-extract_all_prob("G[T>C]A", 4, test_nmf)
-#Sample 4 is ~50% Signature 21
-extract_all_prob("G[T>C]A", 4, test_nmf)
-#Sample 3 is ~30% Signature 21
-extract_all_prob("G[T>C]A", 3, test_nmf)
-#Sample 2 is ~10% Signature 21
-extract_all_prob("G[T>C]A", 2, test_nmf)
-#Sample 1 is ~0% Signature 21
-extract_all_prob("G[T>C]A", 1, test_nmf)
+
+sample.number <- c(1, 2, 3, 4, 5, 6, 7)
+percent.sig21 <- c(0,10,30,50,70,90,100)
+sig18.prob <- vector(mode = "double")
+sig21.prob <- vector(mode = "double")
+
+for (i in 1:7){
+  probs <- extract_all_prob("G[T>C]A", i, test_nmf)
+  sig18.prob[[i]]<- probs[1][[1]]
+  sig21.prob[[i]]<- probs[2][[1]]
+}
+
+x <- data.frame(sample.number, percent.sig21, sig18.prob, sig21.prob)
+colnames(x) <- c("Sample", "Percent Signature 21", "Signature 18 Posterior Probability", "Signature 21 Posterior Probability")
+view(x)
 ```
@@
 ```{r}
-#G[C>A]A mutation is the most common in Signature 18
-#1287 times more likely in signature 18 than 21
-#Sample 1 is ~100% Signature 18
-extract_all_prob("G[C>A]A", 1, test_nmf)
-#Sample 2 is ~90% Signature 18
-extract_all_prob("G[C>A]A", 2, test_nmf)
-#Sample 3 is ~70% Signature 18
-extract_all_prob("G[C>A]A", 3, test_nmf)
-#Sample 3 is ~50% Signature 18
-extract_all_prob("G[C>A]A", 3, test_nmf)
-#Sample 4 is ~30% Signature 18
-extract_all_prob("G[C>A]A", 4, test_nmf)
-#Sample 5 is ~10% Signature 18
-extract_all_prob("G[C>A]A", 5, test_nmf)
-#Sample 6 is ~0% Signature 18
-extract_all_prob("G[C>A]A", 6, test_nmf)
+sample.number2 <- c(1, 2, 3, 4, 5, 6, 7)
+percent.sig21.2 <- c(0,10,30,50,70,90,100)
+sig18.prob2 <- vector(mode = "double")
+sig21.prob2 <- vector(mode = "double")
+
+for (i in 1:7){
+  probs <- extract_all_prob("G[C>A]A", i, test_nmf)
+  sig18.prob2[[i]]<- probs[1][[1]]
+  sig21.prob2[[i]]<- probs[2][[1]]
+}
+
+y <- data.frame(sample.number2, percent.sig21.2, sig18.prob2, sig21.prob2)
+colnames(y) <- c("Sample", "Percent Signature 21", "Signature 18 Posterior Probability", "Signature 21 Posterior Probability")
+view(y)
 ```
@@
 #Sample 3 is 90% Signature 18 and 10% Signature 21
-extract_all_prob("T[C>T]G", 3, test_nmf)
+extract_all_prob("T[C>T]G", 2, test_nmf)
 #Sample 4 is 90% Signature 21 and 10% Signature 18
-extract_all_prob("T[C>T]G", 4, test_nmf)
+extract_all_prob("T[C>T]G", 6, test_nmf)
 ```
@@
 
+cosmic.sigs <- load_cosmic_matrix()
+
 mapping.stats <- read_tsv("supplied_results/master_statistics_table.txt") %>%
@@
 ```{r}
+#plot_contribution(nmf.res$contribution, nmf.res$signature, mode = "relative", coord_flip = TRUE)
+```
+
+```{r}
+#T[C>A]T is one of the peaks of the POLE mutational signature, MSS_6 is a sample with a high proportion of POLE mutations
 extract_all_prob("T[C>A]T", "MSS_6", nmf.res)
+```
+
+```{r}
+#G[C>T]T is a mutation that is present in all 4 mutational signatures, MSI_MLH1HM_15 is a sample with contributions from all 4
+extract_all_prob("G[C>T]T", "MSI_MLH1HM_15", nmf.res)
 ```
@@
 ```{r}
+#function to load ffpe signature
 load_old_mutational_matrix<- function(filename,mut.order="supplied_results/mut_sig.order.txt") {
@@
 #create ffpe sample vector of 1000 mutations
-ffpe.sample.vector <- create_signature_sample_vector(ffpe.signature.matrix, 1000)
+ffpe.sample.vector <- create_signature_sample_vector(ffpe.signature.matrix, 100)
 ffpe.sample.vector
@@
 #create sample matrix to match ffpe signature
-ffpe.sample.matrix <- create_signature_sample_matrix(ffpe.signature.matrix, 10000)
+ffpe.sample.matrix <- create_signature_sample_matrix(ffpe.signature.matrix, 100)
 ffpe.sample.matrix
@@
 ```
+
+```{r}
+cosmic.sig6 <- as.matrix(cosmic.sigs[, 6])
+sample.sig6 <- create_signature_sample_vector(cosmic.sig6, 200)
+
+ffpe.20mutations <- create_signature_sample_vector(ffpe.signature.matrix, 20)
+
+classification.df <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"), list(cosmic.sig6, ffpe.sig))
+
+#view(classification.df)
+mean(classification.df$misclassification)
+```

### 2022-08-10 sandbox/Varying Signature Function Testing.Rmd
Changes to figures including combined effect of FFPE percent and cosine similarity. Cleanup of previous markdowns and files

--- 2021-06-29 Varying Signature Function Testing.Rmd
@@
 ```{r}
-#Load FFPE signature
-##Function to load FFPE signature
-load_old_mutational_matrix<- function(filename,mut.order="supplied_results/mut_sig.order.txt") {
-
-  mutational.signature = read_tsv(filename)
-  mut_mat = read_tsv(mut.order)
-
-  mut.join = mutational.signature %>%
-    inner_join(mut_mat)
-
-  order.tri = unique(mut.join$ext.context)
-
-  mut.join = mut.join %>%
-    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
-    select(ext.context,prob)
-
-  mut.matrix=as.matrix(sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
-  rownames(mut.matrix) = mut.join$ext.context
-
-  return (t(mut.matrix)[,1:96])
-}
-
 ffpe.signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))

### 2022-08-10 sandbox/firevat/firevat_testing.Rmd
Changes to figures including combined effect of FFPE percent and cosine similarity. Cleanup of previous markdowns and files

---
title: "firevat_testing"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(FIREVAT)
library(tidyverse)
library(bedr)
library(stringr)
library(MutationalPatterns)
library(BSgenome)
ref_genome <- "BSgenome.Hsapiens.UCSC.hg38"
library(ref_genome, character.only = TRUE)
library(ROCR)

source("../simulation.R")
source("../sample_conversions.R")
source("../get_classification_df.R")
source("../bayes_function.R")

seq <- getSeq(Hsapiens, "chr1")

knitr::opts_chunk$set(warning=FALSE)
```

```{r echo=FALSE}
df.sigs

(PCAWG.All.Sequencing.Artifact.Signatures)
run <- function(signature = "SBS6", default = 0) {

  # Run FIREVAT
    results <- RunFIREVAT(vcf.file = sample.vcf.file,
                          vcf.file.genome = 'hg38', # for mouse variants: 'mm10'
                          config.file = config.file,
                          df.ref.mut.sigs = df.sigs,
                          target.mut.sigs = c("SBS6", "SBS4", "ffpe.sig"),
                          sequencing.artifact.mut.sigs = c("ffpe.sig"),
                          output.dir = output.dir,
                          objective.fn = Default.Obj.Fn,
                          num.cores = 2,
                          ga.pop.size = 100,
                          ga.max.iter = 5,
                          ga.run = 5,
                          perform.strand.bias.analysis = FALSE,
                          annotate = FALSE)
}

get_roc_point <- function(df) {
  tp <- df %>%
    filter(truth == classify, classify != "FFPE") %>%
    count()
  tp <- tp$n

  tn <- df %>%
    filter(truth == classify, classify == "FFPE") %>%
    count()
  tn <- tn$n

  fp <- df %>%
    filter(truth != classify, classify != "FFPE") %>%
    count()
  fp <- fp$n

  fn <- df %>%
    filter(truth != classify, classify == "FFPE") %>%
    count()
  fn <- fn$n

  tpr <- tp / (tp + fn)
  fpr <- fp / (fp + tn)

  return(c(fpr, tpr))
}
```

```{r echo=FALSE}
cosmic.sigs <- get_known_signatures()
cosmic.sig6 <- as.matrix(cosmic.sigs[,6])
cosmic.sig4 <- as.matrix(cosmic.sigs[,4])
ffpe.sig <- as.matrix(load_old_mutational_matrix("../ROC_Curves/supplied_results/ffpe.signature.txt", "../supplied_results/mut_sig.order.txt"))

rownames(cosmic.sig6) <- rownames(ffpe.sig)
rownames(cosmic.sig4) <- rownames(ffpe.sig)

df.sigs <- data.frame(GetPCAWGMutSigs()[1:3], cosmic.sigs, ffpe.sig)
rownames(df.sigs) <- seq(1, 96)
```

```{r message=FALSE, warning=FALSE}
sample.sig6 <- create_signature_sample_vector(cosmic.sig6, 100)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)

sample1.df <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR", "FFPE"), list(cosmic.sig6, ffpe.sig))

write_sample_to_vcf(sample1.df, "sample1.vcf", Hsapiens, "chr1")

sample.vcf.file <- "sample1.vcf"
config.file <- "../custom_config.json"
output.dir <- "./"

run("SBS6")
```

```{r}
sample1.firevat.df <- create_classify_df("sample1.vcf", seq)

mean(sample1.firevat.df$misclassification)
mean(sample1.df$misclassification)

roc.point <- get_roc_point(sample1.firevat.df)
get_roc_point(sample1.firevat.df)

sample1.df <- sample1.df %>%
  mutate(binaryTruth = ifelse(truth == "MMR", 1, 0))

sample1.prediction <- prediction(sample1.df$MMR, sample1.df$binaryTruth)
sample1.perf <- performance(sample1.prediction, "tpr", "fpr")
plot(sample1.perf, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7))
points(x=roc.point[1], y=roc.point[2])
```

```{r}
sample.sig6 <- create_signature_sample_vector(cosmic.sig6, 500)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 500)

sample2.df <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR", "FFPE"), list(cosmic.sig6, ffpe.sig))

write_sample_to_vcf(sample2.df, "sample2.vcf", Hsapiens, "chr1")

sample.vcf.file <- "sample2.vcf"
config.file <- "../custom_config.json"
output.dir <- "./"

run("SBS6")
```

```{r}
sample2.firevat.df <- create_classify_df("sample2.vcf", seq)

mean(sample2.firevat.df$misclassification)
mean(sample2.df$misclassification)

roc.point <- get_roc_point(sample2.firevat.df)
get_roc_point(sample2.firevat.df)

sample2.df <- sample2.df %>%
  mutate(binaryTruth = ifelse(truth == "MMR", 1, 0))

sample2.prediction <- prediction(sample2.df$MMR, sample2.df$binaryTruth)
sample2.perf <- performance(sample2.prediction, "tpr", "fpr")
plot(sample2.perf, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7))
points(x=roc.point[1], y=roc.point[2])
```

```{r}
sample.sig4 <- create_signature_sample_vector(cosmic.sig4, 100)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)

signature_cosine_similarity(sample.sig4, cosmic.sig4)
plot_96_profile(as.matrix(table(sample.sig4)))

sample3.df <- get_classification_df(list(sample.sig4, sample.ffpe), c("SIG4", "FFPE"), list(cosmic.sig4, ffpe.sig))

write_sample_to_vcf(sample3.df, "sample3.vcf", Hsapiens, "chr1")

sample.vcf.file <- "sample3.vcf"
config.file <- "../custom_config.json"
output.dir <- "./"

run("SBS4")
```

```{r}
sample3.firevat.df <- create_classify_df("sample3.vcf", seq)

mean(sample3.firevat.df$misclassification)
mean(sample3.df$misclassification)

roc.point <- get_roc_point(sample3.firevat.df)
get_roc_point(sample3.firevat.df)

sample3.df <- sample3.df %>%
  mutate(binaryTruth = ifelse(truth == "SIG4", 1, 0))

sample3.prediction <- prediction(sample3.df$MMR, sample3.df$binaryTruth)
sample3.perf <- performance(sample3.prediction, "tpr", "fpr")
plot(sample3.perf, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7))
points(x=roc.point[1], y=roc.point[2])
```

```{r}
sample.sig4 <- create_signature_sample_vector(cosmic.sig4, 900)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)

sample4.df <- get_classification_df(list(sample.sig4, sample.ffpe), c("SIG4", "FFPE"), list(cosmic.sig4, ffpe.sig))

write_sample_to_vcf(sample4.df, "sample4.vcf", Hsapiens, "chr1")

sample.vcf.file <- "sample4.vcf"
config.file <- "../custom_config.json"
output.dir <- "./"

run("SBS4")
```

```{r}
sample4.firevat.df <- create_classify_df("sample4.vcf", seq)

mean(sample4.firevat.df$misclassification)
mean(sample4.df$misclassification)

roc.point <- get_roc_point(sample4.firevat.df)
get_roc_point(sample4.firevat.df)

sample4.df <- sample4.df %>%
  mutate(binaryTruth = ifelse(truth == "SIG4", 1, 0))

sample4.prediction <- prediction(sample4.df$MMR, sample4.df$binaryTruth)
sample4.perf <- performance(sample4.prediction, "tpr", "fpr")
plot(sample4.perf, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7))
points(x=roc.point[1], y=roc.point[2])
```

```{r message=FALSE, warning=FALSE}
sample.sig6 <- create_signature_sample_vector(cosmic.sig6, 900)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)

sample5.df <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR", "FFPE"), list(cosmic.sig6, ffpe.sig))

write_sample_to_vcf(sample5.df, "sample5.vcf", Hsapiens, "chr1")

sample.vcf.file <- "sample5.vcf"
config.file <- "../custom_config.json"
output.dir <- "./"

run("SBS4")
```

```{r}
sample5.firevat.df <- create_classify_df("sample5.vcf", seq)

mean(sample5.firevat.df$misclassification)
mean(sample5.df$misclassification)

roc.point <- get_roc_point(sample5.firevat.df)
get_roc_point(sample5.firevat.df)

sample5.df <- sample5.df %>%
  mutate(binaryTruth = ifelse(truth == "MMR", 1, 0))

sample5.prediction <- prediction(sample5.df$MMR, sample5.df$binaryTruth)
sample5.perf <- performance(sample5.prediction, "tpr", "fpr")
plot(sample1.perf, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7))
points(x=roc.point[1], y=roc.point[2])
```

```{r}
sample.sig4 <- create_signature_sample_vector(cosmic.sig4, 100)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)

sample6.df <- get_classification_df(list(sample.sig4, sample.ffpe), c("SIG4", "FFPE"), list(cosmic.sig4, ffpe.sig))

write_sample_to_vcf(sample6.df, "sample6.vcf", Hsapiens, "chr1")

sample.vcf.file <- "sample6.vcf"
config.file <- "../custom_config.json"
output.dir <- "./"

run("SBS4")
```

```{r}
# two genotype info with one sample only

sample.sig4 <- create_signature_sample_vector(cosmic.sig4, 100)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)

sample7.df <- get_classification_df(list(sample.sig4, sample.ffpe), c("SIG4", "FFPE"), list(cosmic.sig4, ffpe.sig))

write_sample_to_vcf(sample7.df, "sample7.vcf", Hsapiens, "chr1")

sample.vcf.file <- "sample7.vcf"
config.file <- "../custom_config.json"
output.dir <- "./"

run("SBS4")

# Doesn't work
```

```{r}
# No sample, config looks at info column DP and PROB index 1

sample.sig4 <- create_signature_sample_vector(cosmic.sig4, 100)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)

sample8.df <- get_classification_df(list(sample.sig4, sample.ffpe), c("SIG4", "FFPE"), list(cosmic.sig4, ffpe.sig))

write_sample_to_vcf(sample8.df, "sample8.vcf", Hsapiens, "chr1")

sample.vcf.file <- "sample8.vcf"
config.file <- "custom_config8.json"
output.dir <- "./"

run("SBS4")
```

```{r}

# Look at Position

sample.sig4 <- create_signature_sample_vector(cosmic.sig4, 100)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)

sample9.df <- get_classification_df(list(sample.sig4, sample.ffpe), c("SIG4", "FFPE"), list(cosmic.sig4, ffpe.sig))

write_sample_to_vcf(sample9.df, "sample9.vcf", Hsapiens, "chr1")

sample.vcf.file <- "sample9.vcf"
config.file <- "custom_config9.json"
output.dir <- "./"

run("SBS4")

# DOesn't work

```

```{r}
# No sample, config looks at PROB index 1 and 2

sample.sig4 <- create_signature_sample_vector(cosmic.sig4, 100)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)

sample10.df <- get_classification_df(list(sample.sig4, sample.ffpe), c("SIG4", "FFPE"), list(cosmic.sig4, ffpe.sig))

write_sample_to_vcf(sample10.df, "sample10.vcf", Hsapiens, "chr1")

sample.vcf.file <- "sample10.vcf"
config.file <- "custom_config10.json"
output.dir <- "./"

run("SBS4")
```

```{r}
# One sample looking only at GQ from 30 - 60

sample.sig4 <- create_signature_sample_vector(cosmic.sig4, 100)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)

sample11.df <- get_classification_df(list(sample.sig4, sample.ffpe), c("SIG4", "FFPE"), list(cosmic.sig4, ffpe.sig))

write_sample_to_vcf(sample11.df, "sample11.vcf", Hsapiens, "chr1")

sample.vcf.file <- "sample11.vcf"
config.file <- "custom_config11.json"
output.dir <- "./"

run("SBS4")

# Doesn't work
```

```{r}
# two sample looking only at GQ from 30 - 60

sample.sig4 <- create_signature_sample_vector(cosmic.sig4, 100)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)

sample12.df <- get_classification_df(list(sample.sig4, sample.ffpe), c("SIG4", "FFPE"), list(cosmic.sig4, ffpe.sig))

sample12.df <- sample12.df[sample(nrow(sample12.df)),]

write_sample_to_vcf(sample12.df, "sample12.vcf", Hsapiens, "chr1")

sample.vcf.file <- "sample12.vcf"
config.file <- "custom_config12.json"
output.dir <- "./"

run("SBS4")
```

```{r}
# two sample looking only at GQ from 30 - 60 but random ordering

sample.sig4 <- create_signature_sample_vector(cosmic.sig4, 100)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)

sample13.df <- get_classification_df(list(sample.sig4, sample.ffpe), c("SIG4", "FFPE"), list(cosmic.sig4, ffpe.sig))

sample13.df <- sample13.df[sample(nrow(sample13.df)),]

write_sample_to_vcf(sample13.df, "sample13.vcf", Hsapiens, "chr1")

sample.vcf.file <- "sample13.vcf"
config.file <- "custom_config13.json"
output.dir <- "./"

run("SBS4")
```

```{r}
# two sample looking only at GQ with defined distributiona and random ordering

sample.sig4 <- create_signature_sample_vector(cosmic.sig4, 100)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)

sample14.df <- get_classification_df(list(sample.sig4, sample.ffpe), c("SIG4", "FFPE"), list(cosmic.sig4, ffpe.sig))

sample13.df <- sample14.df[sample(nrow(sample14.df)),]
=
write_sample_to_vcf(sample14.df, "sample14.vcf", Hsapiens, "chr1")

sample.vcf.file <- "sample14.vcf"
config.file <- "custom_config13.json"
output.dir <- "./"

run("SBS4")
```

```{r}
as.matrix(table(sample14.df$mutations))

cos_sim_matrix(as.matrix(table(sample14.df$mutations)), as.matrix(GetPCAWGMutSigs()[7]))

as.matrix(GetPCAWGMutSigs()[7]
```

```{r}
sample.sig4 <- create_signature_sample_vector(cosmic.sig4, 100)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)

sample15.df <- get_classification_df(list(sample.sig4, sample.ffpe), c("SIG4", "FFPE"), list(cosmic.sig4, ffpe.sig))

sample15.df <- sample15.df[sample(nrow(sample14.df)),]

write_sample_to_vcf(sample15.df, "sample15.vcf", Hsapiens, "chr1")

sample.vcf.file <- "sample15.vcf"
config.file <- "../custom_config.json"
output.dir <- ""

run("SBS4")
```

```{r}
sample.sig4 <- create_signature_sample_vector(cosmic.sig4, 26)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 26)

sample16.df <- get_classification_df(list(sample.sig4, sample.ffpe), c("SIG4", "FFPE"), list(cosmic.sig4, ffpe.sig))

sample16.df <- sample16.df[sample(nrow(sample16.df)),]

write_sample_to_vcf(sample16.df, "sample16.vcf", Hsapiens, "chr1")

sample.vcf.file <- "sample13.vcf"
config.file <- "custom_config13.json"
output.dir <- ""

run("SBS4")
```

```{r}
sample.sig4 <- create_signature_sample_vector(cosmic.sig4, 100)
sample.ffpe <- create_signature_sample_vector(ffpe.sig, 100)

sample13a.df <- get_classification_df(list(sample.sig4, sample.ffpe), c("SIG4", "FFPE"), list(cosmic.sig4, ffpe.sig))

sample13a.df <- sample13a.df[sample(nrow(sample13a.df)),]

write_sample_to_vcf(sample13a.df, "sample13a.vcf", Hsapiens, "chr1")

sample.vcf.file <- "sample13a.vcf"
config.file <- "custom_config13.json"
output.dir <- ""

run("SBS4")
```

```{r}
sample13.firevat.df <- create_classify_df("sample13a.vcf", seq)

mean(sample13.firevat.df$misclassification)
mean(sample13.df$misclassification)

roc.point <- get_roc_point(sample13.firevat.df)
get_roc_point(sample13.firevat.df)

sample13.df <- sample13.df %>%
  mutate(binaryTruth = ifelse(truth == "SIG4", 1, 0))

sample13.prediction <- prediction(sample13.df$SIG4, sample13.df$binaryTruth)
sample13.perf <- performance(sample13.prediction, "tpr", "fpr")
plot(sample13.perf, print.cutoffs.at=0.5, text.adj=c(-0.2,1.7))
points(x=roc.point[1], y=roc.point[2])
```

### 2022-08-10 sandbox/plots.Rmd
Changes to figures including combined effect of FFPE percent and cosine similarity. Cleanup of previous markdowns and files

--- 2021-07-27 plots.Rmd
@@
 ```{r}
+#Summary Statistics for cos.sim.auc.data
+summary(cos.sim.auc.data)
+IQR(cos.sim.auc.data$cosine_similarity)
+IQR(cos.sim.auc.data$average_auc)
+```
+
+```{r}
 model2 <- lm(average_auc ~ cosine_similarity, data = cos.sim.auc.data)
@@
          residuals <= quantile(cos.sim.auc.data$residuals, 0.10) | 
-         signature %in% c("SBS4"))
+         signature %in% c("SBS4", "SBS6"))
 ```
@@
   theme(plot.title = element_text(hjust = 0.5))
+
+auc.scatter
 
@@
 sens.spec.label.points <- sens.spec.data %>%
-  filter(signature %in% c("SBS38", "SBS4", "SBS89", "SBS3", "SBS33", "SBS44", "SBS31", "SBS40", "SBS42", "SBS23", "SBS19", "SBS11", "SBS30"))
+  filter(signature %in% c("SBS38", "SBS4", "SBS6","SBS89", "SBS3", "SBS33", "SBS44", "SBS31", "SBS40", "SBS42", "SBS23", "SBS19", "SBS11", "SBS30"))
 ```
@@
 ```
+
+```{r}
+sensitivity.model <- lm(sensitivity ~ cosine_similarity, data = sens.spec.data)
+summary(sensitivity.model)
+```
+
+```{r}
+specificity.model <- lm(specificity ~ cosine_similarity, data = sens.spec.data)
+summary(specificity.model)
+```
+
+```{r}
+auc.model <- lm(average_auc ~ cosine_similarity , data = cos.sim.auc.data)
+summary(auc.model)
+```
+
+```{r}
+summary(sens.spec.data)
+IQR(sens.spec.data$sensitivity)
+IQR(sens.spec.data$specificity)
+```
+
+```{r}
+#outer for loop: loop through each COSMIC v3 signature
+average.accuracies <- c()
+for (x in 1:60){
+  signature.accuracy.vec <- c()
+  #inner for loop: calculate misclassification rate 10x for each signature
+  for (y in 1:10){
+    #create a sample vector of 500 mutations for the COSMIC signature corresponding to number x
+    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],500)
+    #Create FFPE sample vector of 500 mutations
+    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 500)
+    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
+    accuracy = 1 - mean(classification.df$misclassification)
+    signature.accuracy.vec[y] <- accuracy
+  }
+  average.signature.accuracy <- mean(signature.accuracy.vec)
+  average.accuracies[x] <- average.signature.accuracy
+}
+```

### 2022-08-10 sandbox/Mayo Data Analysis/bayes_function.R

@@ -1,0 +1,56 @@
+
+
+extract_all_prob <- function(mutation, sample, nmf_res) {
+  
+  if (class(colnames(nmf_res$signature)) == "NULL") {
+    stop("Column names for nmf_res$signature is empty")
+  } else if (class(rownames(nmf_res$contribution)) == "NULL") {
+    stop("Row names for nmf_res$contribution is empty")
+  }
+  
+  # Creates empty list of length equal to amount of signatures
+  prob = vector(mode = "list", length=length(colnames(nmf_res$signatures)))
+  
+  # initailize accumulator variable and denominator of Bayes Theorem
+  i <- 1
+  denominator = 0
+  
+  # Fills prob with prior and likelihood for each signature
+  for (signature in colnames(nmf_res$signatures)) {
+    # Naming each element of prob
+    names(prob)[i] <- signature
+    
+    # Pr(Mutation|Signature) - likelihood
+    a <- nmf_res$signatures[mutation,signature]/sum(nmf_res$signatures[,signature])
+    
+    # Pr(Signature) - prior
+    b <- nmf_res$contribution[signature,sample]/sum(nmf_res$contribution[,sample])
+    
+    # Appending likelihood and prior into prob
+    prob[[i]] <- c(a, b)
+    
+    # Adding product of likelihood and prior to denominator
+    denominator = denominator + a * b
+    
+    # Accumulating
+    i <- i + 1
+  }
+  
+  # Intialize acummulator and vector to store posteriors
+  i <- 1
+  posteriors <- vector(mode = "list", length = length(colnames(nmf_res$signatures)))
+  
+  # Fills posteriors
+  for (x in colnames(nmf_res$signatures)) {
+    names(posteriors)[i] <- x
+    # Calculates numerator of Bayes formula for each signature
+    numerator <- prob[[x]][1] * prob[[x]][2]
+    # Finishes Bayes calculation and inserts value into the vector posteriors
+    posteriors[[i]] <- numerator / denominator
+    
+    # Accumulating
+    i <- i + 1
+  }
+  
+  return (posteriors)
+}

### 2022-08-10 Mayo Data Analysis/bayes_function.R
moving files to sandbox

@@ -1,56 +1,0 @@
-
-
-extract_all_prob <- function(mutation, sample, nmf_res) {
-  
-  if (class(colnames(nmf_res$signature)) == "NULL") {
-    stop("Column names for nmf_res$signature is empty")
-  } else if (class(rownames(nmf_res$contribution)) == "NULL") {
-    stop("Row names for nmf_res$contribution is empty")
-  }
-  
-  # Creates empty list of length equal to amount of signatures
-  prob = vector(mode = "list", length=length(colnames(nmf_res$signatures)))
-  
-  # initailize accumulator variable and denominator of Bayes Theorem
-  i <- 1
-  denominator = 0
-  
-  # Fills prob with prior and likelihood for each signature
-  for (signature in colnames(nmf_res$signatures)) {
-    # Naming each element of prob
-    names(prob)[i] <- signature
-    
-    # Pr(Mutation|Signature) - likelihood
-    a <- nmf_res$signatures[mutation,signature]/sum(nmf_res$signatures[,signature])
-    
-    # Pr(Signature) - prior
-    b <- nmf_res$contribution[signature,sample]/sum(nmf_res$contribution[,sample])
-    
-    # Appending likelihood and prior into prob
-    prob[[i]] <- c(a, b)
-    
-    # Adding product of likelihood and prior to denominator
-    denominator = denominator + a * b
-    
-    # Accumulating
-    i <- i + 1
-  }
-  
-  # Intialize acummulator and vector to store posteriors
-  i <- 1
-  posteriors <- vector(mode = "list", length = length(colnames(nmf_res$signatures)))
-  
-  # Fills posteriors
-  for (x in colnames(nmf_res$signatures)) {
-    names(posteriors)[i] <- x
-    # Calculates numerator of Bayes formula for each signature
-    numerator <- prob[[x]][1] * prob[[x]][2]
-    # Finishes Bayes calculation and inserts value into the vector posteriors
-    posteriors[[i]] <- numerator / denominator
-    
-    # Accumulating
-    i <- i + 1
-  }
-  
-  return (posteriors)
-}

### 2022-08-10 Mayo Data Analysis/get_classification_df.R

@@ -1,129 +1,0 @@
-
-#Arguments:
-#1) list where each element is a sample generated from a certain signature and
-#   each subelement is a trinucleotide mutation type (as a string)
-#2) vector of strings with names of signatures in the same order as the elements
-#   in the list from argument 1
-#3) list where each element is the true mutational signature vector
-
-
-get_classification_df <- function(samples, sig.names, signatures) {
-  
-  #Empty vectors for names of mutation sources (signatures) and randomly generated samples
-  names <- vector(mode = "character")
-  signature.sample <- vector(mode = "character")
-  
-  for (i in 1:length(samples)) {
-    #Adds each signature name for every element/profile in the sample to empty vector
-    names <- append(names, rep(sig.names[i], length(samples[[i]])))
-    #Adds each element/profile in the sample to empty vector
-    signature.sample <- append(signature.sample, samples[[i]])
-  }
-  
-  #Puts the long generated mutations and their source into a data frame and names columns
-  df.final <- data.frame(signature.sample, names)
-  colnames(df.final) <- c("mutations", "truth")
-  
-  
-  #converts sample vector into a table of probabilities
-  tab <- table(signature.sample)/sum(table(signature.sample))
-
-  #converts table of probabilities to a data frame and defines column names
-  sample.probs.df <- data.frame(tab)
-  colnames(sample.probs.df) <- c("mutations", "frequencies")
-
-  #defines a function that joins two data frames and replaces NA with 0
-  left_join_NA <- function(x, y, by) {
-    left_join(x = x, y = y, by = by) %>% 
-      mutate_each(funs(replace(., which(is.na(.)), 0)))
-  }
-  
-  #creates a vector of mutation types
-  mutations <- rownames(signatures[[1]])
-  
-  #creates a data.frame of mutation types to join with sample probabilities
-  df <- data.frame(mutations)
-
-  #joins together sample probabilities with data frame of mutation types so that all 96 mutation types are included
-  sample.df <- left_join_NA(df, sample.probs.df, by = "mutations")
-
-  #Add pseudocount to column of probabilities 
-  sample.matrix<- as.matrix(sample.df$frequencies) + 0.0001
-  rownames(sample.matrix)= mutations
-  
-  #Computes exposure matrix using sample matrix and true signature matrix
-  nmf_res <- fit_to_signatures(sample.matrix, do.call(cbind, signatures))
-  #Add element with original signatures so that Bayes function accepts this object
-  nmf_res$signatures <-do.call(cbind, signatures)
-  
[…]

### 2022-08-18 paper_analysis.Rmd
Added files for the analysis with real samples

@@
 
+# Using real data
+
+We are going use 6 files from this study ...
+
+## Preprocessing
+
+Let's start by listing the 6 VCF files from the `supplied_results` folder:
+
+```{r}
+library(vcfR)
+library(R.utils)
+library(tidyverse)
+vcf_files <- list.files(path="supplied_results",pattern = "vcf", full.names = TRUE)
+```
+
+We will be subsetting each of those files to the variants for which we have confident information in the DNA. We can assure that by requiring `GQ >30` (Genotype Quality) on the exome side
+
+```{r}
+for (i in 1:length(vcf_files)) {
+  name <- vcf_files[i]
+  vcf <- read.vcfR(name)
+  filter_name <- str_replace_all(name,"vcf","high.vcf.gz")
+
+  gq <-  extract.gt(vcf, element="GQ", as.numeric=TRUE)
+  write.vcf(vcf[which(gq[,1]>30),], file=filter_name)
+  gunzip(filter_name)
+}
+```
+
+## Running excerno
+
+We ran excerno on the 6 vcf files and use the NMF option
+
+```{r}
+library(MutationalPatterns)
+library(BSgenome)
+library(excerno)
+ref_genome <- "BSgenome.Hsapiens.UCSC.hg38"
+library(ref_genome, character.only = TRUE)
+
+vcf_files <- list.files(path="supplied_results",pattern = "high.vcf", full.names = TRUE)
+excerno_vcf(vcf_files,method="nmf",2)
+```
+
+## Calculating performance metrics
+
+We first create a function that allows to collect information on a filtered vcf file
+
+```{r}
+get_metrics <- function (file, sample_name) {
+  vcf <- read.vcfR(file)
+  gt <- extract.gt(vcf)
+
+  # Calculate the errors in the true variants  
+  true_variants <- vcf[which(gt[,1]!="0/0"),]
+  all_true <- as.numeric(dim(true_variants)[1])
+  true_classify <- getFIX(true_variants)[,7]
+  fn <- sum(is.na(true_classify))
+  tp <- all_true -fn
+
+  # Calculate the errors in the false variants
+  false_variants <- vcf[which(gt[,1]=="0/0"),]
+  all_false <- as.numeric(dim(false_variants)[1])
+  false_classify <- getFIX(false_variants)[,7]
+  tn <- sum(is.na(false_classify))
+  fp <- all_false - tn
+  sum(is.na(true_classify))
+
+  # Calculate sensitivity and speficifty
+  (sens <-  tp /all_true)
+  (spec <- tn /all_false)
+  (ffpe_pct <- all_false/(all_false+all_true))
+  (total <- all_false+all_true)
+
+  tibble(name=sample_name, sens=sens, spec=spec, 
+         ffpe.pct = ffpe_pct*100, total=total)
+}
+```
+
+Now we list the classified files from `excerno`
+
+```{r}
+vcf_files <- list.files(pattern = "vcf", full.names = TRUE)
+sample_names = c("MLHG10_10", "MSH6_1","MSH6_2", "MSH6_8", "PMS2_10", "PMS2_09")
+```
+
+And we generate the table with informaton from the samples
+
+```{r}
+1:6 %>%
+  map_dfr(~get_metrics(vcf_files[.x], sample_names[.x])) %>%
+  arrange(desc(ffpe.pct)) %>%
+  mutate_if(is.numeric, format, digits=1)%>%
+  kbl() %>%
+  kable_styling()
+
+```
+
 # Conclusion

### 2022-08-18 paper_analysis.Rmd
Updated analysis with 6 real samples

@@
 
-```{r}
+```{r eval=FALSE}
 for (i in 1:length(vcf_files)) {
@@
   vcf <- read.vcfR(name)
-  filter_name <- str_replace_all(name,"vcf","high.vcf.gz")
+  filter_name <- str_replace_all(name,"vcf","dna.vcf.gz")
+  out_vcf_name <- str_replace_all(name,"vcf", "dna.vcf")
 
@@
   write.vcf(vcf[which(gq[,1]>30),], file=filter_name)
-  gunzip(filter_name)
+  gunzip(filter_name, out_vcf_name)
 }
@@
 
-vcf_files <- list.files(path="supplied_results",pattern = "high.vcf", full.names = TRUE)
+vcf_files <- list.files(path="supplied_results",pattern = "dna.vcf", full.names = TRUE)
 excerno_vcf(vcf_files,method="nmf",2)
@@
   (total <- all_false+all_true)
+  accuracy <- (tp + tn)/(all_true+all_false)
 
-  tibble(name=sample_name, sens=sens, spec=spec, 
-         ffpe.pct = ffpe_pct*100, total=total)
+  tibble(Name=sample_name, `% FFPE`= round(ffpe_pct*100), `# Mutations`=total,
+         Accuracy=accuracy, Specificity=spec, Sensitivity=sens)
 }
@@
 vcf_files <- list.files(pattern = "vcf", full.names = TRUE)
-sample_names = c("MLHG10_10", "MSH6_1","MSH6_2", "MSH6_8", "PMS2_10", "PMS2_09")
+sample_names = c("MLH1G_10", "MSH6_1","MSH6_2", "MSH6_8", "PMS2_10", "PMS2_9")
 ```
 
-And we generate the table with informaton from the samples
+And we generate the table with information from the samples
 
 ```{r}
-1:6 %>%
+(perf_tbl <- 1:6 %>%
   map_dfr(~get_metrics(vcf_files[.x], sample_names[.x])) %>%
-  arrange(desc(ffpe.pct)) %>%
-  mutate_if(is.numeric, format, digits=1)%>%
-  kbl() %>%
-  kable_styling()
+  arrange(desc(`% FFPE`)) %>%
+  mutate(`% FFPE`=str_c(`% FFPE`,"%")) %>%
+  mutate(across(c("Accuracy","Sensitivity", "Specificity"),
+                ~format(.x,digits=2, justify="right")))%>%
+  mutate(across(c(`# Mutations`), ~format(.x,big.mark=",", justify="right"))) %>%
+  kbl(caption = "Table 1. Assessment of excerno on real samples", 
+      align=c("l", rep("r",5))) %>%
+  kable_classic(full_width = F, html_font = "Cambria"))
+
+#save_kable(perf_tbl,"paper_figures/table_1.pdf")
 
